### 2018-10-08 paper.tex
Added initial style and files

\documentclass[conference]{IEEEtran}

\begin{document}
\title{}

\author{
\IEEEauthorblockN{Nicholas Hazekamp}
\IEEEauthorblockA{University of Notre Dame \\
Notre Dame, Indiana 46556 \\
nhazekam@nd.edu}
\and
\IEEEauthorblockN{Douglas Thain}
\IEEEauthorblockA{University of Notre Dame \\
Notre Dame, Indiana 46556 \\
dthain@nd.edu}
}

\maketitle

\begin{abstract}
Something about portability and packaging of software.
\end{abstract}

\section{Introduction}

% Context: 
% Goal: 
% Problem: 
% Solution: 
% Implementation: 
% Evaluation:

\section{Background and Challenges}

\section{Results}

\section{Related Work}

\section{Conclusion}

\section{Reproducibility Data}

In an effort to provide consistent, reproducible results outlined are the
resources utilized in this paper and where they can be found.
Specific commits are mentioned to provide the exact version that was used.

All of these repositories are open source and contain Makefiles
and instructions on how to build and run them.

\section*{Acknowledgment}

\bibliographystyle{IEEEtran}
\bibliography{otherpapers,cclpapers}

\end{document}

### 2018-11-19 paper.tex
Full rough draft

@@
 
+\usepackage{graphicx}
+\usepackage{mathtools}
+\usepackage{framed}
+\usepackage{cleveref}
+
 \begin{document}
-\title{}
+\title{Projecting Global Filesystems into Local Containers for High-throughput Computing on HPC Resources}
+
+\if 0
+
+Abstract
+
+Intro
+- CVMFS & HEP applications
+- NERSC/HPC resources
+- mismatch, new requirements on HPC
+
+Background and Challenges
+- CVMFS requirements
+- no FUSE, sudo, etc. on HPC
+- Frequent updates to CMVFS
+- need for past versions and present
+- version consistency issues/mismatch with shared filesystem semantics
+- DVMFS
+
+-------------------
+
+Identifying Dependencies
+- tracing vs. static analysis
+- represent deps. as a set of files -> equivalence between the two approaches
+- possibility to include more than requested
+- WLCG as testbed for tracing
+- overhead of tracing
+- sampling ratio for live applications
+
+Projecting a Global FS
+- projections vs. images
+- SquashFS images vs. Docker, ...
+- Shrinkwrap
+- quantify image size vs. projection size
+- data, metadata, cache costs
+- creation costs
+
+--------------------
+
+Managing CVMFS Containers at an HPC Site
+- multiple snapshots at different versions
+- cost of large number of container images (storage, distribution, etc.)
+- extreme cases: massive NERSC image vs. loads of tiny containers
+- possibility of replacing an image with an augmented version to serve 2 apps
+- tradeoffs in space, creation time, etc.
+- online problem: create vs. augment
+
+---------------------
+
+PoC Implementation
+- consider (path, ID) to deal with versions
+- Jacard distance metric
+- group images with common parts
+- fast to (approximately) compute
+- putting too many disjoint pieces in an image pushes it farther and farther away
+- LRU to clean up
+
+Evaluation
+- sample workloads
+- image vs. projection vs. cache size
+- space savings
+- image re-use
+- construction time
+- choice of alpha
+
+----------------------
+
+Conclusion
+
+\fi
 
@@
 nhazekam@nd.edu}
+\and
+\IEEEauthorblockN{Tim Shaffer}
+\IEEEauthorblockA{University of Notre Dame \\
+Notre Dame, Indiana 46556 \\
+tshaffe1@nd.edu}
+\and
+\IEEEauthorblockN{Jakob Blomer}
+\IEEEauthorblockN{\&}
+\IEEEauthorblockN{P. Samuel M. Teuber}
 \and
@@
 \begin{abstract}
-Something about portability and packaging of software.
+High energy physics research at CERN has a large and rapidly growing need for computation.
+To bridge the gap, may scientists are looking to HPC resources.
+HPC environments provides a platform for large
+batch computation, similar to the WLCG.
+However, the restrictions on computation,
+such as limited internet access and prohibiting FUSE,
+clash with HEP's reliance on CVMFS as global filesystem
+for software.
+As most LHC jobs uses 
+only a small portion of CVMFS,
+a more fine-grained approach to preparing container images 
+is needed to limit creation time and storage.
+Shrinkwrap introduces a method for creating and
+updating projections, but a framework for managing
+theses is needed.
+...
 \end{abstract}
@@
 
-% Context: 
-% Goal: 
-% Problem: 
-% Solution: 
-% Implementation: 
-% Evaluation:
+High energy physics research at CERN has a large and rapidly growing need for computation.
+With upgrades to the LHC scheduled in the coming years,
+the amount of computation required will increase dramatically.
+The simulation and analysis workloads at CERN are primarily performed using software distributed
+via the CernVM File System (CVMFS).
+CMVFS provides a read-only, distributed file system using HTTP for network transport and extensive local
+caching to rapidly distribute updates in software across the WLCG.
+The Worldwide LHC Computing Grid (WLCG) currently satisfies most of the computational needs of the LHC experiments.
+With the upcoming high luminosity upgrades to the LHC,
+the computational resources needed will nearly double (?? check this) by 2026,
+% On CERN's page it estimate 50-100X need, argonne had a report that was 10-20X
+prompting researchers to turn to available high performance computing (HPC) resources to meet demand.
+
+HPC environments provides a platform for large
+batch computation, similar to the WLCG.
+In contrast with the WLCG's single slot submissions,
+HPC applications typically consist of parallel jobs using several cores up to a full machine for each submission.
+Most HPC systems have site-specific configuration and security requirements.
+These can take the form of limited network availability at computation nodes,
+limited container support,
+and restricted access to privileged system operations (such as FUSE).
+These restrictions present challenges when porting CERN's HEP applications,
+such as difficulties submitting tasks,
+inability to access required software,
+and additional time spent converting between image and container formats.
+
+Dependence on CVMFS to distribute experiment software currently makes it impossible to take advantage of many HPC sites.
+CVMFS requires unrestricted network access and is implemented as a FUSE module,
+which sites often disallow for security and management reasons.
+Some sites, such as NERSC,
+have attempted workarounds for providing CVMFS,
+such as creating large images containing all of the software stored in CVMFS (500GB-1TB) or
+maintaining separate specialize systems for distributing 
+CVMFS, such as DVMFS.
+However, both of these existing solutions require 
+hours to days of attention from system administrators to deploy or update,
+which leads to lag between the available software and the current version.
+Additionally for systems that utilize large container images, large amounts of storage are needed to maintain different versions.
+
+Since each LHC experiment job uses 
+only a relatively small portion of CVMFS,
+a more fine-grained approach to preparing container images would reduce the time spent creating images and allow for more precise control of the software versions in use.
+This could also reduce the storage required at each worker node and the network bandwidth spent transferring images.
+We modified the reference implementation of CVMFS to allow us to capture the filesystem interactions of a live HEP applications.
+Using this live tracing or static analysis,
+we are able to determine exactly which parts of CVMFS must be included in a container image for a particular job.
+We developed a tool, Shrinkwrap, that can \emph{project} CVMFS (or a subset thereof) at a particular version into a traditional filesystem.
+Such a \emph{projection} of CVMFS is suitable for inclusion in a static container image that does not depend on FUSE or network access.
+
+When providing CMVFS at an HPC site,
+the cost of creating, distributing, and storing customized images for a large number of jobs and users imposes additional challenges.
+The na\"{i}ve approach of creating a projection of CVMFS for each job results in significant computation, storage, and bandwidth overhead.
+Acknowledging that jobs frequently have significant overlap in dependencies,
+we propose a metric for determining the similarity between projections.
+This gives the possibility of identifying existing container images that could be re-used to fulfill a new submission.
+In addition, we explore the trade-offs in automatically augmenting existing projections to allow container images to support multiple distinct job types.
+We bring these tools together to propose a potential enhancement to a batch scheduler for making online decisions about whether to create, reuse, or augment projections of CVMFS to efficiently support high-throughput HEP jobs on HPC resources.
+
+%TODO summarize results
 
@@
 
-\section{Results}
+\subsection{CVMFS}
+The CernVM File System (CVMFS) is a globally distributed filesystem designed for providing efficient,
+read-only access to scientific software.
+CVMFS is primarily used for distribution of High Energy Physics (HEP) software used as part of the Large Hadron Collider (LHC) at CERN.
+Researchers at CERN use CVMFS as the primary means of distributing the analysis and simulation software they develop to the Worldwide LHC Computing Grid (WLCG).
+
+Each WLCG site commits to providing some fixed amount of computational resources for analyzing particle collision events and running simulations.
+Worker nodes at each site must be configured to mount CVMFS as a prerequisite to performing any computational tasks.
+Since CVMFS is implemented as a FUSE module,
+WLCG sites are expected to support FUSE on their worker nodes.
+CVMFS also relies on global configuration files,
+requiring administrative involvement in preparing worker nodes for WLCG jobs.
+In addition, sites generally provision some infrastructure to support CVMFS,
+such as caching proxies to reduce load created by local workers on publicly accessible filesystem servers.
+
+As a result, it is difficult to contribute computational resources toward the LHC experiments without making some site-wide provisioning and configuration changes.
+This is not an issue for the WLCG sites,
+as they provide dedicated computing resources and can customize their setups to meet the requirements of CVMFS.
+For a professor attempting to harness a campus cluster,
+on the other hand, these requirements present a serious barrier.
+Parrot, a debugging tool and personal filesystem developed at the University of Notre Dame,
+can be used in cases where privilege on the worker nodes is limited and FUSE is not available but still requires unrestricted network access.
+For computing at scale, it is important to provide a caching proxy as well,
+which is likely to require assistance from administrators.
+
+\subsection{HPC}
+
+With the rapidly increasing demands of the LHC experiments at CERN,
+HPC resources are an appealing source of computing power to supplement the WLCG.
+Leadership class machines such as Summit (?) will by themselves be able to perform a substantial portion of the work of the entire WLCG today.
+Use of national-scale HPC resources across the United States and Europe such as XXX, YYY, ZZZ could be key in meeting the computational demands following the LHC's high-luminosity upgrade.
+
+Unfortunately, HPC sites are generally more restrictive than the previously mentioned campus resources.
+In addition to limited privilege on workers and lack of FUSE support,
+HPC sites often impose restrictions on network activity,
+including outbound firewalls preventing access to CVMFS.
+
+The US collaboration of the ATLAS project is currently taking advantage of computing resources at various supercomputers in the United States including Cori at NERSC.
+For the deployment of their software the \texttt{uncvmfs} utility is used to unpack entire CVMFS repositories into standalone images (usually in ext4 or squashfs format) which in turn can be used to build Shifter or Singularity images.
+While these images can very well be scaled out onto a large number of nodes inside the NERSC infrastructure,
+the image build process takes around 24 hours which makes it difficult to deploy up-to-date versions of the software on a regular basis.
+As these images contain complete copies of everything in CVMFS,
+each takes on the order of terabytes,
+necessitating careful storage management.
+In addition, the process requires administrators for image creation, deployment, and cleanup.
+Currently, this approach is used in production on the ATLAS and CMS experiments.
+As additional experiments want to take advantage of the resources at NERSC,
+the administrative burden of managing multiple CVMFS images on multiple software versions increases accordingly.
+For the experiments to use the same images (saving time and cost at NERSC),
+multiple groups would have to come to agreement about a single common set of software versions to use and adjust their pipelines elsewhere.
+Both solutions introduce new inconveniences and administrative burden.
+
+More broadly, the use of static, manually managed container images breaks user assumptions about the operation of CVMFS.
+When running jobs using CVMFS,
+researchers can specify the version of each repo to use,
+allowing precise control over software and dependencies.
+Each experiment has its own policies and procedures for choosing a software version;
+some may use bleeding-edge analysis software or another experiment may prefer older, stable versions.
+CVMFS is able to provide multiple customized views of the global filesystem without incurring much additional cost.
+CVMFS also supports some more exotic features,
+such as symlink resolution controlled by environment variables.
+CVMFS as a whole does not map well to the semantics of a traditional filesystem.
+Projecting CVMFS into such a filesystem can capture at most a single, static version of a repo and must necessarily eliminate some features,
+such as environment-based symlink resolution at runtime.
+
+% https://indico.cern.ch/event/587955/contributions/2937411/attachments/1683156/2705423/WahidBhimji-CHEP18-NERSC.pdf
+Another similar approach in use at NERSC is DVMFS.
+Since creating static container images containing CVMFS is costly and difficult to customize,
+DVMFS instead uses Cray DVS IO forwarders to mount a copy of CVMFS served over NFS.
+This allows for less administrator involvement,
+since the mirrored copy of CVMFS can be automatically updated on a regular basis.
+After dealing with several implementation problems,
+DVMFS is currently deployed at NERSC as an alternative to static images.
+While this approach affords more flexibility in choice of repos and software versions,
+it increases the job startup time and places additional requirements on the infrastructure.
+To be usable in production, DVMFS requires a sufficient number of DVS servers to be provisioned.
+DVMFS is nonetheless a viable solution as long as the costs of developing/maintaining the service and allocating infrastructure do not grow too high.
+
+\section{File Specifications}
+
+The first step in addressing the high footprint 
+of CVMFS repositories is to provide accurate 
+specifications of the files, libraries, and packages
+needed for each experiment.
+These file specifications can then be used when 
+projecting CVMFS into realized filesystems.
+In general, the ability to project global filesystems 
+hinges on the clarity of the file specification.
+File specifications can be derived from several
+different sources, such as 
+usage traces, static analysis, or package listings.
+The end goal of any of these methods is to 
+create a comprehensive list of needed files and structure.
+The nature of global filesystems flexible behavior
+further extends file specifications by requiring
+not just the files and structure, but also what
+version/revision was used to verify the contents as well.
+
+\subsection{Creation of Specification}
+File specifications can be created several ways,
+with the obvious option being handwritten.
+However, this quickly becomes unfeasible with
+even a moderate number of requirements.
+This leaves the need for an automate method to create
+specifications.
+The first method is to create a sample trace for your
+jobs. 
+Creating a trace from a sample job allows 
+for a listing of actual files used, which
+can be more accurate than handwritten listings.
+As these jobs are typically on the order of thousands
+to millions of jobs, sampling a small subset
+can provide good software coverage with little
+additional overhead. % the computation is still used
+The burden of creating these traces is
+dependant on the method used,
+which could be an LD\_PRELOAD library,
+using Parrot to track used files, % I think parrot was mentioned previously
+or building the tracing in to the application.
+As we will see later, we chose application
+level tracing as it provides more context
+for the use.
+
+The second automated method is to use static
+analysis for discover requirements.
+This would allow the application to be analyzed
+prior to running to determine what packages are
+linked.
+Static analysis can be done at several levels,
+each providing different levels of control 
+and flexibility.
+This could be more language specific in the 
+form of following import lines in python or
+include lines in C.
+The alternative is to use repository specific 
+methods for setting up the environment,
+such as bash scripts for pointing to correct 
+software versions, or 
+analyzing SCRAM\cite{DBLP:journals/corr/cs-OH-0306014} lines
+to determine the configuration.
+With either of these methods the issues
+that arise come from bleeding edge or
+experiment specific code that is not directly
+handled by the configuration.
+There is a significant mix between languages,
+configurations methods, and software organization
+practices even within a single repository,
+which compounds the above issues in this method.
+
+An combination of tracing and static analysis is the
+use of package listings. 
+Package listings provide a high level indication
+of the software needed, and allow for 
+flexibility on the specifying user about the
+exact files needed.
+Using high level packages provides a method
+for handwritten/tuned specifications that 
+include all aspects of the package.
+The downside of using packages is that in
+their flexible, forgiving specification
+they can be bloated and include unnecessary
+files and data.
+Packages also require that there is consistent
+software organization, which often differs
+between repositories.
+
+For the analysis and creation pipeline discussed
+later in the paper, tracing was used to provide
+the file specifications. 
+Tracings flexibility between repositories
+and more precises file inclusion allow
+for a lean package that provides coverage for
+an experiment.
+
+\subsection{Trace Sampling}
+Tracing was selected as it provides a consistent 
+behavior between experiments and use cases.
+This allows the solution to port between applications
+quickly without having to delve in to the applications
+specific configuration.
+As previously mentioned, there are several methods for
+obtaining a trace from a file.
+
+Tracing was implemented in the CVMFS software and
+interacts via the FUSE interface.
+The tracing utility has been implemented in a way that 
+minimizes the overhead if the tracer is deactivated and 
+also minimizes the measurable overhead in the FUSE calls for an activated tracer 
+by delegating the actual log writing to a separate thread to avoid blocks 
+during the accesses to CVMFS. 
+In fact, for a deactivated tracer the entire overhead consists of 
+a single inline if for a subset of the FUSE calls.
+This allows the actual observed CVMFS usage to  be tracked
+and then used for determining the file specification.
+
+Despite assumption that both CVMFS and FUSE will be limited
+on the HPC site, this methods was used as it is reasonable to
+allow for sampling of these jobs on WLCG. 
+WLCG provides the flexibility of using CVMFS and FUSE, without
+having to implement or provision a separate tracing infrastructure
+at each site.
+Additionally, as many of these jobs would have originally been 
+destine to run on WLCG, the sampling sizes are unlikely to 
+create significant overhead.
+Using CVMFS for the tracing allows CVMFS's just-in-time deployment
+of files to patch the environment and give an accurate view without
+the risk of missing files.
+
+Regardless of how the trace was produced,
+there are several ways a trace needs to communicate 
+which files and how they were used.
+Below is a table describing different axes by which a 
+specification can be tracked:
+\begin{table}[h]
+    \centering
+    \begin{tabular}{|c|c|}
+    \hline
+    Number of files & The number of files in a given subtree. \\ \hline
+    Depth of Tree & The largest number of nested directories. \\ \hline
+    Width of Tree & The number of directory splits in a subtree.\\ \hline
+    Size of files & Sum of the files contents in subtree.\\ \hline
+    Actual files & The actual file names.\\ \hline
+    Revision/Release Number & The version of subtree/compatibility. \\ \hline
+    Checksum & The files computed checksum. \\ \hline
+    File type & Directories vs Files.\\ \hline
+    Usage type & File open/stat/directory listing.\\ \hline
+    \end{tabular}
+    \caption{Description of different axes in File specification distance}
+    \label{tab:distance_axes}
+\end{table}
+
+\subsection{Defined Consistency}
+In order for a file specification to accommodate 
+the need for describing the structure and contents
+the specification goes beyond a generic file listing.
+A file specification needs to include enough information
+to establish the directory structure and verify the contents.
+This can be done in several ways, 
+such as computing the checksum of contents,
+using global revision numbers, 
+or running verification software.
+Computed checksums provide a system agnostic approach
+to verification by allowing the checksum to be computed
+at specification and verified before use.
+Computed checksums provide a
+consistency guarantee to the byte level,
+but are often expensive at runtime for large
+production systems.
+This approach is alleviated in systems such as 
+Git and CVMFS as the contents' checksums are 
+computed at ingestion into the repository.
+This method relies on the underlying system computing
+and providing this information for verification.
+
+The use of a global revision number provides a quick reference
+on the state of the repository at the highest level.
+Global revision numbers can instantly verify the expected
+contents, but does not give evidence of compatibility between
+revisions.
+As the revision number if global, any change to the repository
+is reflected in a change to the revision number, which means 
+some software is compatible between revisions.
+This method leaves much to be desired when using common stable
+packages that change irregularly.
+
+Aside from computing and checking the contents of the
+repository or it's files, it is possible to verify the
+software using experiment computation.
+This is a common method used to testing software
+and works equally when the target software provides
+coverage of the features used by the application.
+The difficulty here is in providing tests that
+behave consistently, verify if computation is
+incorrect, and cover the desired features.
+Though it is possible for the submitting user
+to provide test with coverage, application
+developers have a better grasp of it usage
+and how to provide coverage in testing.
+An additional method is to run the application
+on a determined site for initial results and
+compare in the future for verification, but
+this relies on deterministic code which is 
+not always the assumption in HEP computing.
+
+I AM NOT SURE WHICH WE WILL USE. WRAP THIS UP LATER.
+
+\section{Realization of CVMFS Projection}
+
+\subsection{Definition of Projection}
+
+In the case of global filesystems,
+every time the filesystem is used locally
+the flexible global state has to be 
+locked for that concrete use case.
+In the case of CVMFS,
+this includes specifying the date-time
+or revision number of the repository.
+This state is used to determine 
+what the global filesystem's directory
+structure and file contents are, 
+allowing for future use of the same state.
+CVMFS is able to flexibly move
+between these revision using the FUSE module
+to determine the structure and files and
+relies on HTTP to pull any missing files.
+However, with neither the HTTP access or
+FUSE permissions, a method is needed to 
+create and track realizations of this
+state for future use.
+
+We call these realizations of the global filesystem
+projections.
+A projection is a concrete filesystem that can be
+mounted, loaded, or copied into an existing system.
+Projections can then be used to run experiments
+and analysis using the global filesystem without
+the previously mentioned permissions.
+Projections have been used with CVMFS before,
+in the form of unCVMFS and rsync.
+UnCVMFS is a utility that projects an entire
+CVMFS repository, at a set revision, into the local
+filesystem. 
+UnCVMFS utilizes CVMFS's data/file separation
+to deduplicate the underlying data, but requires
+entire repositories to be projected.
+When used with several repositories for large projects,
+unCVMFS can easily consume 500GB for single projection.
+As is often the case, different experiments/analyses 
+require different revisions that may conflict.
+
+Rsync allows for file lists to specify and download
+subsets of CMVFS repositories, but because it
+was designed for general filesystem synchronization
+rsync flattens the files introducing duplicate data.
+This direct representation of the file tree also
+excludes the possibility of shared data repositories
+between different projections.
+
+% Probably need to mention that many applications assume
+% CVMFS software is installed/mounted at /cvmfs
+
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{drawings/shrinkwrap-structure.png}
+\caption{Shrinkwrap Approach}
+\label{figure:shrinkwrap-arch}
+\end{figure}
+
+We built the CVMFS Shrinkwrap utility,
+a solution that leverages the best of both
+unCVMFS and rsync.
+Shrinkwrap uses the CVMFS's client interface to
+create and update CVMFS projections.
+Shrinkwrap uses a CVMFS configuration to specify the
+revision of the repository, a file-specification list
+to determine the repository subset to project, and
+deduplicates the data in a user specified location.
+Shrinkwrap can be configured to share the data
+repository between several repositories and even
+projections of the same repositories.
+This allows for multiple realized projections
+to coexist, with the desired projection
+mounted for use.
+In \Cref{figure:shrinkwrap-arch} you can see an outline
+of the design of the Shrinkwrap utility.
+
+These methods were compared to showcase Shrinkwraps
+behavior and evaluate usage performance.
+The first analysis compares all three methods pulling 
+an entire repository, to allow comparision between the
+three.
+GRAPH OF DEDUPLICATED DATA/FULL REPO SIZE/TIME TO CREATE.
+The second looks at the time to update existing repositories 
+between verison.
+GRAPH UPDATE TIME ON FULL REPO AND MAYBE SUB REPO(SUB REPO ONLY COMPARES RSYNC AND SHRINKWRAP)
+The third looks at hosting several projection at once and
+switching between them.
+COMPARES RSYNC AND SHRINKWRAP. UNCVMFS CANT COMPETE.
+
+\begin{figure}[h]
+\caption{COMPARISON OF SHRINKWRAP unCMVFS and RSYNC}
+\label{figure:projection-comp}
+\end{figure}
+
+\subsection{Packing Projections}
+
+All of the above methods describe utilize CVMFS
+and rely on the internet connection to create
+the projections.
+A core assumption of this paper is that internet
+access is limited at the compute nodes, so a method
+is needed to transform and use these projections
+for computation.
+For this we looked at several methods for providing
+these projections and compared using 
+Container images (images specific to Docker, Singularity, ect.)
+versus SquashFS images.
+The goal is to find a flexible, compatible intermediate
+format that allows projections to be moved between sites
+and technologies.
+For this purpose, DOcker images, Singularity images, and 
+SqaushFS are compared on their time to create,
+storage requirements, and most importantly compatibility
+with several used technologies, namely 
+Docker, Singularity, and Shifter.
+
+The first aspect is the creation of these images.
+From a defined projection of a repository
+what are the steps needed to create each of these image types.
+
+GRAPH COMPARING BUILD TIME
+
+GRAPH COMPARING STORAGE SPACE
+
+TABLE COMPARING COMPATIBILITY AND SUPPORT
+
+%MERGING PROJECTIONS
+
+%INCOMPATIBLE PROJECTIONS
+%It is often the case when using CVMFS
+%an analysis job requires pieces of several repositories.
+%As a result a method for merging and verifying that these
+%repositories are compatible is needed.
+
+When creating these images for analysis jobs its often the case that several
+of these images may be created and used in the system.
+Using Shrinkwrap to organize the repository data and projections,
+the impact of an increasing number of images can be see in 
+\Cref{figure:spec-size}.
+This graph illustrates a new challenge that needs to be addressed.
+As the number of experiments run and projections hosted by Shrinkwrap
+increase. 
+The raw number of images that need to be host increases to an unmanageable
+size.
+A solution is needed for how to manage these images and still
+provide responsive execution of HEP analysis on HPC.
+
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{plots/specification_size.png}
+\caption{Additive effect of host many images.}
+\label{figure:spec-size}
+\end{figure}
+
+\subsection{Use Cases}
+Here we illustrate several real world use cases.
+These use cases are being used as the existing solutions
+were not feasible and a new approach was needed.
+In analyzing these use cases we can see the beginnings of
+the problem outlined in the previous subsection.
+
+\subsubsection{Atlas at NERSC}
+
+We are currently discussing options to integrate our shrinkwrap utility into the HPC deployment processes of ATLAS as we believe that the proposed software solution can help to reduce the overhead of HPC software delivery significantly.
+On the one hand the actual uncvmfs export process can be optimized as the current export workflow loads the entire CVMFS repository in a first step and later on removes certain files in the exported image which were not in fact needed.
+This is something that can be entirely avoided through the definition of exclusion rules passed to the shrinkwrap utility.
+
+On the other hand, we believe that an export client which relies on the libcvmfs library might reduce the total overhead and therefore improve the export performance in comparison to both uncvmfs and rsync (for uncvmfs this is something that we haven’t had yet the chance to evaluate in an experiment).
+As a longer-term goal, it might be useful to start tracing ATLAS simulation workflows on a few machines to collect data that might allow further export optimizations in a later stage through automated specification building.
+
+\subsubsection{Benchmarking Containers}
+For benchmarking purposes, the CERN IT-Department is interested in a ”Benchmark in a Container” solution which allows the distribution of standalone container images for system performance evaluations.
+For such benchmarks the IT-Department typically relies on simulation work- flows provided by CERN’s experiment collaborations.
+To execute these benchmarks in an easily portable and reproducible manner, it is of high interest to package these software workflows into containers.
+Since the workflows’ software stacks are usually stored inside CVMFS repositories, this is a perfect use case for the shrinkwrap utility and its Docker injection functionality.
+
+For the image export the workflow’s resource-needs would be examined through the CVMFS tracing feature in a first step.
+This would then result in a log containing detailed insights on the paths which were accessed by the benchmarking workload.
+In a second step we can then produce a specification based on the tracing log which will allow an easy, efficient export of the necessary files from CVMFS.
+Afterwards the Docker injector can be used to inject the exported files into a prepared benchmarking container stored in a registry.
+Later, the same utility can be used to update the current container to an up-to-date software version or — depending on the exact use case — the utility could even be integrated into a continuous deployment mechanism which produces new benchmarking containers on every software update.
+
+\section{Managing Projections as a Site-wide Resource}
+
+A major queston not addressed by the Shrinkwrap tool is how to manage images after creation.
+In the case of a small number of manually created container images,
+this is not a serious problem.
+Thus for integrating with the existing systems at NERSC,
+Shrinkwrap alone might be sufficient.
+We would also like to support CVMFS as a site-wide service,
+e.g.\ with a batch system plugin that can automatically prepare container images on job submission.
+This both minimizes direct administrator burden and more closely approximates user experience on other resources like the WLCG.
+With static analysis and dynamic tracing of tasks,
+we have ways to choose subsets of CVMFS to project to run specific tasks.
+The user can collect this information themself,
+or tracing, too, could be handled automatically.
+We have seen two approaches to this in section ??.
+If the same types of jobs are also run on other resources such as the WLCG,
+it is possible to instrument a fraction of runs to build a profile of the type of task.
+Alternatively, we could integrate tracing with image creation by automatically running exemplar tasks on the WLCG or on dedicated bridge nodes and collecting traces as necessary.
+Assuming that accurate dependency information is available via some means,
+we are interested in automatically providing suitable projections of CVMFS to users.
+
+\subsection{An initial approach}
+
+The simplest approach for automatically projecting container images is to create a suitable image on each batch job submission.
+Unfortunately, this solution is very wasteful of computing time.
+The average simulation or analysis job runs for XXX seconds,
+while preparing an image for it using Shrinkwrap and mksquashfs takes on average YYY seconds.
+This image creation is a serial step that must be completed before any actual processing occurs.
+In addition, freshly created images must be transferred and stored on any nodes that will use them,
+increasing load on the network and storage infrastructure.
+
+The first step toward improving performance is to re-use previously created container images.
+A single experiment submitting computational tasks is likely to request the same projection of CVMFS,
+at least for jobs submitted around the same time.
+It is straightforward to check for exact equality between projections as long as the system retains the set of files that went into their creation.
+In particular, it is important to retain both the paths and the hash of the file contents or some other unique identifier.
+This ensures that multiple versions of the same file are recognized as distinct entities.
+
+Next, worker nodes need to be able to cache previously used container images.
+Technologies like Docker currently operate this way (often to the chagrin of users and administrators doing scientific computing).
+Since projections of CVMFS can be re-created on demand,
+there is no need to carefully track which container images to keep.
+For our evaluation,
+we chose LRU as a simple caching policy.
+Worker nodes are assumed to have some fixed amount of storage available for caching images.
+Node-local scratch storage is sufficient for temporarily caching these images.
+We also assumed that durable storage is available on the bridge machine for storing the data cache and projections associated with Shrinkwrap,
+as well as a cache of container images derived from these projections.
+Again, we chose LRU as a caching policy for our evaluation,
+though administrators would be free to experiment with different policies.
+
+Under this design,
+on job submission the bridge machine first checks if the same projection had been requested before.
+If so, it arranges to transfer its cached container image to worker nodes that lack that image locally.
+If the bridge node does not have that projection built already,
+it uses Shrinkwrap to create the requested projection,
+then invokes mksquashfs to create the container image.
+Once the bridge node places the newly created image in its cache,
+the process proceeds as in the previous case.
+The architecture of this initial approach is shown in figure ???.
+
+\subsection{Evaluation of the Initial Approach}
+
+We wanted to evaluate the viability of the simple design discussed in the previous section,
+so we analyzed of a stream of actual job submissions to the WLCG.
+Figure ?? gives a summary of the results.
+We observed that for a single experiment at a given point in time,
+the requested projection is fairly stable.
+However, updates to software sometimes overlapped with previous job submissions,
+leading to multiple projected versions of repos at the same time.
+In addition, submissions from different experiments generally request different versions of repos.
+Thus for a site servicing only a single LHC experiment and with ample storage for caching images,
+the simple approach outlined above is sufficient.
+For a more diverse workload like that of the WLCG,
+however, the amount of resources spent creating and storing projected images rapidly increases.
+
+This presents a problem for sites attempting to significantly augment existing computational capacity,
+as they must be able to efficiently support jobs that vary widely both across experiment and time.
+Since NERSC was willing to devote QQ~TB of storage to complete copies of CVMFS for their current container images,
+the storage cost may be bearable.
+It is not safe to assume,
+however, that this will generally be the case.
+Thus we need to find a way to more efficiently handle and store images while still retaining the flexibility of fine-grained per-job customization.
+
+It is important to note that while the cumulative size of the generated container images grows quickly,
+the projections themselves take up much less space.
+This is due to the deduplication feature of the Shrinkwrap tool.
+When creating a projection,
+Shrinkwrap fetches any missing file data and stores it in a common content-addressable cache directory.
+Shrinkwrap then hard links into the data cache to create the projection.
+This allows multiple projections with contents in common to share their underlying storage.
+Thus even with a large number of projections,
+as long as the differences between projections are small the storage space required is modest.
+When creating container images,
+however, this deduplication is broken as it is not possible to share components between disk images.
+When creating each image all data in the projection must be copied in.
+We observed that the overlap between images was in many cases quite large.
+We therefore investigated strategies for deduplication between static disk images,
+as this would greatly reduce the storage and transfer overhead observed here.
+
+\subsection{Quantifying Similarity between Projections}
+
+Because of the difficulties in applying normal data deduplication across multiple disk images,
+we developed an optimization strategy using additional knowledge about the behavior of CVMFS.
+To better understand the overlap and similarity between projections,
+we sought to quantify the degree to which two projections are related.
+This would allow us to identify projections that are ``close'' (for some definition of close) as candidates for optimization.
+We chose the Jaccard distance for this metric because it has a number of useful properties.
+First, it is very simple to compute and operates on general sets.
+For two sets $A$ and $B$,
+the Jaccard distance $d_j$ is defined as follows.
+\[
+d_j(A, B) = \frac{|A \cup B| - |A \cap B|}{|A \cup B|}
+\]
+Second, there exist approximations such as MinHash that can quickly estimate the Jaccard distance between two sets.
+Third, the Jaccard distance is a metric in the mathematical sense,
+making it easier to reason about distances.
+For the purposes of comparisons,
+we represent a projection as a set of files in CVMFS.
+It is not sufficient to represent a file solely by its path,
+as different versions of a repo can have different data at the same path over time.
+Therefore we represent each file as a tuple consisting of (path, hash) where the path is the absolute path of the file and the hash is a unique, content based identifier.
+CVMFS already stores files by the hashes of their contents,
+so this information is readily available.
+
+For this application in particular,
+the Jaccard distance captures several desirable properties of projections.
+In the case of two projections that differ only by one file,
+the Jaccard distance will be small.
+Likewise for a par of projections with nothing in common,
+the distance will be high.
+And since the Jaccard distance is a proper metric,
+we can find clusters of related projections;
+If two projections are both found to be extremely close to a third projection,
+we can conclude that the two are themselves close by the triangle inequality.
+
+\subsection{Image Management as an Online Problem}
+
+\begin{figure*}[t]
+\centering
+\fbox{\includegraphics[width=0.95\textwidth]{drawings/system_architecture.pdf}}
+\caption{System Architecture}
+\label{figure:sys-arch}
+\end{figure*}
+
+Combining the tools and techniques introduced so far,
+we are now ready to address the central problem in managing projected images as a site-wide service.
+Rather than relying on administrators to manually manage images,
+we must define a method for making a decision online as to how to efficiently satisfy the dependency requirements for batch jobs.
+Our method must be suitable for inclusion as part of the setup for batch job scheduling and must be efficient in terms of both computation time and storage space.
+We will extend the initial implementation discussed in section ?? by tracking distances between projections and allowing existing images to be augmented rather than creating new ones for each request.
+\Cref{figure:sys-arch} gives an overview of the structure of this system.
+
+Upon receiving a request for a projection,
+we first hash the new projection according to the MinHash algorithm.
+We always store the hash values computed for every projection we build to allow for fast clustering.
+Next, we use MinHash to approximate the Jaccard distance between the new projection and all projections previously built.
+Since comparisons using MinHash take constant time for a given probability bound independent of the size of the sets involved,
+we can quickly identify cached projections that are similar to the new request.
+
+To decide if two projections are ``close'',
+we define the parameter $\alpha$ as the maximal Jaccard distance between closely related projections.
+Since Jaccard distance is by definition between zero and one,
+$\alpha$ must be in the same range.
+Choice of $\alpha$ is left to the system administrators.
+Very small choices of $\alpha$ require that projections be extremely close before considering them for merging.
+In the extreme case with $\alpha = 0$,
+only \emph{identical} images will be considered close so no images will be merged.
+This results in a larger number of independent images.
+Choosing $\alpha$ to be larger makes it more likely for images to be considered similar and merged.
+This results in more augmented images that serve multiple tasks.
+In the extreme case of $\alpha = 1$,
+\emph{every} pair of images is considered close and merged if possible.
+This results in large container images that have accumulated many projections.
+We explore tradeoffs when choosing $\alpha$ and give recommendations based on observed patterns of activity in section ??.
+
+Having selected a set of candidate projections that are close enough to the new request,
+we next check each in turn to find either a superset of the requested projection or one without version incompatibilities.
+In the case of finding a superset,
+we can simply use the corresponding cached container image.
+Presumably this superset was created by the merging process below,
+so we are done.
+Otherwise, we continue to look for a compatible projection
+As discussed in section ??,
+two projections are incompatible if they both include the same path but point to different data.
+This situation can arise if for example a library in CVMFS is updated in place.
+Old and new projections would include an identical filesystem path,
+but the new version would have different contents.
+This situation can arise in more subtle ways,
+for example with symlinks controlled by environment variables.
+Since projections of CVMFS must statically resolve such links,
+incompatibilities can arise despite paths appearing compatible under CVMFS.
+And candidate projections that are incompatible with the new request are removed from further consideration.
+
+If no compatible projections were found among the candidates,
+we create a container image for the projection as requested and add it to the cache.
+Otherwise, on finding a compatible cached projection the next step is to augment it with the contents of the new request.
+As previously discussed,
+Shrinkwrap can update a projection that was previously created;
+since we keep the projection used to generate each container image in the cache,
+we can simply augment the selected projection with the new files from the request.
+Next, we build a container image based on the newly updated projection.
+The computational cost of this step could become high since we are creating larger combined projections.
+Appropriate choice of $\alpha$, however,
+bounds the distance between the combined projections such that the cost of creating the augmented container is arbitrarily close to cost of simply fulfilling the original request.
+Since the augmented projection is the union of two different projections,
+we can substitute it for both.
+The container image for the augmented projection can replace the previous image,
+so that the net increase in storage space comes only from the newly added files.
+Since a single augmented image now serves two different batch job submissions,
+we avoid making two copies of the common files and data.
+In section ?? we explore the storage and computational costs of this approach in a more concrete setting.
+
+\section{Evaluation}
+
+\begin{table}
+\centering
+% not factually accurate...
+\begin{tabular}{c c}
+CPU & Xeon something 2 \\
+Cores & 12 \\
+Memory & 24 GB \\
+Disk & 1 TB \\
+Image cache & 32 GB \\
+\end{tabular}
+\caption{Description of each of the 24 worker nodes}
+\label{tab:exp-setup}
+\end{table}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/IMG_20181116_101513.jpg}
+\caption{Data and metadata costs with $\alpha=A$}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/IMG_20181116_101907.jpg}
+\caption{Data and metadata costs with $\alpha=B$}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/IMG_20181116_102111.jpg}
+\caption{Data and metadata costs with $\alpha=C$}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/IMG_20181116_102506.jpg}
+\caption{Image size vs. requested projection size with $\alpha=A$}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/IMG_20181116_102623.jpg}
+\caption{Image size vs. requested projection size with $\alpha=B$}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/IMG_20181116_102706.jpg}
+\caption{Image size vs. requested projection size with $\alpha=C$}
+\end{figure}
+
+We implemented the pipeline described in section ?? on a cluster of RR machines at the Center for Research Computing~(CRC),
+a high-performance computing site at the University of Notre Dame.
+\Cref{tab:exp-setup} details the provisioned resources used in this test.
+We used the sequence of jobs from WLCG submission log discussed in section ?? as a realistic source of events for our evaluation.
+For each submission,
+we used the simplified pipeline from section ?? to build an image for running each job.
+We recorded the sizes of the image cache, projections, and data cache over the course of the workload.
+Figure ?? shows the data and metadata costs observed.
+We indeed observed rapid growth of the image cache with much slower increase in the underlying data cache.
+Figure ?? shows the amount of time spent preparing images versus the size of the requested projection.
+We see that preparation time is strongly correlated with the amount of data involved.
+
+Next, we repeated this analysis using the projection pipeline described in section ?? and $\alpha = ??$.
+We chose these values to illustrate the effects of extreme values as well as one we consider to be a good choice.
+The results with $\alpha = W$ are very similar to those of the simplified pipeline.
+This is to be expected,
+since setting $\alpha$ to be very low limits the amount of merging performed.
+Setting $\alpha = 0$ would approximate the simplified pipeline,
+in that little to no merging occurs.
+
+With $\alpha = X$,
+projections were merged very frequently.
+Figure ?? shows the steady increase in image size under this configuration.
+We see a corresponding increase in time spent constructing images in figure ??.
+With $\alpha$ set this high,
+the time spent searching for a compatible image becomes much more significant,
+hence the super-linear increase in build time.
+Approximate Jaccard distance is extremely fast to compute in memory via MinHash without interacting with the filesystem.
+As $\alpha$ is increased,
+there is less filtering from this fast approximation.
+Additional time is spent while Shrinkwrap walks the filesystem on a larger number of candidate images.
+For very homogeneous workloads,
+choice of $\alpha$ has less impact.
+In this case, however,
+different jobs are often incompatible with each other,
+so eagerly merging projections resulted in poorly chosen combinations that prevented better fits in the future.
+Thus we observe that the character of the workload is important in choosing $\alpha$.
+
+Figures ?? and ?? show the same space and time costs, respectively,
+as previously discussed for $\alpha = X$.
+Here we note a much less pronounced increase in image storage cost than figure ??.
+The increase in construction time is also less pronounced than in figure ??.
+We chose this value of $\alpha$ to be more selective than Y to avoid ballooning images but still allow more merging than Z.
+It is difficult to suggest values of $\alpha$ suitable for all cases since the machine configuration, workload, etc. all affect the patterns of behavior.
+Based on experience,
+we have found $\alpha$ values in the range V-W to be acceptable across a number of workloads and configurations.
+
+\begin{table}
+\centering
+\begin{tabular}{c|c c}
+& Cache Life (jobs) & Reuse count \\ \hline
+$\alpha=a$ & small & small? \\
+$\alpha=b$ & big & big \\
+$\alpha=c$ & small? & big \\
+\end{tabular}
+\caption{Median cache life and number of reuses for container images}
+\label{tab:cachelife}
+\end{table}
+
+We also measured the number of times a projection was reused or modified and how long images remained in the cache in an effort to estimate the reusability and relevance of images.
+\Cref{tab:cachelife} tabulates these results.
+With $\alpha = P$,
+images were not reused as often as with the other values.
+We suspect that this is because eager and frequent merges resulted in large images that were often incompatible with new jobs,
+and that drifted away such that they were no longer close enough to any projection to be used.
+The Jaccard distance metric penalizes elements in one set but not the other,
+so as more and more projections are merged,
+the resulting conglomerate moves farther and farther away from any single projection.
+Eventually, it will fall out of use and age out of the cache.
+
+In the other direction,
+the simplified pipeline and $\alpha = Q$ also had many short-lived images.
+This is likely because new images were often/always created for jobs,
+so there was little reuse.
+Due to the large number of images being created,
+images were pushed out of the cache very frequently.
+
+As part of the choice of $\alpha = Z$,
+we tried to maximize the amount of reuse and image lifetimes.
+This has the effect of indirectly minimizing storage and image construction time.
+These are also very easy metrics to track as part of batch submission.
+As such, they provide a simple way for administrators to optimize $\alpha$ to fit the system or workload.
+Some manual tuning of this form is likely necessary for applying this sort of projection pipeline in situations significantly different from our test setup.
 

### 2019-01-17 paper.tex
Update with abstract and more indepth arguments.

@@
 \usepackage{cleveref}
+\usepackage{hyperref}
 
@@
 \author{
+\IEEEauthorblockN{Tim Shaffer}
+\IEEEauthorblockA{University of Notre Dame \\
+Notre Dame, Indiana 46556 \\
+tshaffe1@nd.edu}
+\and
 \IEEEauthorblockN{Nicholas Hazekamp}
@@
 \and
-\IEEEauthorblockN{Tim Shaffer}
-\IEEEauthorblockA{University of Notre Dame \\
-Notre Dame, Indiana 46556 \\
-tshaffe1@nd.edu}
-\and
 \IEEEauthorblockN{Jakob Blomer}
-\IEEEauthorblockN{\&}
-\IEEEauthorblockN{P. Samuel M. Teuber}
 \and
@@
 \begin{abstract}
-High energy physics research at CERN has a large and rapidly growing need for computation.
-To bridge the gap, may scientists are looking to HPC resources.
-HPC environments provides a platform for large
-batch computation, similar to the WLCG.
-However, the restrictions on computation,
-such as limited internet access and prohibiting FUSE,
-clash with HEP's reliance on CVMFS as global filesystem
-for software.
-As most LHC jobs uses 
-only a small portion of CVMFS,
-a more fine-grained approach to preparing container images 
-is needed to limit creation time and storage.
-Shrinkwrap introduces a method for creating and
-updating projections, but a framework for managing
-theses is needed.
-...
+Scientific applications often depend on large, complex software repositories containing libraries, tools, and frameworks in multiple versions and platforms to support many researchers.
+As a practical necessity, software repos are most commonly distributed via shared or global filesystems.
+In moving existing research to HPC environments, however,
+scientists must work around restrictions that can interfere with global software delivery methods.
+To illustrate these challenges,
+we examine CVMFS, the CernVM File System.
+CVMFS is a global, distributed filesystem that provides read-only access to experiment software and configurations for high energy physics research at CERN.
+CVMFS requires a number of network and OS configurations that are incompatible with a restrictive HPC environment.
+In this work, we introduce a set of tools and procedures for \emph{projecting} the global filesystem (or parts thereof) into local container images suitable for HPC sites.
+As part of our evaluation,
+we found that using local approximations of a global filesystem tends to introduce problems of its own in terms of compute and storage overhead.
+We thus examined several declarative and profiling-based approaches to creating more fine-grained projections to fit application requirements,
+and propose a strategy for efficiently managing these projected container images at an HPC site.
+Using our projection tools and image management strategy,
+we implemented a proof-of-concept submission system for running existing HEP analysis tasks in an HPC environment.
 \end{abstract}
@@
 
-High energy physics research at CERN has a large and rapidly growing need for computation.
-With upgrades to the LHC scheduled in the coming years,
-the amount of computation required will increase dramatically.
-The simulation and analysis workloads at CERN are primarily performed using software distributed
-via the CernVM File System (CVMFS).
+Scientific research has a large and rapidly growing need for computation.
+Researchers have a number of options at their disposal to meet those needs,
+including campus-local or organization-local resources,
+cloud providers,
+grid computing,
+and large HPC sites.
+Moving applications between these classes of resources can present problems,
+as the assumptions made during development may no longer hold.
+For example, organization-local resources and some grid sites commonly provide access to shared filesystems at each compute node.
+In addition, there is some flexibility in system configuration to support an organization's primary research activity.
+A new computing environment may be much less open to changes for a specific application,
+especially large, multi-tenant resources such as HPC sites.
+In these cases, it is left to scientists to adapt their application software to the available resources.
+
+Many of our research collaborators have experienced significant frustration in preparing an existing scientific application to run on a new type of resource.
+As a concrete example for this work,
+we consider the challenges in running simulation and analysis workloads at CERN on HPC sites.
+These applications assume several pieces of infrastructure to be in place and require some configuration and customization to the compute environment.
+This is not an issue when running on Worldwide LHC Computing Grid~(WLCG),
+which is a dedicated resource built to meet CERN's substantial computing needs.
+The WLCG currently satisfies most of the computational needs of the LHC experiments at CERN.
+With the upcoming high luminosity upgrades to the LHC,
+the computational resources needed will will increase by at least an order of magnitude by 2026,
+% On CERN's page it estimate 50-100X need, argonne had a report that was 10-20X
+prompting researchers to turn to available high performance computing (HPC) resources to meet demand.
+%In contrast with the WLCG's single slot submissions,
+%HPC applications typically consist of parallel jobs using several cores up to a full machine for each submission.
+
+Most HPC systems, however, have site-specific configuration and security requirements.
+These can take the form of limited network availability at computation nodes,
+limited container support,
+and restricted access to privileged system features (such as FUSE support).
+These new restrictions present challenges when porting existing scientific applications,
+which expect required software to be readily available,
+make a number of assumptions about how the system should be configured,
+and may not be designed with tasks such as converting between image and container formats in mind.
+
+In the case of the simulation and analysis workloads at CERN,
+unrestricted network access and FUSE support are critical for providing global filesystem access.
+Nearly all research software used at CERN assumes the availability of the CernVM File System (CVMFS),
+a distributed global filesystem built for distributing software and configuration data.
 CMVFS provides a read-only, distributed file system using HTTP for network transport and extensive local
 caching to rapidly distribute updates in software across the WLCG.
-The Worldwide LHC Computing Grid (WLCG) currently satisfies most of the computational needs of the LHC experiments.
-With the upcoming high luminosity upgrades to the LHC,
-the computational resources needed will nearly double (?? check this) by 2026,
-% On CERN's page it estimate 50-100X need, argonne had a report that was 10-20X
-prompting researchers to turn to available high performance computing (HPC) resources to meet demand.
-
-HPC environments provides a platform for large
-batch computation, similar to the WLCG.
-In contrast with the WLCG's single slot submissions,
-HPC applications typically consist of parallel jobs using several cores up to a full machine for each submission.
-Most HPC systems have site-specific configuration and security requirements.
-These can take the form of limited network availability at computation nodes,
-limited container support,
-and restricted access to privileged system operations (such as FUSE).
-These restrictions present challenges when porting CERN's HEP applications,
-such as difficulties submitting tasks,
-inability to access required software,
-and additional time spent converting between image and container formats.
-
 Dependence on CVMFS to distribute experiment software currently makes it impossible to take advantage of many HPC sites.
 CVMFS requires unrestricted network access and is implemented as a FUSE module,
-which sites often disallow for security and management reasons.
+both of which sites often disallow for security and management reasons.
+
 Some sites, such as NERSC,
 have attempted workarounds for providing CVMFS,
-such as creating large images containing all of the software stored in CVMFS (500GB-1TB) or
-maintaining separate specialize systems for distributing 
-CVMFS, such as DVMFS.
+such as creating large images several terabytes in size that contain all of the software stored in CVMFS or
+maintaining local mirrors of CVMFS available via site-local shared filesystems.
 However, both of these existing solutions require 
@@
 This could also reduce the storage required at each worker node and the network bandwidth spent transferring images.
-We modified the reference implementation of CVMFS to allow us to capture the filesystem interactions of a live HEP applications.
-Using this live tracing or static analysis,
+It is often possible to determine dependencies for particular applications statically by examining submission commands, setup scripts, etc.
+This approach, while reliable, tends to be labor-intensive and tied to specific experiments and software stacks.
+We therefore modified the reference implementation of CVMFS to allow us to capture the filesystem interactions of a live HEP applications.
+Using live tracing and static analysis,
 we are able to determine exactly which parts of CVMFS must be included in a container image for a particular job.
-We developed a tool, Shrinkwrap, that can \emph{project} CVMFS (or a subset thereof) at a particular version into a traditional filesystem.
+We developed a tool, Shrinkwrap, that can efficiently \emph{project} CVMFS (or a subset thereof) at a particular version into a traditional filesystem.
 Such a \emph{projection} of CVMFS is suitable for inclusion in a static container image that does not depend on FUSE or network access.
 
-When providing CMVFS at an HPC site,
-the cost of creating, distributing, and storing customized images for a large number of jobs and users imposes additional challenges.
-The na\"{i}ve approach of creating a projection of CVMFS for each job results in significant computation, storage, and bandwidth overhead.
-Acknowledging that jobs frequently have significant overlap in dependencies,
-we propose a metric for determining the similarity between projections.
-This gives the possibility of identifying existing container images that could be re-used to fulfill a new submission.
-In addition, we explore the trade-offs in automatically augmenting existing projections to allow container images to support multiple distinct job types.
-We bring these tools together to propose a potential enhancement to a batch scheduler for making online decisions about whether to create, reuse, or augment projections of CVMFS to efficiently support high-throughput HEP jobs on HPC resources.
-
-%TODO summarize results
+While working with projected container images and Shrinkwrap,
+we found that the cost of creating, distributing, and storing customized images for a large number of jobs and users creates unexpected challenges as well.
+Shrinkwrap is designed to efficiently manage projections by deduplicating data in the filesystem.
+Disk images, however, must contain complete copies of all included data.
+The na\"{i}ve approach of creating a projection of CVMFS for each application results in frequent duplication of data and significant overhead in terms of computation and bandwidth.
+Depending on the granularity of the projection,
+image creation time can easily exceed individual job's actual compute time.
+The alternative extreme, i.e.\ building a single large image that satisfies the requirements of all applications,
+forces all jobs to use the same repository revision,
+which can result in incompatibilities across applications.
+Updates to the repository or changes in applications and requirements can again result in large amounts of overhead due to frequently creating, transferring, and storing these very large images.
+
+Observing that jobs frequently have significant overlap in dependencies,
+we designed a system to more efficiently manage container images on job submission.
+First, we defined a metric for determining the similarity between projections.
+This gives the possibility of identifying existing container images that could be re-used to fulfill a new request.
+We also determined that in some cases,
+it is less expensive to augment an existing, similar projection to allow a single container image to support multiple distinct applications.
+We bring these tools together as a proof-of-concept submission system for making online decisions about whether to create, reuse, or augment container images to efficiently support high-throughput HEP jobs on HPC resources.
+
+Compared to existing solutions,
+Shrinkwrap created more compact projections than traditional methods such as rsync copies,
+and allowed for more efficient updates to existing projections.
+Based on the existing projects to produce static CVMFS container images and traces of benchmark applications,
+we examined the sizes and structures of several projected container images.
+Based on these properties,
+we simulated the behavior of our image management system over a large number of job submissions.
+We found that our approach can effectively manage the container images for a continuous stream of submissions with varying software dependencies and versions with less storage overhead than creating a new image per application.
+In addition, our results show that applications can still take advantage of precise dependency and version specifications to reduce image creation and transfer overhead compared to working with large/complete mirrored copies of CVMFS.
 
@@
 
-\section{File Specifications}
-
-The first step in addressing the high footprint 
-of CVMFS repositories is to provide accurate 
-specifications of the files, libraries, and packages
-needed for each experiment.
-These file specifications can then be used when 
+\section{Defining Application Dependencies}
+
+While the entirety of CVMFS is fairly large,
+a single application would never use every piece of data.
+Rather than attempting to provide a complete copy of CVMFS locally,
+which entails significant storage costs,
+we could instead provide a limited subset of the complete filesystem that is sufficient for a particular job.
+This approach is more in line with traditional software packaging practices,
+where software is provided as needed in the form of RPM files or another similar format.
+Working with software packages is simpler in that dependencies are explicitly specified in the package file or database.
+With CVMFS, no such definitive specification exists.
+This issue is common to shared/global filesystems,
+since users expect the complete contents to be available on demand and so devote little time to explicitly defining dependencies and other requirements.
+
+In the case of CVMFS,
+building an accurate 
+specification of the files, libraries, and packages
+needed for each experiment is the first step toward creating usable container images.
+Such file specifications can be used when 
 projecting CVMFS into realized filesystems.
-In general, the ability to project global filesystems 
-hinges on the clarity of the file specification.
+The correctness and storage overhead of the projected filesystem hinges on the clarity and accuracy of the file specification.
 File specifications can be derived from several
@@
 \subsection{Trace Sampling}
-Tracing was selected as it provides a consistent 
+Tracing is attractive as it provides a consistent 
 behavior between experiments and use cases.
+In addition, it is easier to automate and sample arbitrary applications without assistance from each user.
 This allows the solution to port between applications
@@
 
-\subsection{Defined Consistency}
-In order for a file specification to accommodate 
-the need for describing the structure and contents
-the specification goes beyond a generic file listing.
+\subsection{Unique Consistency Requirements}
+CVMFS differs from traditional POSIX filesystems in that previous versions of all content remain available.
+A researcher may specify the precise revision under which an application is known to run.
+Thus the same path may refer to two completely different files for a pair of applications.
+This is beneficial from the standpoint of reliability and reproducibility,
+but complicates the process of projection.
 A file specification needs to include enough information
-to establish the directory structure and verify the contents.
+to establish the directory structure and verify the contents at a specific revision.
 This can be done in several ways, 
@@
 Aside from computing and checking the contents of the
-repository or it's files, it is possible to verify the
+repository or its files, it is possible to verify the
 software using experiment computation.
@@
 
+\begin{table*}[t]
+\begin{center}
+\begin{tabular}{|c|r r r|r r r|}
+\hline
+CVMFS & & ls -lr & & & Shrinkwrap & \\
+Repository & Total Size (B) & inode Count & Time to walk & 
+  Data Files & Deduped Data Files & Dedup Bytes \\ \hline
+%alice-ocdb & 482,642,307,428 & 1,621,661 & 2:22 &
+% 1,621,282 & 1,595 & 24,433,131,203\\
+alice & 3,980,875,747,986 & 27,322,350 & 51:53 & 
+  797,404 & 23,996,541 & 2,098,509,015,172\\
+%atlas-condb & 719,548,941,240 & 11,781 & 0:04 &
+%  10,940 & 229 & 471,492,086 \\
+atlas & 6,195,013,689,056 & 120,598,601 & 3:48:07 &
+ 8,823,179 & 85,554,444 & 3,760,712,059,548 \\
+boss & 120,048,527,315 & 2,418,025 & 4:18 &
+ 256,061 & 1,868,192 & 72,301,955,666\\
+cms & 581,789,406,847 & 26,224,567 & 1:07:30 &
+ 18,667,178 & 101,448,867 & *2,242,658,825,701\\
+cvmfs-config & 16,546 & 9 & 0:00 &
+ 2 & 0 & 0 \\
+geant4 & 297,571,283,648 & 7,128,024 & 10:36 &
+ 323,280 & 6,219,891 & 68,381,969,660 \\
+grid & 30,642,826,259 & 714,235 & 1:17 &
+ 141,622 & 583,923 & 18,804,003,040 \\
+lhcb & 2,134,893,079,440 & 33,736,802 & 1:01:26 &
+ 10,032,339 & 20,406,235 & 368,070,277,634 \\
+na61 & 5,222,967,734 & 25,602 & 0:02 &
+ 8,609 & 15,001 & 2,425,507,431 \\
+sft & 3,996,527,275,237 & 128,608,792 & 4:17:58 &
+ 6,821,699 & 63,210,139 & 2,016,609,810,573 \\ 
+ \hline
+\end{tabular}
+\caption{CVMFS repository sizes.}
+\label{tab:repo-sizes}
+This table shows the size, number of inodes, and the amount
+of duplicate data that is eliminated using CVMFS.
+\end{center}
+\end{table*}
+
+\if 0
 I AM NOT SURE WHICH WE WILL USE. WRAP THIS UP LATER.
+\fi
 
@@
 
-\subsection{Packing Projections}
+\subsection{Packaging Projections}
 
@@
 and technologies.
-For this purpose, DOcker images, Singularity images, and 
+For this purpose, Docker images, Singularity images, and 
 SqaushFS are compared on their time to create,
@@
 GRAPH COMPARING STORAGE SPACE
-
-TABLE COMPARING COMPATIBILITY AND SUPPORT
-
-%MERGING PROJECTIONS
-
-%INCOMPATIBLE PROJECTIONS
-%It is often the case when using CVMFS
-%an analysis job requires pieces of several repositories.
-%As a result a method for merging and verifying that these
-%repositories are compatible is needed.
 
@@
 
-\subsection{Use Cases}
-Here we illustrate several real world use cases.
-These use cases are being used as the existing solutions
-were not feasible and a new approach was needed.
-In analyzing these use cases we can see the beginnings of
-the problem outlined in the previous subsection.
+\subsection{Use Cases of Shrinkwrap}
+%Here we illustrate several real world use cases.
+%These use cases are being used as the existing solutions
+%were not feasible and a new approach was needed.
+%In analyzing these use cases we can see the beginnings of
+%the problem outlined in the previous subsection.
 
@@
 
-A major queston not addressed by the Shrinkwrap tool is how to manage images after creation.
+With static analysis and dynamic tracing of tasks,
+we have methods to choose subsets of CVMFS to project to run specific tasks.
+Users can collect this information themselves,
+or tracing, too, could be handled automatically.
+We have seen two approaches to this in section ??.
+For applications already running on the WLCG,
+it is also possible to instrument a fraction of runs to build a profile of the type of task.
+Alternatively, we could integrate tracing with image creation by automatically running a small number of sample tasks on the WLCG or on dedicated bridge nodes and collecting traces as necessary.
+These approaches give a fair amount of flexibility in determining accurate dependency information.
+Based on this information,
+we are interested in automatically providing suitable projections of CVMFS to users.
+
+Shrinkwrap provides a mechinsm for projecting parts of CVMFS,
+but is not by itself a complete solution as it does not integrate with job submission or manage images after creation.
 In the case of a small number of manually created container images,
 this is not a serious problem.
+Administrators could manage a small collection of images in this case and adjust jobs to work with their setup.
 Thus for integrating with the existing systems at NERSC,
 Shrinkwrap alone might be sufficient.
-We would also like to support CVMFS as a site-wide service,
-e.g.\ with a batch system plugin that can automatically prepare container images on job submission.
+To support CVMFS as a site-wide service with a potentially varied user base and workload,
+a more automated solution
+(e.g.\ a batch system plugin that can automatically prepare container images on job submission)
+is required.
 This both minimizes direct administrator burden and more closely approximates user experience on other resources like the WLCG.
-With static analysis and dynamic tracing of tasks,
-we have ways to choose subsets of CVMFS to project to run specific tasks.
-The user can collect this information themself,
-or tracing, too, could be handled automatically.
-We have seen two approaches to this in section ??.
-If the same types of jobs are also run on other resources such as the WLCG,
-it is possible to instrument a fraction of runs to build a profile of the type of task.
-Alternatively, we could integrate tracing with image creation by automatically running exemplar tasks on the WLCG or on dedicated bridge nodes and collecting traces as necessary.
-Assuming that accurate dependency information is available via some means,
-we are interested in automatically providing suitable projections of CVMFS to users.
+Our first attempt at such a system,
+described in the next sections,
+took the simplest possible approach to image.
+As we were developing this approach, however,
+a number of issues in real-world usage led us to conculde that a somewhat more sophisticated solution was required.
 
@@
 The simplest approach for automatically projecting container images is to create a suitable image on each batch job submission.
-Unfortunately, this solution is very wasteful of computing time.
-The average simulation or analysis job runs for XXX seconds,
-while preparing an image for it using Shrinkwrap and mksquashfs takes on average YYY seconds.
-This image creation is a serial step that must be completed before any actual processing occurs.
-In addition, freshly created images must be transferred and stored on any nodes that will use them,
-increasing load on the network and storage infrastructure.
-
-The first step toward improving performance is to re-use previously created container images.
+Unfortunately, this solution is very wasteful of computing time and storage space.
+Even a minimal image consisting of a precise set of required libraries and frameworks ranges in the tens to hundreds of gigabytes,
+depending on the applicaion.
+The full-repo images being tested at NERSC are multiple terabytes in size and take the majority of a day to create.
+In addition, image creation must be completed before any actual processing occurs,
+i.e.\ there is no way to start with a partial image that supports the first steps of a pipeline and finish building the image as the pipeline runs.
+There is also the cost of transferring and storing freshly created images on any nodes that will use them,
+further delaying job dispatch and increasing load on the network and storage infrastructure.
+
+The disk I/O alone required to create even a targeted image is likely to take more time than a single simulation or analysis job.
+Thus a simple performance improvement is to re-use previously created container images for multiple jobs.
 A single experiment submitting computational tasks is likely to request the same projection of CVMFS,
@@
 It is straightforward to check for exact equality between projections as long as the system retains the set of files that went into their creation.
-In particular, it is important to retain both the paths and the hash of the file contents or some other unique identifier.
-This ensures that multiple versions of the same file are recognized as distinct entities.
-
-Next, worker nodes need to be able to cache previously used container images.
-Technologies like Docker currently operate this way (often to the chagrin of users and administrators doing scientific computing).
-Since projections of CVMFS can be re-created on demand,
-there is no need to carefully track which container images to keep.
-For our evaluation,
-we chose LRU as a simple caching policy.
+Note that it is important to retain both the paths and the hash of the file contents used in images.
+This ensures that multiple versions of the same file over time are recognized as distinct entities.
+If there are few CVMFS users at a site,
+this method could be sufficient.
+
+After creating a set of static images,
+the next step is to transfer them to local caches on the workers.
 Worker nodes are assumed to have some fixed amount of storage available for caching images.
-Node-local scratch storage is sufficient for temporarily caching these images.
-We also assumed that durable storage is available on the bridge machine for storing the data cache and projections associated with Shrinkwrap,
-as well as a cache of container images derived from these projections.
-Again, we chose LRU as a caching policy for our evaluation,
-though administrators would be free to experiment with different policies.
+If the size of the ``working set'' of images is smaller than the node-local storage,
+then each worker can keep a complete set of images.
+While ideal, this case is not necessarily realistic beyond a few images.
+With small/medium images in the tens-to-hundreds of gigabytes range,
+maintining a complete set of projected CVMFS images at each worker requires substantial resource provisioning.
+Thus it is important for workers be able to to cache only a subset of the available projected images.
+There are numerous caching policies and options available,
+but choice of worker cache strategy is not critically important for this application.
+Since projections and images of CVMFS can be re-created on demand,
+cache evictions might result in computation overhead but would not otherwise affect the functionality or correctness of the system.
+In the absence of a compelling and site-specific reason,
+a simple caching policy such as LRU would be sufficient.
+As we will see in section ??,
+LRU will also be an important piece of our final proposed approach.
 
@@
 
-\subsection{Evaluation of the Initial Approach}
-
-We wanted to evaluate the viability of the simple design discussed in the previous section,
-so we analyzed of a stream of actual job submissions to the WLCG.
-Figure ?? gives a summary of the results.
-We observed that for a single experiment at a given point in time,
-the requested projection is fairly stable.
-However, updates to software sometimes overlapped with previous job submissions,
-leading to multiple projected versions of repos at the same time.
-In addition, submissions from different experiments generally request different versions of repos.
-Thus for a site servicing only a single LHC experiment and with ample storage for caching images,
-the simple approach outlined above is sufficient.
-For a more diverse workload like that of the WLCG,
-however, the amount of resources spent creating and storing projected images rapidly increases.
-
-This presents a problem for sites attempting to significantly augment existing computational capacity,
-as they must be able to efficiently support jobs that vary widely both across experiment and time.
-Since NERSC was willing to devote QQ~TB of storage to complete copies of CVMFS for their current container images,
-the storage cost may be bearable.
-It is not safe to assume,
-however, that this will generally be the case.
-Thus we need to find a way to more efficiently handle and store images while still retaining the flexibility of fine-grained per-job customization.
-
-It is important to note that while the cumulative size of the generated container images grows quickly,
-the projections themselves take up much less space.
-This is due to the deduplication feature of the Shrinkwrap tool.
-When creating a projection,
-Shrinkwrap fetches any missing file data and stores it in a common content-addressable cache directory.
-Shrinkwrap then hard links into the data cache to create the projection.
-This allows multiple projections with contents in common to share their underlying storage.
-Thus even with a large number of projections,
-as long as the differences between projections are small the storage space required is modest.
-When creating container images,
-however, this deduplication is broken as it is not possible to share components between disk images.
-When creating each image all data in the projection must be copied in.
-We observed that the overlap between images was in many cases quite large.
-We therefore investigated strategies for deduplication between static disk images,
-as this would greatly reduce the storage and transfer overhead observed here.
+%Node-local scratch storage is sufficient for temporarily caching these images.
+%We also assumed that durable storage is available on the bridge machine for storing the data cache and projections associated with Shrinkwrap,
+%as well as a cache of container images derived from these projections.
+%Again, we chose LRU as a caching policy for our evaluation,
+%though administrators would be free to experiment with different policies.
+
+%Technologies like Docker currently operate this way (often to the chagrin of users and administrators doing scientific computing).
+
+\subsection{Problems with the initial approach}
+
+In developing our initial approach,
+we found that a number of real-world usage patterns are poorly handled.
+A small change to a projection,
+such as adding a library or updating dependencies,
+results in the creation of a second, unrelated image.
+Even a small change such as removing an unnecessary documentation file results in a completely new container image.
+This is a manageable nuisance when manually testing or developing a system,
+but potentially results in problematic proliferation of images in a system managing images automatically and receiving updated specifications over time from tracing jobs.
+
+In addition to unnecessary duplication for a single user/application,
+there is also significant duplication across multiple users.
+In the simulation and analysis pipelines at CERN,
+a number of large components such as ROOT are used nearly universally.
+These components provide basic functionality and infrastructure that each experiment's software stack builds upon.
+Moreover, these same common infrastructure components must be included regardless of user or application.
+Thus even if every user of the system only creates a single container image,
+there is a duplicate copy of all such common infrastructure components for each user.
+This is problematic if we would like to design a system that can support many users at a site.
+
+These two sources of duplication tend to compound each other,
+leading us to conclude that a more sophisticated approach is necessary to work with realistically-provisioned workers and system infrastructure.
+These issues are fundamentally related to unnecessary duplication of data among container images.
+The same types of issues motivate the use of dynamic libraries for everyday applications.
+With statically linked executables,
+each application includes a complete set of required libraries.
+While very stable and reliable,
+OS distributions prefer to use dynamic linking.
+This allows applications to share a single copy of each library,
+so that foundational components such as libc are not copied into every executable,
+and so that updates and changes to the set of installed packages do not require rebuilding the entire system.
+For this approach to work,
+package maintainers must arrange all software packages in the system to ensure compatibility,
+and the OS distribution must set policies for system organization, update procedures, etc.
+Container images here function similarly to statically linked executables,
+but since multiple complete OS distributions are stored within CVMFS,
+and since we are dealing with multi-terabyte repos and entire frameworks and software stacks rather than individual OS libraries,
+the scale involved is of course larger.
+In the case of CVMFS, there is no notion of a ``package mainainer'' or central authority over all software.
+Each experiment manages its own software repo as it chooses,
+with some components and repos shared between multiple applications.
+Since we would like to be able to automatically handle existing applications that are not designed with central management in mind,
+we must instead look at other approaches for dealing with data duplication.
+
+\subsection{Deduplicating Container Images}
+
+The representation of projections as disk images is problematic for deduplicating data.
+Internally, Shrinkwrap uses hard links to a common data store so that a single file is only stored once across all projections in the system.
+On creating disk images, however,
+these links are broken.
+Each disk image must contain a complete copy of all data,
+so that a file must be included (and take up space) in every image that includes it.
+Docker's solution is to build container from layers of archives that can be reused.
+Docker layers are explicitly defined when writing a Dockerfile.
+With CVMFS, however, there is no existing layer-like organization or universal rule that could derive layers for existing applicaions.
+Working with layers defined from the outset is also a poor fit if we would like to support partial, gradually improved information from traces.
+We therefore desire some sort of deduplication that can be applied for existing applications to an assortment of container images after creation.
+
+To develop a deduplication strategy,
+we started with some observations about the projection and image properties.
+For ``large'' images,
+there is limited opportunity for deduplication.
+Each large image is a fairly complete copy of a particular repo.
+While there is bound to be some overlap between versions of the same repo,
+an updated image will contain a large number of updated software components,
+making a ``deduplicated'' between two versions significantly larger than either individually.
+There are also likely to be some conflicts in the filesystem if software components re-use the same paths for different files between versions.
+This case is akin to combining RHEL~6 and RHEL~7 in an attempt to save space.
+While there are similarities between the two,
+the significant changes and conflicts make this a poor strategy.
+Thus for our strategy we would like to avoid ``deduplicating'' images that have large non-overlapping portions.
+
+Likewise, there is little reason to attempt to deduplicate large images built from different repos.
+Such images would have little to no overlap,
+so the only benefit to combining the two would be reducing the number of images by one.
+This comes at the cost of needlessly storing and transferring the unused half.
+This case is akin to bundling a complete Debian installation alongside RHEL.
+Our deduplication strategy should only take effect for images with significant overlap.
+
+It is also undesirable to combine images of drastically different size.
+As an example, suppose a user creates a small projection as a subset of a repo for running a specific application.
+This small image overlaps completely with a large image of the whole repo,
+and there is likely no conflict.
+Automatically combining the large and small image, however,
+can have unintended consequences.
+Small, tailored images minimize transfer and storage overhead,
+and can be important for operating at scale.
+It is therefore unwise to surprise users by suddenly making their container images several orders of magnitude larger.
+This is akin to distributing a library exclusively by placing it within a disk image of a Debian installation.
+We would therefore like to attempt deduplication only for
+\begin{itemize}
+\item similarly-sized images.
+\item with large overlap.
+\item and with no conflicting paths.
+\end{itemize}
 
@@
 
-Because of the difficulties in applying normal data deduplication across multiple disk images,
-we developed an optimization strategy using additional knowledge about the behavior of CVMFS.
-To better understand the overlap and similarity between projections,
-we sought to quantify the degree to which two projections are related.
+Having set out the conditions under which to consider deduplicating projected images,
+we are in a position to define a metric to quantify the similarity between images.
 This would allow us to identify projections that are ``close'' (for some definition of close) as candidates for optimization.
-We chose the Jaccard distance for this metric because it has a number of useful properties.
-First, it is very simple to compute and operates on general sets.
+We chose the Jaccard distance under appropriate choice of set elements as it conveniently satisfies the desired qualities set out above,
+and is very well used and studied.
+In our case, we must describe a container image as consisting of a set of (path, content hash) tuples.
+CVMFS already stores files by the hashes of their contents,
+so this information is readily available.
+Considering both the file names and contents is important as CVMFS allows for updates to file contents over time.
+We chose fine-grained tracking of file contents rather than simply recording the version of the CVMFS repo to afford additional opportunities for dedupliaction.
+Since each update to a repo does not change the contets of every (or even the majority of) files,
+data from multiple repo versions may well be able to coexist in a single container image.
+
 For two sets $A$ and $B$,
@@
 \[
-d_j(A, B) = \frac{|A \cup B| - |A \cap B|}{|A \cup B|}
+d_j(A, B) = 1 - \frac{|A \cap B|}{|A \cup B|} = \frac{|A \cup B| - |A \cap B|}{|A \cup B|}
 \]
-Second, there exist approximations such as MinHash that can quickly estimate the Jaccard distance between two sets.
-Third, the Jaccard distance is a metric in the mathematical sense,
-making it easier to reason about distances.
+The Jaccard distance is a metric in the mathematical sense,
+making many intitive assumptions about distance and relationships between elements mathematically sound.
 For the purposes of comparisons,
-we represent a projection as a set of files in CVMFS.
-It is not sufficient to represent a file solely by its path,
-as different versions of a repo can have different data at the same path over time.
-Therefore we represent each file as a tuple consisting of (path, hash) where the path is the absolute path of the file and the hash is a unique, content based identifier.
-CVMFS already stores files by the hashes of their contents,
-so this information is readily available.
 
@@
 And since the Jaccard distance is a proper metric,
-we can find clusters of related projections;
-If two projections are both found to be extremely close to a third projection,
+related projections will tend to occur in clusters;
+if two projections are both found to be extremely close to a third projection,
 we can conclude that the two are themselves close by the triangle inequality.
+Other choices of metric may well be viable,
+but we found the Jaccard distance to be both well-studied and amenable to efficient approximate impelementation.
+As described in section ??,
+we used MinHash~\cite{minhash} as a constant-time approximation of the Jaccard metric as a first pass at selecting similar images.
+Such an approximation is important in practice due to the sizes of the data involved;
+metadata listings alone (not including any file data) consumed multiple gigibytes of storage in some of the container images we evaluated.
 
@@
 $\alpha$ must be in the same range.
-Choice of $\alpha$ is left to the system administrators.
-Very small choices of $\alpha$ require that projections be extremely close before considering them for merging.
+This $\alpha$ parameter is something like the ``globbiness'' of the system.
+Very small choices of $\alpha$ requires that projections be extremely close before considering them for merging.
 In the extreme case with $\alpha = 0$,
@@
 incompatibilities can arise despite paths appearing compatible under CVMFS.
-And candidate projections that are incompatible with the new request are removed from further consideration.
+Any candidate projections that are incompatible with the new request are removed from further consideration.
 
@@
 we avoid making two copies of the common files and data.
-In section ?? we explore the storage and computational costs of this approach in a more concrete setting.
+
+One potential issue with this merging strategy is that there is no way to prevent poor merging decisions.
+In the case of a one-off submission,
+the system might merge an existing, frequently used image.
+This is not problematic for the one-off job itself,
+but increases overhead for subsequent jobs.
+The original, frequently used image would increase in size with no additional benefit to other tasks.
+Images might become ``bloated'' in this way,
+accumulating infrequently used dependencies and increasing overhead indefinitely for future tasks.
+We can take advantage of LRU here to ensure that bad merges are eventually removed.
+Since limited storage is available for container images,
+our original approach removed infrequently used images by LRU when necessary.
+We thus define the age of an image to be the maximum of the ages of its constituent images.
+In the case of the frequently used container image merged with a one-off job,
+the one-off job would define the age of the container image since it is never used again.
+Despite having frequently used components,
+this image would be one of the first candidates for removal.
+Removing old images in this way therefore helps to limit the increased overhead of bad merging decisions.
+For a merged image to remain in cache indefinitely,
+its components would need to all be used regularly.
 
@@
 
-\begin{table}
-\centering
-% not factually accurate...
-\begin{tabular}{c c}
-CPU & Xeon something 2 \\
-Cores & 12 \\
-Memory & 24 GB \\
-Disk & 1 TB \\
-Image cache & 32 GB \\
-\end{tabular}
-\caption{Description of each of the 24 worker nodes}
-\label{tab:exp-setup}
-\end{table}
+To evaluate the viability of our approach in automatically deduplicating a container image store,
+we simulated the behavior of the system over time for a range of $\alpha$ values.
+Choice of $\alpha$ is important here to ensure that common components can be detected and merged in container images,
+and that old images or poor decisions are eventually removed from the system.
+Choosing $\alpha$ too far in either direction will result in excessive storage overhead due to duplicated components,
+or excessive compute and network overhead repeatedly merging largely unrelated images.
+Our goal in this evaluation is therefore to choose $\alpha$ so as to minimize the storage and compute costs associated with maintaining a collection of images.
+
+\subsection{Simulating Job Submissions}
+
+The first step in simulating system behavior over time was generating synthetic requests for images.
+We consulted with the developers of CVMFS,
+as well as HEP researchers at the University of Notre Dame collaborating with CERN.
+Since CVMFS is not currently widely deployed on HPC resources,
+the prospective user base for our container management system is limited.
+We were, however, informed of some researchers and groups experimenting with static mirrors of CVMFS for container images,
+including projects at NERSC and Stanford University,
+with one of the setups available publicly at \url{https://github.com/wyang007/atlas-fat-container}.
+Based on correspondence with researchers and existing applications running on the WLCG,
+we made some general observations about the behaviors and requirements of applications we need to support.
+First, there are some components such as ROOT that serve as common, foundational components to almost every other simulation and analysis stack.
+These core components are often stored separately from the rest of the software collections available in CVMFS.
+Second, there are shared components such as setup scripts, configuration files, and calibration or condition data (required to interpret readings from detectors) that are also nearly universally required but are more particular to a given experiment or group.
+Finally, there is the collection of software maintained by each experiment.
+Software collections may contain multiple copies of the same software for different architectures and versions.
+A single analysis job does not use the majority of the available software.
+There is some overlap among these categories of software,
+for example some software components are widely enough used to be considered core components,
+and usage and availability of components varies over time.
+
+We use these observations about the structure and behavior of applications on CVMFS to a simplified simulation of dependency requirements for evaluating our container image management scheme.
+For the purpose of simulation,
+we assume that the requirements of a job are given as discrete components or collections of files.
+While Shrinkwrap can operate at the granularity of individual files,
+both static analysis and runtime tracing produce file collections to include.
+We are intentionally somewhat vague here to allow for both explicitly defined software components and tracing outputs.
+We examined the software collections available in the ATLAS repo and determined that the median size of a software component around 150~MB,
+which we used as the baseline for simulated components.
+This is a significant simplification,
+as CVMFS hosts some very large components and files.
+Looking at the distribution of file sizes we measured for the major repos (figures M-N),
+we observed that these extreme outlier files are fairly rare.
+Our next assumption is that certain components are included in most container images.
+While multiple versions and variations might exist,
+these components have a very high likelihood of appearing in every container image.
+These components correspond to the core components, setup scripts, calibration data, etc.\ described previously.
+For the purposes of simulation,
+we expect these components to be very widely duplicated in jobs across all users and experiments.
+Corresponding to the full software collections maintained by each experiment,
+we assume that there will be a larger set of components that is less uniformly included in images.
+Each user or group might have a set of tools that they use heavily,
+or jobs will request particular versions of libraries and components out of many avialable in the software collection.
+
+To simulate requests similar to these usage patterns,
+we first created two pools of components/packages: core and extra components.
+We surveyed the software collections available on CVMFS and chose pools of 200 core components and 3000 extra components to approximate the size and number of software components currently available on CVMFS.
+Of these core components,
+due to the availability of multiple versions and some duplication across experiments,
+we estimate that roughly half of these core components would be required to make a functional container image.
+Thus for the simulated requests,
+we included a random selection of 100 core components.
+This represents a fairly pathological case,
+as actual depencencies would show less random behavior and variation here.
+The simulation therefore tests near-worst case behavior.
+For the extra components,
+we surveyed available software collections,
+especially those of the ATLAS and CMS experiments,
+and chose a simulated pool size of 3000 extra components.
+For simulated image requests,
+we included a random selection of up to 10\% of the extra pool.
+This is again a worst case behavior,
+as actual requests would more ordered.
+We did not allow for the entire pool to be included as there is no compelling use case for this behavior.
+Even the multi-terabyte images being tested at NERSC are taken only from a single repo,
+the CMS repo.
+There is no analysis or simulation software that uses complete software stacks from multiple repos;
+each experiment maintains its own self-contained software collection.
+Thus aside from the core components that may be used in common across experiments,
+we do not expect to see images that require more than a fraction of the total available on CVMFS.
+
+\subsection{Quantifying Image Management Costs}
+
+The next step in evaluation is to define metrics describing the collection of cached containers.
+Here we will define metrics in terms of discrete software components/packages to match the simulated streams,
+but the same definitions would apply equally well at the granularity of individual files.
+We will suppose that there is a collection of requested container images,
+from either our simulated stream or from real user requests based on job analysis and/or dynamic tracing.
+It is the responsibility of the image management system to provide a container image satisfying the requirements of each request.
+Let the set $C$ be the union of all components in all requests.
+$C$ provides a lower bound on the amount of space required to serve all requests at the same time.
+If the storage space required to hold all of $C$ is greater than the total available storage,
+then new requests will requre removal of container image(s) for previous requests.
+Since we have been considering image management as an online problem,
+we cannot place an upper bound on the sizes of requests and thus require some caching policy
+(in our case LRU) to ensure that $C$ is finite and fits available storage.
+The large, full-repo images tested at NERSC correspond to single images containing all of $C$.
+The image size is very large and any changes/updates are expensive both in terms of compute and bandwidth.
+This approach does, however,
+avoid duplication of data as there is a single image that serves all users and applications.
+We next define the multiset $A$ as the union (with repetition) of all components in all requested images.
+If multiple images request the same component,
+it will occur multiple times in $A$.
+The size required to store $A$,
+including all duplicated elements,
+is the storage space that would be required if a separate image were created for each job request.
+$A$ captures the amount of (potentially) unnecessary duplication in the collection of images.
+The initial approach described in section ?? would have a storage cost of $A$.
+Note that it is always the case that $|C| \leq |A|$,
+and if $|C| = |A|$ then there are no common components among the requested container images.
+While large, full-repo images provide the best storage efficiency, being based solely on $|C|$,
+they are expensive and unweildy to use in practice.
+The initial approach in section ?? is extremely simple and friendly to use in practice,
+but its storage cost is determined by $A$.
+Our goal, therefore, in evaluating our improved image management scheme detailed in section ??
+is to achieve storage cost as close to $C$ as possible without excessive overhead in terms of compute and network bandwidth to keep the image collection up to date.
+We will quantify this duplication ratio $d$ as $|A| / |C|$.
+For a system with no duplicated data in its container image(s),
+$d=1$.
+Larger values of $d$ indicate duplicated data.
+
+\subsection{Effect of $\alpha$ Threshold}
+
+Using these metrics for cost and overhead,
+we carried out an evaluation of our container image management scheme as follows.
+First, we somewhat arbitrarily chose a storage limit of 1~TB for our simulated image cache.
+This is not an unreasonable storage cost for an HPC site;
+for the test setup in place at NERSC,
+the full repo images already consume multiple terabytes per worker.
+We experimented with different storage limits around the same order of magnitude,
+but did not find marked differences in the behavioral trends observed below.
+The important aspect is that some definite limit exist.
+This does not take into account the storage used by Shrinkwrap,
+which is automatically deduplicated,
+nor does it take into account the metadata cost of creating projections.
+
+The next step was to generate a synthetic stream of requests for container images.
+These were generated using the procedure described previously.
+To ensure that both the merging strategy and the caching policy come into play,
+and to try to capture the steady state behavior of the system,
+we continued the stream until the total amount of unique data handled was twice the storage capacity.
+Equivalently, the request streams were long enough such that $C$ for the entire stream exceeded 2~TB.
+We then tabulated the duplication ratio $d$ for the image cache,
+as well as the total amount of data written over the course of the simulation.
+Our goal is to minimize both duplication and data written,
+which will be proportional to the compute and bandwidth cost of creating and updating container images.
+We repeated this simulation W times and computed the averages and standard deviations of these metrics.
+We repeated this procedure for a range of $\alpha$ thresholds in the range 0 to 1,
+with the results shown in figure ??.
+
+Looking at the results in figure ??,
+we can see that small values of $\alpha$ result in the highest level of data duplication,
+with $d$ decreasing as $\alpha$ increases.
+Since $\alpha=0$ corresponds to the initial approach of creating a unique container image for each job submission,
+we expect to see the highest level of duplication there.
+Looking at the cost in terms of write activity,
+we see that large values of $\alpha$ result in the most write I/O,
+and thus compute and bandwidth overhead.
+Since $\alpha=1$ allows all requests to be merged into a single large image,
+we expect minimal duplication.
+This configuration, however,
+has substantial overhead for image creation and maintenance.
+This is a real complaint from administrators at NERSC testing full-repo images.
+The high cost of creation means that such large images can only be infrequently created or updated.
+
+\section{Related Work}
+
+\section{Conclusion}
 
 \begin{figure}
-\includegraphics[width=\linewidth]{plots/IMG_20181116_101513.jpg}
-\caption{Data and metadata costs with $\alpha=A$}
+\includegraphics[width=\linewidth]{plots/file-hist/alice-ocdb.pdf}
+\caption{Distribution of file sizes in the \texttt{alice-ocdb.cern.ch} repo.
+The vertical line at $2^{20}$~B marks a file size of 1~MB.}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{plots/IMG_20181116_101907.jpg}
-\caption{Data and metadata costs with $\alpha=B$}
+\includegraphics[width=\linewidth]{plots/file-hist/alice.pdf}
+\caption{Distribution of file sizes in the \texttt{alice.cern.ch} repo.
+The vertical line at $2^{20}$~B marks a file size of 1~MB.}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{plots/IMG_20181116_102111.jpg}
-\caption{Data and metadata costs with $\alpha=C$}
+\includegraphics[width=\linewidth]{plots/file-hist/atlas-condb.pdf}
+\caption{Distribution of file sizes in the \texttt{atlas-condb.cern.ch} repo.
+The vertical line at $2^{20}$~B marks a file size of 1~MB.}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{plots/IMG_20181116_102506.jpg}
-\caption{Image size vs. requested projection size with $\alpha=A$}
+\includegraphics[width=\linewidth]{plots/file-hist/atlas.pdf}
+\caption{Distribution of file sizes in the \texttt{atlas.cern.ch} repo.
+The vertical line at $2^{20}$~B marks a file size of 1~MB.}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{plots/IMG_20181116_102623.jpg}
-\caption{Image size vs. requested projection size with $\alpha=B$}
+\includegraphics[width=\linewidth]{plots/file-hist/boss.pdf}
+\caption{Distribution of file sizes in the \texttt{boss.cern.ch} repo.
+The vertical line at $2^{20}$~B marks a file size of 1~MB.}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{plots/IMG_20181116_102706.jpg}
-\caption{Image size vs. requested projection size with $\alpha=C$}
-\end{figure}
-
-We implemented the pipeline described in section ?? on a cluster of RR machines at the Center for Research Computing~(CRC),
-a high-performance computing site at the University of Notre Dame.
-\Cref{tab:exp-setup} details the provisioned resources used in this test.
-We used the sequence of jobs from WLCG submission log discussed in section ?? as a realistic source of events for our evaluation.
-For each submission,
-we used the simplified pipeline from section ?? to build an image for running each job.
-We recorded the sizes of the image cache, projections, and data cache over the course of the workload.
-Figure ?? shows the data and metadata costs observed.
-We indeed observed rapid growth of the image cache with much slower increase in the underlying data cache.
-Figure ?? shows the amount of time spent preparing images versus the size of the requested projection.
-We see that preparation time is strongly correlated with the amount of data involved.
-
-Next, we repeated this analysis using the projection pipeline described in section ?? and $\alpha = ??$.
-We chose these values to illustrate the effects of extreme values as well as one we consider to be a good choice.
-The results with $\alpha = W$ are very similar to those of the simplified pipeline.
-This is to be expected,
-since setting $\alpha$ to be very low limits the amount of merging performed.
-Setting $\alpha = 0$ would approximate the simplified pipeline,
-in that little to no merging occurs.
-
-With $\alpha = X$,
-projections were merged very frequently.
-Figure ?? shows the steady increase in image size under this configuration.
-We see a corresponding increase in time spent constructing images in figure ??.
-With $\alpha$ set this high,
-the time spent searching for a compatible image becomes much more significant,
-hence the super-linear increase in build time.
-Approximate Jaccard distance is extremely fast to compute in memory via MinHash without interacting with the filesystem.
-As $\alpha$ is increased,
-there is less filtering from this fast approximation.
-Additional time is spent while Shrinkwrap walks the filesystem on a larger number of candidate images.
-For very homogeneous workloads,
-choice of $\alpha$ has less impact.
-In this case, however,
-different jobs are often incompatible with each other,
-so eagerly merging projections resulted in poorly chosen combinations that prevented better fits in the future.
-Thus we observe that the character of the workload is important in choosing $\alpha$.
-
-Figures ?? and ?? show the same space and time costs, respectively,
-as previously discussed for $\alpha = X$.
-Here we note a much less pronounced increase in image storage cost than figure ??.
-The increase in construction time is also less pronounced than in figure ??.
-We chose this value of $\alpha$ to be more selective than Y to avoid ballooning images but still allow more merging than Z.
-It is difficult to suggest values of $\alpha$ suitable for all cases since the machine configuration, workload, etc. all affect the patterns of behavior.
-Based on experience,
-we have found $\alpha$ values in the range V-W to be acceptable across a number of workloads and configurations.
-
-\begin{table}
-\centering
-\begin{tabular}{c|c c}
-& Cache Life (jobs) & Reuse count \\ \hline
-$\alpha=a$ & small & small? \\
-$\alpha=b$ & big & big \\
-$\alpha=c$ & small? & big \\
-\end{tabular}
-\caption{Median cache life and number of reuses for container images}
-\label{tab:cachelife}
-\end{table}
-
-We also measured the number of times a projection was reused or modified and how long images remained in the cache in an effort to estimate the reusability and relevance of images.
-\Cref{tab:cachelife} tabulates these results.
-With $\alpha = P$,
-images were not reused as often as with the other values.
-We suspect that this is because eager and frequent merges resulted in large images that were often incompatible with new jobs,
-and that drifted away such that they were no longer close enough to any projection to be used.
-The Jaccard distance metric penalizes elements in one set but not the other,
-so as more and more projections are merged,
-the resulting conglomerate moves farther and farther away from any single projection.
-Eventually, it will fall out of use and age out of the cache.
-
-In the other direction,
-the simplified pipeline and $\alpha = Q$ also had many short-lived images.
-This is likely because new images were often/always created for jobs,
-so there was little reuse.
-Due to the large number of images being created,
-images were pushed out of the cache very frequently.
-
-As part of the choice of $\alpha = Z$,
-we tried to maximize the amount of reuse and image lifetimes.
-This has the effect of indirectly minimizing storage and image construction time.
-These are also very easy metrics to track as part of batch submission.
-As such, they provide a simple way for administrators to optimize $\alpha$ to fit the system or workload.
-Some manual tuning of this form is likely necessary for applying this sort of projection pipeline in situations significantly different from our test setup.
-
-\section{Related Work}
-
-\section{Conclusion}
+\includegraphics[width=\linewidth]{plots/file-hist/cms.pdf}
+\caption{Distribution of file sizes in the \texttt{cms.cern.ch} repo.
+The vertical line at $2^{20}$~B marks a file size of 1~MB.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/file-hist/cvmfs-config.pdf}
+\caption{Distribution of file sizes in the \texttt{cvmfs-config.cern.ch} repo.
+The vertical line at $2^{20}$~B marks a file size of 1~MB.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/file-hist/geant4.pdf}
+\caption{Distribution of file sizes in the \texttt{geant4.cern.ch} repo.
+The vertical line at $2^{20}$~B marks a file size of 1~MB.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/file-hist/grid.pdf}
+\caption{Distribution of file sizes in the \texttt{grid.cern.ch} repo.
+The vertical line at $2^{20}$~B marks a file size of 1~MB.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/file-hist/lhcb.pdf}
+\caption{Distribution of file sizes in the \texttt{lhcb.cern.ch} repo.
+The vertical line at $2^{20}$~B marks a file size of 1~MB.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/file-hist/na61.pdf}
+\caption{Distribution of file sizes in the \texttt{na61.cern.ch} repo.
+The vertical line at $2^{20}$~B marks a file size of 1~MB.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/file-hist/sft.pdf}
+\caption{Distribution of file sizes in the \texttt{sft.cern.ch} repo.
+The vertical line at $2^{20}$~B marks a file size of 1~MB.}
+\end{figure}
 

### 2019-01-17 paper.tex
Updated style to ACM sigconf

@@
-\documentclass[conference]{IEEEtran}
+\documentclass[sigconf]{acmart}
 
@@
 \usepackage{hyperref}
+
+\copyrightyear{2019}
+\acmYear{2019}
+\setcopyright{acmlicensed}
+\acmConference[HPDC '19]{HPDC '19: International ACM Symposium on High-Performance Parallel and Distributed Computing}{June 24--28, 2019}{Phoenix, AZ}
 
@@
 
-\author{
-\IEEEauthorblockN{Tim Shaffer}
-\IEEEauthorblockA{University of Notre Dame \\
-Notre Dame, Indiana 46556 \\
-tshaffe1@nd.edu}
-\and
-\IEEEauthorblockN{Nicholas Hazekamp}
-\IEEEauthorblockA{University of Notre Dame \\
-Notre Dame, Indiana 46556 \\
-nhazekam@nd.edu}
-\and
-\IEEEauthorblockN{Jakob Blomer}
-\and
-\IEEEauthorblockN{Douglas Thain}
-\IEEEauthorblockA{University of Notre Dame \\
-Notre Dame, Indiana 46556 \\
-dthain@nd.edu}
+\author{Tim Shaffer}
+\email{tshaffe1@nd.edu}
+\orcid{}
+\affiliation{%
+  \institution{University of Notre Dame}
+  \city{Notre Dame}
+  \state{Indiana}
+  \postcode{46556}
 }
 
-\maketitle
+\author{Nicholas Hazekamp}
+\email{nhazekam@nd.edu}
+\orcid{}
+\affiliation{%
+  \institution{University of Notre Dame}
+  \city{Notre Dame}
+  \state{Indiana}
+  \postcode{46556}
+}
+
+\author{Jakob Blomer}
+\email{jblomer@cern.ch}
+\orcid{}
+\affiliation{%
+  \institution{CERN}
+  \city{Geneva}
+  \country{Switzerland}
+}
+
+\author{Douglas Thain}
+\email{dthain@nd.edu}
+\orcid{}
+\affiliation{%
+  \institution{University of Notre Dame}
+  \city{Notre Dame}
+  \state{Indiana}
+  \postcode{46556}
+}
+
+\renewcommand{\shortauthors}{Shaffer et al.}
 
@@
 \end{abstract}
+
+%
+% The CCSXML code is generated by the tool at http://dl.acm.org/ccs.cfm.
+% Please copy and paste the code.
+%
+
+%
+% Keywords. The author(s) should pick words that accurately describe the work being
+% presented. Separate the keywords with commas.
+\keywords{global filesystem}
+
+\maketitle
 

### 2019-01-17 micro/benchmark/benchmark_setup.sh
Added benchmarking script

@@ -1,0 +1,107 @@
+#!/bin/bash
+
+CVMFS=pwd
+
+build_cvmfs () {
+  if [ $# -eq 1" ]
+    CVMFS=$1
+  fi
+  cd $CVMFS
+  git clone https://github.com/cvmfs/cvmfs.git
+  cd cvmfs
+  mkdir build
+  cd build
+  cmake -DBUILD_SHRINKWRAP=on ../cvmfs
+  make
+}
+
+create_trace () {
+  bash $CVMFS/cvmfs/cvmfs/shrinkwrap/scripts/trace.sh --trace_dir=$1 --spec_dir=$2 run_benchmark
+}
+
+create_image () {
+  local REPO=$1
+  cvmfs_shrinkwrap -r $REPO.cern.ch -f $REPO.cern.ch.config -t $REPO.cern.ch.spec --dest-base $2 -j $3
+}
+
+create_sqfs () {
+  mksquashfs $1 $2
+}
+
+#native_cvmfs
+
+#shrinkwrap_tmp
+
+#shrinkwrap_sqfs
+
+#docker_native_cvmfs
+
+#docker_shrinkwrap_tmp
+
+#docker_shrinkwrap_mnt
+
+#docker_shrinkwrap_sqfs
+
+#parrot_cvmfs?
+
+create_id_map () {
+  echo "* $2" > $1\.map
+}
+
+create_config () {
+    local REPO=$1
+    local CONFIG=$REPO\.cern.ch.config
+    local DIR=$2
+    local KEYS=$3
+    echo "CVMFS_REPOSITORIES=$REPO.cern.ch" > $CONFIG
+    echo "CVMFS_REPOSITORY_NAME=$REPO.cern.ch" >> $CONFIG
+    echo "CVMFS_CONFIG_REPOSITORY=cvmfs-config.cern.ch" >> $CONFIG
+    echo "CVMFS_DEFAULT_DOMAIN=cern.ch" >> $CONFIG
+    echo "CVMFS_SERVER_URL='http://cvmfs-stratum-zero-hpc.cern.ch/cvmfs/$REPO.cern.ch;http://cvmfs-stratum-one.cern.ch/cvmfs/$REPO.cern.ch'" >> $CONFIG
+    echo "CVMFS_HTTP_PROXY=DIRECT" >> $CONFIG
+    echo "CVMFS_MOUNT_DIR=/cvmfs" >> $CONFIG
[…]

### 2019-01-18 simulation/dat.py
Dump in simulator

@@ -1,0 +1,40 @@
+#!/usr/bin/env python
+
+import sys
+import json
+
+import sim
+
+INST = ['overhead', 'efficiency', 'images', 'imagesizes', 'pkgs',
+        'realsize', 'reqsize', 'requests', 'size', 'upkgs', 'usize',
+        'txcost', 'pool', 'activeimgs', 'workers', 'tx']
+CUM = ['byteswritten', 'deletes', 'inserts', 'merges', 'hits']
+
+def sample_final(samples, field):
+    return samples[-1][field]
+
+def sample_median(samples, field):
+    values = [s[field] for s in samples]
+    values.sort()
+    return sim.median(values)
+
+raw = json.load(sys.stdin)[sys.argv[1]]
+dat = dict()
+
+for alpha, runs in raw.items():
+    dat[float(alpha)] = runs
+    for stream in runs:
+        for sample in stream:
+            sample['overhead'] = sample['realsize'] - sample['reqsize']
+            sample['efficiency'] = float(sample['reqsize'])/sample['realsize']
+            sample['txcost'] = sample['tx'] * sample['realsize']
+
+print(' '.join(['alpha'] + INST + CUM))
+
+alphas = dat.keys()
+alphas.sort()
+
+for i in alphas:
+    cc = [str(sim.smedian([sample_final(stream, f) for stream in dat[i]])) for f in CUM]
+    ci = [str(sim.smedian([sample_median(stream, f) for stream in dat[i]])) for f in INST]
+    print(' '.join([str(i)] + ci + cc))

### 2019-01-18 simulation/deps.py

@@ -1,0 +1,29 @@
+#!/usr/bin/env python
+
+import fileinput
+import json
+import sys
+import re
+
+with open(sys.argv[1]) as f:
+    builds = json.load(f)
+
+sizes = dict()
+with open(sys.argv[2]) as f:
+    for line in f.readlines():
+        (s, p) = line.split('\t')
+        sizes[p.strip()] = int(s)
+
+deps = dict()
+for p, b in builds.items():
+    pkg = '{}-{}-{}'.format(b["NAME"], b["VERSION"], b["HASH"])
+    if not pkg in deps:
+        deps[pkg] = {"size": list(), "deps": set()}
+    deps[pkg]["size"].append(sizes[re.sub(r'[^/]*$', '', p)])
+    if "DEPENDS" in b:
+        deps[pkg]["deps"].update(b["DEPENDS"].split(","))
+
+for e in deps.values():
+    e["deps"] = [x for x in e["deps"] if x in deps]
+
+json.dump(deps, sys.stdout, indent=2)

### 2019-01-18 simulation/graph.py

@@ -1,0 +1,31 @@
+#!/usr/bin/env python
+
+import sys
+import json
+
+def closure(graph, deps, pkg):
+    if pkg in graph: return
+    graph[pkg] = deps[pkg]
+    for p in deps[pkg]['deps']:
+        closure(graph, deps, p)
+
+with open(sys.argv[1]) as f:
+    deps = json.load(f)
+with open(sys.argv[2]) as f:
+    recent = json.load(f)
+
+graph = dict()
+
+for p, (v, h) in recent.items():
+    pkg = '{}-{}-{}'.format(p, v, h)
+    closure(graph, deps, pkg)
+
+print('digraph {')
+
+for k, v in graph.items():
+    #print('"{}";'.format(k))
+
+    for d in v['deps']:
+        print('"{}" -> "{}";'.format(k, d))
+
+print('}')

### 2019-01-18 simulation/picks.py

@@ -1,0 +1,28 @@
+#!/usr/bin/env python
+
+import sys
+import json
+import random
+
+import sim
+
+SAMPLES = 100
+PICKS = 1000
+
+sizes = [list() for i in range(PICKS)]
+counts = [list() for i in range(PICKS)]
+
+deps = json.load(sys.stdin)
+
+for i in range(SAMPLES):
+    out = set() 
+    sys.stderr.write('{} '.format(i))
+    for j in range(PICKS):
+        sim.closure(deps, random.choice(deps.keys()), out)
+        counts[j].append(len(out))
+        sizes[j].append(sim.size(deps, out))
+
+for i in range(PICKS):
+    sizes[i].sort()
+    counts[i].sort()
+    print('{} {} {}'.format(i, sim.median(counts[i]), sim.median(sizes[i])))

### 2019-01-18 simulation/read_buildinfo.py

@@ -1,0 +1,37 @@
+#!/usr/bin/env python
+
+import re
+import json
+import sys
+import fileinput
+
+def label(s):
+    m = re.search(r'\b[A-Z]+$', s)
+    if not m: return
+    return m.group(0)
+
+def body(s):
+    m = re.match(r'(.*)(?=\b[A-Z]+$)', s)
+    if m: 
+        out = m.group(1)
+    else:
+        out = s
+    return out.strip().strip(',')
+
+def parse(s):
+    info = dict()
+    chunks = s.split(':')
+    for i in range(len(chunks) - 1):
+        l = label(chunks[i])
+        if not l: continue
+        info[l] = body(chunks[i + 1])
+    return info
+
+builds = dict()
+
+for line in fileinput.input():
+    line = line.strip()
+    with open(line) as f:
+        builds[line] = parse(f.read())
+
+json.dump(builds, sys.stdout, indent=2)

### 2019-01-18 simulation/recent.py

@@ -1,0 +1,26 @@
+#!/usr/bin/env python
+
+from distutils.version import LooseVersion, StrictVersion
+import json
+import sys
+
+builds = json.load(sys.stdin)
+
+packages = dict()
+recent = dict()
+
+for b in builds.values():
+    if ',' in b['VERSION']: continue
+    if not b['NAME'] in packages:
+        packages[b['NAME']] = dict()
+    packages[b['NAME']][b['VERSION']] = b['HASH']
+
+for pkg,vers in packages.items():
+    for v, h in vers.items():
+        if pkg in recent:
+            if LooseVersion(v) > LooseVersion(recent[pkg][0]):
+                recent[pkg] = (v, h)
+        else:
+            recent[pkg] = (v, h)
+
+json.dump(recent, sys.stdout, indent=2)

### 2019-01-18 simulation/ser.py

@@ -1,0 +1,26 @@
+#!/usr/bin/env python
+
+import sys
+import json
+
+import sim
+
+
+def extract(samples, field, step):
+    values = [s[step][field] for s in samples]
+    values.sort()
+    return sim.median(values)
+
+dat = json.load(sys.stdin)[sys.argv[1]][sys.argv[2]]
+
+for run in dat:
+    for sample in run:
+        sample['efficiency'] = float(sample['reqsize']) / sample['realsize']
+
+fields = dat[0][0].keys()
+fields.sort()
+print(' '.join(['step'] + fields))
+
+for i in range(len(dat[0])):
+    cols = [str(extract(dat, f, i)) for f in fields]
+    print(' '.join([str(i)] + cols))

### 2019-01-18 simulation/sim.py

@@ -1,0 +1,301 @@
+#!/usr/bin/env python
+
+import sys
+import json
+import random
+import numbers
+import itertools
+import collections
+import multiprocessing
+
+CAPACITY = 1e12
+WORKERS = 1000
+MAXREQ = 100
+REUSE = 9
+JOBS = 100
+
+def median(lst):
+    # lst must already be sorted!
+    if len(lst) == 0: return 0
+    if len(lst) == 1: return lst[0]
+    i = len(lst) // 2
+    if len(lst) % 2 == 0:
+        return (lst[i] + lst[i - 1]) / 2
+    else:
+        return lst[i]
+
+def smedian(lst):
+    lst.sort()
+    return median(lst)
+
+def jaccard(a, b):
+    return 1 - float(len(a & b))/len(a | b)
+
+def real_img(v):
+    return isinstance(v, numbers.Integral)
+
+def size(deps, img):
+    accum = 0
+    for pkg in img:
+        accum += median(deps[pkg]['size'])
+    return accum
+
+def closure(deps, pkg, res):
+    if pkg in res: return
+    res.add(pkg)
+    for p in deps[pkg]['deps']:
+        closure(deps, p, res)
+
+def blind(deps, img):
+    return set(random.sample(deps.keys(), len(img)))
+
+class Stream:
+    def __init__(self, deps):
+        self.deps = deps
+    def __iter__(self):
+        return self
+    def next(self):
+        out = set()
+        count = random.randrange(1, MAXREQ)
+        for pkg in random.sample(self.deps.keys(), count):
+           closure(deps, pkg, out) 
+        return frozenset(out)
+
+class BlindStream:
+    def __init__(self, deps):
+        self.deps = deps
+        self.stream = Stream(deps)
+    def __iter__(self):
+        return self
+    def next(self):
+        return frozenset(blind(self.deps, self.stream.next()))
+
+class Cache:
+    def __init__(self, deps, alpha):
+        self.deps = deps
+        self.size = 0
+        self.alpha = alpha
+        self.bytes_written = 0
+        self.merges = 0
+        self.inserts = 0
+        self.deletes = 0
+        self.hits = 0
+        self.contents = collections.OrderedDict()
+        self.log = []
+        self.workers = {(): WORKERS}
+        self.pool = {}
+
+    def unique(self):
+        return {item for subset in self.contents.keys() for item in subset}
+
+    def total(self):
[…]

### 2019-01-19 micro/benchmark/benchmark_setup.sh
Added spec and make for micro benchmarking

@@ -1,15 +1,18 @@
 #!/bin/bash
 
-CVMFS=pwd
+CVMFS=$(pwd)
 
 build_cvmfs () {
-  if [ $# -eq 1" ]
-    CVMFS=$1
+  if [ $# -eq 1 ]
+    then
+      CVMFS=$1
   fi
   cd $CVMFS
-  git clone https://github.com/cvmfs/cvmfs.git
-  cd cvmfs
-  mkdir build
+  if [ ! -d "cvmfs" ]
+    then
+      git clone https://github.com/cvmfs/cvmfs.git
+  fi
+  mkdir -p build
   cd build
   cmake -DBUILD_SHRINKWRAP=on ../cvmfs
   make
@@ -21,7 +24,7 @@ create_trace () {
 
 create_image () {
   local REPO=$1
-  cvmfs_shrinkwrap -r $REPO.cern.ch -f $REPO.cern.ch.config -t $REPO.cern.ch.spec --dest-base $2 -j $3
+  $CVMFS/build/cvmfs/cvmfs_shrinkwrap -r $REPO.cern.ch -f $REPO.cern.ch.config -t $REPO.cern.ch.spec --dest-base $2 -j $3
 }
 
 create_sqfs () {
@@ -61,7 +64,7 @@ create_config () {
     echo "CVMFS_HTTP_PROXY=DIRECT" >> $CONFIG
     echo "CVMFS_MOUNT_DIR=/cvmfs" >> $CONFIG
     echo "CVMFS_CACHE_BASE=$DIR/cache/cvmfs/shrinkwrap" >> $CONFIG
-    echo "CVMFS_SHARED_CACHE=no # Important as libcvmfs does not support shared caches" >> $CONFIG
+    echo "CVMFS_SHARED_CACHE=no" >> $CONFIG
     echo "CVMFS_USER=$USER" >> $CONFIG
     echo "CVMFS_DEBUGLOG=$DIR/cvmfs.log" >> $CONFIG
     echo "CVMFS_SYSLOG_LEVEL=LOG_DEBUG" >> $CONFIG
@@ -92,8 +95,8 @@ build_benchmark () {
 }
 
 run_benchmark () {
-    time cern-benchmark --benchmarks="kv;DB12;whetstone" -o
-    time cern-benchmark --benchmarks="kv;DB12;whetstone" -o --mp_num=1
+    time cern-benchmark --benchmarks="kv;DB12;whetstone" -o;
[…]

### 2019-01-20 simulation/picks.py
Add usage distribution

@@ -26,3 +26,5 @@ for i in range(PICKS):
     sizes[i].sort()
     counts[i].sort()
     print('{} {} {}'.format(i, sim.median(counts[i]), sim.median(sizes[i])))
+
+sys.stderr.write('\n')

### 2019-01-20 simulation/read_usage.py

@@ -1,0 +1,47 @@
+#!/usr/bin/env python
+
+import re
+import json
+import sys
+import fileinput
+
+def package(s):
+    path_pieces = s.split('/')
+    if len(path_pieces) >= 4: 
+        out = path_pieces[3]
+        if len(path_pieces) >= 5:
+            out += '-'
+            out += path_pieces[4]
+    else:
+        out = s
+    return out
+
+def parse(f):
+    info = dict()
+    s = f.readline()
+    while s:
+        chunks = s.split(":")
+	if len(chunks) < 2:
+            s = f.readline()
+            continue
+        pack = chunks[1].split(' ')
+	if len(pack) < 2:
+            s = f.readline()
+            continue
+        p = package(pack[0])
+        if p:
+            info[p] = info.get(p, 0) + 1
+            packages[p] = packages.get(p, 0) + 1
+        s = f.readline()
+    return info
+
+builds = dict()
+packages = dict()
+
+for line in fileinput.input():
+    line = line.strip()
+    with open(line) as f:
+        builds[line] = parse(f)
+
+json.dump(packages, sys.stdout, indent=2)
+

### 2019-01-20 simulation/sim.py

@@ -8,10 +8,12 @@ import itertools
 import collections
 import multiprocessing
 
-CAPACITY = 1e12
+from numpy.random import choice
+
+CAPACITY = 2.5e11
 WORKERS = 1000
 MAXREQ = 100
-REUSE = 9
+REUSE = 5
 JOBS = 100
 
 def median(lst):
@@ -50,21 +52,22 @@ def blind(deps, img):
     return set(random.sample(deps.keys(), len(img)))
 
 class Stream:
-    def __init__(self, deps):
+    def __init__(self, deps, deps_freq):
         self.deps = deps
+        self.dep_freq = deps_freq
     def __iter__(self):
         return self
     def next(self):
         out = set()
         count = random.randrange(1, MAXREQ)
-        for pkg in random.sample(self.deps.keys(), count):
+        for pkg in choice(self.deps.keys(), count, False, self.dep_freq):
            closure(deps, pkg, out) 
         return frozenset(out)
 
 class BlindStream:
-    def __init__(self, deps):
+    def __init__(self, deps, deps_freq):
         self.deps = deps
-        self.stream = Stream(deps)
+        self.stream = Stream(deps, deps_freq)
     def __iter__(self):
         return self
     def next(self):
@@ -258,13 +261,13 @@ class Cache:
         return self.log
 
 def run(params):
-    alpha, deps = params
+    alpha, deps, deps_freq = params
     out = {}
     for i in range(3):
         c = Cache(deps, alpha)
         d = Cache(deps, alpha)
-        c.process(itertools.islice(Stream(deps), JOBS))
-        d.process(itertools.islice(BlindStream(deps), JOBS))
+        c.process(itertools.islice(Stream(deps, deps_freq), JOBS))
+        d.process(itertools.islice(BlindStream(deps, deps_freq), JOBS))
         out['tree'] = c.log
         out['blind'] = d.log
     sys.stderr.write('{} '.format(alpha))
@@ -276,19 +279,29 @@ if __name__ == '__main__':
     for d in deps.values():
         d['size'].sort()
 
+    deps_freq = []
+    total = 0.0
+    usage = open("sft.usage.json", 'r')
+    freq = json.load(usage)
+    for d in deps.keys():
+	total += freq.get(d, 1.0)
+        deps_freq.append(freq.get(d, 1))
+
+    deps_freq[:] = [x / total for x in deps_freq]
+
     out = {
         "tree": {},
[…]

### 2019-01-20 simulation/dat.py
Switch to Jaccard-heavy algorithm

@@ -6,7 +6,7 @@ import json
 import sim
 
 INST = ['overhead', 'efficiency', 'images', 'imagesizes', 'pkgs',
-        'realsize', 'reqsize', 'requests', 'size', 'upkgs', 'usize',
+        'realsize', 'reqsize', 'size', 'upkgs', 'usize',
         'txcost', 'pool', 'activeimgs', 'workers', 'tx']
 CUM = ['byteswritten', 'deletes', 'inserts', 'merges', 'hits']
 

### 2019-01-20 simulation/sim.py

@@ -33,9 +33,6 @@ def smedian(lst):
 def jaccard(a, b):
     return 1 - float(len(a & b))/len(a | b)
 
-def real_img(v):
-    return isinstance(v, numbers.Integral)
-
 def size(deps, img):
     accum = 0
     for pkg in img:
@@ -87,12 +84,21 @@ class Cache:
         self.log = []
         self.workers = {(): WORKERS}
         self.pool = {}
+        self.distances = {}
+
+    def lazy_jaccard(self, a, b):
+        if (a, b) in self.distances:
+            return self.distances[(a, b)]
+        out = jaccard(a, b)
+        self.distances[(a, b)] = out
+        self.distances[(b, a)] = out
+        return out
 
     def unique(self):
         return {item for subset in self.contents.keys() for item in subset}
 
     def total(self):
-        return [item for (k, v) in self.contents.items() for item in k if real_img(v)]
+        return [item for subset in self.contents.keys() for item in subset]
 
     def pkgs(self):
         return len(self.total())
@@ -101,18 +107,13 @@ class Cache:
         return len(self.unique())
 
     def unique_size(self):
-        return sum([size(self.deps, [x]) for x in self.unique()])
-
-    def requests(self):
-        return self.contents.keys()
+        return size(self.deps, self.unique())
 
     def images(self):
-        return [k for k, v in self.contents.items() if real_img(v)]
+        return self.contents.keys()
 
     def image_sizes(self):
-        out = [v for k, v in self.contents.items() if real_img(v)]
-        out.sort()
-        return out
+        return self.contents.values()
 
     def stats(self):
         return {
@@ -124,9 +125,8 @@ class Cache:
             "inserts": self.inserts,
             "deletes": self.deletes,
             "usize": self.unique_size(),
-            "requests": len(self.requests()),
             "images": len(self.images()),
-            "imagesizes": median(self.image_sizes()),
+            "imagesizes": smedian(self.image_sizes()),
             "activeimgs": len(self.workers),
             "pool": len(self.pool),
             "hits": self.hits,
@@ -146,22 +146,9 @@ class Cache:
         self.workers[img] = avail + transfers
         return transfers
 
-    def tidy(self, img=None):
-        if img is None:
-            for img in self.contents.keys():
-                self.tidy(img)
-            return
-        if not img in self.contents: return False
-        if real_img(self.contents[img]): return True
[…]

### 2019-01-21 simulation/dat.py
Fixups

@@ -7,8 +7,9 @@ import sim
 
 INST = ['overhead', 'efficiency', 'images', 'imagesizes', 'pkgs',
         'realsize', 'reqsize', 'size', 'upkgs', 'usize',
-        'txcost', 'pool', 'activeimgs', 'workers', 'tx']
-CUM = ['byteswritten', 'deletes', 'inserts', 'merges', 'hits']
+        'pool', 'activeimgs', 'workers']
+CUM = ['byteswritten', 'deletes', 'inserts', 'merges', 'hits',
+       'tx', 'txcost', 'peaksize', 'cost']
 
 def sample_final(samples, field):
     return samples[-1][field]
@@ -18,23 +19,30 @@ def sample_median(samples, field):
     values.sort()
     return sim.median(values)
 
+def cost2(samples):
+    return sample_median(samples,'size') / sample_median(samples, 'usize') * sample_final(samples,'peaksize') + sample_final(samples, 'txcost')/1000
+
 raw = json.load(sys.stdin)[sys.argv[1]]
 dat = dict()
 
 for alpha, runs in raw.items():
     dat[float(alpha)] = runs
     for stream in runs:
+        p = 0
         for sample in stream:
             sample['overhead'] = sample['realsize'] - sample['reqsize']
             sample['efficiency'] = float(sample['reqsize'])/sample['realsize']
-            sample['txcost'] = sample['tx'] * sample['realsize']
+            p = max(sample['size'], p)
+            sample['peaksize'] = p
+            sample['cost'] = sample['peaksize'] + sample['txcost']/1000
 
-print(' '.join(['alpha'] + INST + CUM))
+print(' '.join(['alpha', 'cost2'] + INST + CUM))
 
 alphas = dat.keys()
 alphas.sort()
 
 for i in alphas:
+    other = str(sim.smedian([cost2(stream) for stream in dat[i]]))
     cc = [str(sim.smedian([sample_final(stream, f) for stream in dat[i]])) for f in CUM]
     ci = [str(sim.smedian([sample_median(stream, f) for stream in dat[i]])) for f in INST]
-    print(' '.join([str(i)] + ci + cc))
+    print(' '.join([str(i), other] + ci + cc))

### 2019-01-21 simulation/sim.py

@@ -10,10 +10,10 @@ import multiprocessing
 
 from numpy.random import choice
 
-CAPACITY = 2.5e11
+CAPACITY = 1e12
 WORKERS = 1000
 MAXREQ = 100
-REUSE = 5
+REUSE = 9
 JOBS = 100
 
 def median(lst):
@@ -80,9 +80,12 @@ class Cache:
         self.inserts = 0
         self.deletes = 0
         self.hits = 0
+        self.tx = 0
+        self.tx_cost = 0
         self.contents = collections.OrderedDict()
         self.log = []
-        self.workers = {(): WORKERS}
+        self.workers = collections.OrderedDict()
+        self.workers[()] = WORKERS
         self.pool = {}
         self.distances = {}
 
@@ -130,6 +133,8 @@ class Cache:
             "activeimgs": len(self.workers),
             "pool": len(self.pool),
             "hits": self.hits,
+            "tx": self.tx,
+            "txcost": self.tx_cost,
         }
 
     def push_workers(self, img, count):
@@ -137,10 +142,11 @@ class Cache:
         avail = self.workers.pop(img, 0)
         count -= avail
         while count > 0:
-            other = random.choice(self.workers.keys())
-            self.workers[other] -= 1
-            count -= 1
-            transfers += 1
+            other = self.workers.keys()[0]
+            tx = min(count, self.workers[other])
+            self.workers[other] -= tx
+            count -= tx
+            transfers += tx
             if self.workers[other] <= 0:
                 self.workers.pop(other)
         self.workers[img] = avail + transfers
@@ -179,6 +185,16 @@ class Cache:
         if img in self.contents:
             return self.hit(img, img)
 
+        # for worker simulation
+        # prioritize stuff already on workers
+        pushed = [(self.lazy_jaccard(img, x), x) for x in self.workers.keys() if x in self.contents]
+        pushed = [x for x in pushed if x[0] < self.alpha]
+        pushed.sort(key=lambda x: x[0])
+
+        for a in pushed:
+            if img.issubset(a[1]):
+                return self.hit(a[1], img)
+
         candidates = [(self.lazy_jaccard(img, x), x) for x in self.contents.keys()]
         candidates = [x for x in candidates if x[0] < self.alpha]
         candidates.sort(key=lambda x: x[0])
@@ -197,8 +213,10 @@ class Cache:
         self.log.append(self.stats())
[…]

### 2019-01-21 simulation/sim.py
Added dist stream

@@ -48,7 +48,7 @@ def closure(deps, pkg, res):
 def blind(deps, img):
     return set(random.sample(deps.keys(), len(img)))
 
-class Stream:
+class DistStream:
     def __init__(self, deps, deps_freq):
         self.deps = deps
         self.dep_freq = deps_freq
@@ -61,10 +61,23 @@ class Stream:
            closure(deps, pkg, out) 
         return frozenset(out)
 
-class BlindStream:
-    def __init__(self, deps, deps_freq):
+
+class Stream:
+    def __init__(self, deps):
         self.deps = deps
-        self.stream = Stream(deps, deps_freq)
+    def __iter__(self):
+        return self
+    def next(self):
+        out = set()
+        count = random.randrange(1, MAXREQ)
+        for pkg in random.sample(self.deps.keys(), count):
+           closure(deps, pkg, out) 
+        return frozenset(out)
+
+class BlindStream:
+    def __init__(self, deps):
+        self.deps = deps
+        self.stream = Stream(deps)
     def __iter__(self):
         return self
     def next(self):
@@ -248,10 +261,13 @@ def run(params):
     alpha, deps, deps_freq = params
     out = {}
     for i in range(10):
+        b = Cache(deps, alpha)
         c = Cache(deps, alpha)
         d = Cache(deps, alpha)
-        c.process(itertools.islice(Stream(deps, deps_freq), JOBS))
-        d.process(itertools.islice(BlindStream(deps, deps_freq), JOBS))
+        b.process(itertools.islice(DistStream(deps, deps_freq), JOBS))
+        c.process(itertools.islice(Stream(deps), JOBS))
+        d.process(itertools.islice(BlindStream(deps), JOBS))
+        out['dist'] = b.log
         out['tree'] = c.log
         out['blind'] = d.log
     sys.stderr.write('{} '.format(alpha))

### 2019-01-21 simulation/sim.py
Remove workers, speed things up

@@ -11,7 +11,6 @@ import multiprocessing
 from numpy.random import choice
 
 CAPACITY = 1e12
-WORKERS = 1000
 MAXREQ = 100
 REUSE = 9
 JOBS = 100
@@ -93,21 +92,24 @@ class Cache:
         self.inserts = 0
         self.deletes = 0
         self.hits = 0
-        self.tx = 0
-        self.tx_cost = 0
         self.contents = collections.OrderedDict()
         self.log = []
-        self.workers = collections.OrderedDict()
-        self.workers[()] = WORKERS
-        self.pool = {}
         self.distances = {}
+        self.subsets = {}
 
-    def lazy_jaccard(self, a, b):
+    def jaccard(self, a, b):
         if (a, b) in self.distances:
             return self.distances[(a, b)]
         out = jaccard(a, b)
         self.distances[(a, b)] = out
         self.distances[(b, a)] = out
+        return out
+
+    def issubset(self, a, b):
+        if (a, b) in self.subsets:
+            return self.subsets[(a, b)]
+        out = a.issubset(b)
+        self.subsets[(a, b)] = out
         return out
 
     def unique(self):
@@ -143,27 +145,8 @@ class Cache:
             "usize": self.unique_size(),
             "images": len(self.images()),
             "imagesizes": smedian(self.image_sizes()),
-            "activeimgs": len(self.workers),
-            "pool": len(self.pool),
             "hits": self.hits,
-            "tx": self.tx,
-            "txcost": self.tx_cost,
         }
-
-    def push_workers(self, img, count):
-        transfers = 0
-        avail = self.workers.pop(img, 0)
-        count -= avail
-        while count > 0:
-            other = self.workers.keys()[0]
-            tx = min(count, self.workers[other])
-            self.workers[other] -= tx
-            count -= tx
-            transfers += tx
-            if self.workers[other] <= 0:
-                self.workers.pop(other)
-        self.workers[img] = avail + transfers
-        return transfers
 
     def merge(self, existing, img):
         self.size -= self.contents.pop(existing)
@@ -198,22 +181,12 @@ class Cache:
         if img in self.contents:
             return self.hit(img, img)
 
-        # for worker simulation
-        # prioritize stuff already on workers
-        pushed = [(self.lazy_jaccard(img, x), x) for x in self.workers.keys() if x in self.contents]
-        pushed = [x for x in pushed if x[0] < self.alpha]
-        pushed.sort(key=lambda x: x[0])
[…]

### 2019-01-21 simulation/sim.py
Updated plots

@@ -7,13 +7,12 @@ import numbers
 import itertools
 import collections
 import multiprocessing
+import argparse
 
 from numpy.random import choice
 
-CAPACITY = 1e12
+CAPACITY = 1.5e12
 MAXREQ = 100
-REUSE = 9
-JOBS = 100
 
 def median(lst):
     # lst must already be sorted!
@@ -83,7 +82,8 @@ class BlindStream:
         return frozenset(blind(self.deps, self.stream.next()))
 
 class Cache:
-    def __init__(self, deps, alpha):
+    def __init__(self, deps, alpha, reuse):
+        self.reuse = reuse
         self.deps = deps
         self.size = 0
         self.alpha = alpha
@@ -211,7 +211,7 @@ class Cache:
     def process(self, stream):
         pool = []
         for img in stream:
-            for i in range(REUSE):
+            for i in range(self.reuse):
                 pool.append(img)
         random.shuffle(pool)
         for img in pool:
@@ -220,15 +220,15 @@ class Cache:
         return self.log
 
 def run(params):
-    alpha, deps, deps_freq = params
+    alpha, deps, deps_freq, reuse, jobs = params
     out = {}
     for i in range(10):
-        b = Cache(deps, alpha)
-        c = Cache(deps, alpha)
-        d = Cache(deps, alpha)
-        b.process(itertools.islice(DistStream(deps, deps_freq), JOBS))
-        c.process(itertools.islice(Stream(deps), JOBS))
-        d.process(itertools.islice(BlindStream(deps), JOBS))
+        b = Cache(deps, alpha, reuse)
+        c = Cache(deps, alpha, reuse)
+        d = Cache(deps, alpha, reuse)
+        b.process(itertools.islice(DistStream(deps, deps_freq), jobs))
+        c.process(itertools.islice(Stream(deps), jobs))
+        d.process(itertools.islice(BlindStream(deps), jobs))
         out['dist'] = b.log
         out['tree'] = c.log
         out['blind'] = d.log
@@ -237,6 +237,11 @@ def run(params):
     
 
 if __name__ == '__main__':
+    parser = argparse.ArgumentParser()
+    parser.add_argument('--reuse', type=int)
+    parser.add_argument('--jobs', type=int)
+    args = parser.parse_args()
+
     deps = json.load(sys.stdin)
     for d in deps.values():
         d['size'].sort()
[…]

### 2019-01-21 simulation/dat.py
Add option for cache size

@@ -6,10 +6,9 @@ import json
 import sim
 
 INST = ['overhead', 'efficiency', 'images', 'imagesizes', 'pkgs',
-        'realsize', 'reqsize', 'size', 'upkgs', 'usize',
-        'pool', 'activeimgs', 'workers']
+        'realsize', 'reqsize', 'size', 'upkgs', 'usize']
 CUM = ['byteswritten', 'deletes', 'inserts', 'merges', 'hits',
-       'tx', 'txcost', 'peaksize', 'cost']
+       'peaksize']
 
 def sample_final(samples, field):
     return samples[-1][field]
@@ -19,8 +18,6 @@ def sample_median(samples, field):
     values.sort()
     return sim.median(values)
 
-def cost2(samples):
-    return sample_median(samples,'size') / sample_median(samples, 'usize') * sample_final(samples,'peaksize') + sample_final(samples, 'txcost')/1000
 
 raw = json.load(sys.stdin)[sys.argv[1]]
 dat = dict()
@@ -34,15 +31,13 @@ for alpha, runs in raw.items():
             sample['efficiency'] = float(sample['reqsize'])/sample['realsize']
             p = max(sample['size'], p)
             sample['peaksize'] = p
-            sample['cost'] = sample['peaksize'] + sample['txcost']/1000
 
-print(' '.join(['alpha', 'cost2'] + INST + CUM))
+print(' '.join(['alpha'] + INST + CUM))
 
 alphas = dat.keys()
 alphas.sort()
 
 for i in alphas:
-    other = str(sim.smedian([cost2(stream) for stream in dat[i]]))
     cc = [str(sim.smedian([sample_final(stream, f) for stream in dat[i]])) for f in CUM]
     ci = [str(sim.smedian([sample_median(stream, f) for stream in dat[i]])) for f in INST]
-    print(' '.join([str(i), other] + ci + cc))
+    print(' '.join([str(i)] + ci + cc))

### 2019-01-21 simulation/sim.py

@@ -11,7 +11,6 @@ import argparse
 
 from numpy.random import choice
 
-CAPACITY = 1.5e12
 MAXREQ = 100
 
 def median(lst):
@@ -82,8 +81,9 @@ class BlindStream:
         return frozenset(blind(self.deps, self.stream.next()))
 
 class Cache:
-    def __init__(self, deps, alpha, reuse):
+    def __init__(self, deps, alpha, reuse, capacity):
         self.reuse = reuse
+        self.capacity = capacity
         self.deps = deps
         self.size = 0
         self.alpha = alpha
@@ -203,7 +203,7 @@ class Cache:
     def shrink(self):
         dead_img = None
         dead_size = None
-        while self.size > CAPACITY:
+        while self.size > self.capacity:
             dead_img, dead_size = self.contents.popitem(False)
             self.deletes += 1
             self.size -= dead_size
@@ -220,18 +220,18 @@ class Cache:
         return self.log
 
 def run(params):
-    alpha, deps, deps_freq, reuse, jobs = params
+    alpha, deps, deps_freq, reuse, jobs, capacity = params
     out = {}
     for i in range(10):
-        b = Cache(deps, alpha, reuse)
-        c = Cache(deps, alpha, reuse)
-        d = Cache(deps, alpha, reuse)
+        b = Cache(deps, alpha, reuse, capacity)
+        #c = Cache(deps, alpha, reuse, capacity)
+        #d = Cache(deps, alpha, reuse, capacity)
         b.process(itertools.islice(DistStream(deps, deps_freq), jobs))
-        c.process(itertools.islice(Stream(deps), jobs))
-        d.process(itertools.islice(BlindStream(deps), jobs))
+        #c.process(itertools.islice(Stream(deps), jobs))
+        #d.process(itertools.islice(BlindStream(deps), jobs))
         out['dist'] = b.log
-        out['tree'] = c.log
-        out['blind'] = d.log
+        #out['tree'] = c.log
+        #out['blind'] = d.log
     sys.stderr.write('{} '.format(alpha))
     return out
     
@@ -240,6 +240,7 @@ if __name__ == '__main__':
     parser = argparse.ArgumentParser()
     parser.add_argument('--reuse', type=int)
     parser.add_argument('--jobs', type=int)
+    parser.add_argument('--cache', type=float)
     args = parser.parse_args()
 
     deps = json.load(sys.stdin)
[…]

### 2019-01-21 simulation/sim.py
Fix output

@@ -277,8 +277,8 @@ if __name__ == '__main__':
     sys.stderr.flush()
 
     for i in range(len(alphas)):
-        out['tree'][alphas[i]].append(results[i]['tree'])
-        out['blind'][alphas[i]].append(results[i]['blind'])
+        #out['tree'][alphas[i]].append(results[i]['tree'])
+        #out['blind'][alphas[i]].append(results[i]['blind'])
         out['dist'][alphas[i]].append(results[i]['dist'])
 
     json.dump(out, sys.stdout, indent=2)

### 2019-01-22 paper.tex
Added acm ref file to allow off overleaf builds. NOT TESTED
Updates from ShareLaTeX

@@
 \begin{document}
-\title{Projecting Global Filesystems into Local Containers for High-throughput Computing on HPC Resources}
+\title[Projecting Globel Filesystems into Local Containers.]{Projecting Global Filesystems into Local Containers for High-throughput Computing on HPC Resources}
 
@@
 This approach, while reliable, tends to be labor-intensive and tied to specific experiments and software stacks.
-We therefore modified the reference implementation of CVMFS to allow us to capture the filesystem interactions of a live HEP applications.
+We therefore modified the reference implementation of CVMFS to allow us to capture the filesystem interactions of live HEP applications.
 Using live tracing and static analysis,
-we are able to determine exactly which parts of CVMFS must be included in a container image for a particular job.
+we are able to determine exactly which parts of the \texttt{/cvmfs} file system tree CVMFS must be included in a container image for a particular job.
 We developed a tool, Shrinkwrap, that can efficiently \emph{project} CVMFS (or a subset thereof) at a particular version into a traditional filesystem.
@@
 Leadership class machines such as Summit (?) will by themselves be able to perform a substantial portion of the work of the entire WLCG today.
-Use of national-scale HPC resources across the United States and Europe such as XXX, YYY, ZZZ could be key in meeting the computational demands following the LHC's high-luminosity upgrade.
+Use of national-scale HPC resources across the United States and Europe such as XXX, YYY, Piz Daint at the Swiss CSCS HPC center could be key in meeting the computational demands following the LHC's high-luminosity upgrade.
 
@@
 
-Regardless of how the trace was produced,
-there are several ways a trace needs to communicate 
-which files and how they were used.
-Below is a table describing different axes by which a 
-specification can be tracked:
-\begin{table}[h]
-    \centering
-    \begin{tabular}{|c|c|}
-    \hline
-    Number of files & The number of files in a given subtree. \\ \hline
-    Depth of Tree & The largest number of nested directories. \\ \hline
-    Width of Tree & The number of directory splits in a subtree.\\ \hline
-    Size of files & Sum of the files contents in subtree.\\ \hline
-    Actual files & The actual file names.\\ \hline
-    Revision/Release Number & The version of subtree/compatibility. \\ \hline
-    Checksum & The files computed checksum. \\ \hline
-    File type & Directories vs Files.\\ \hline
-    Usage type & File open/stat/directory listing.\\ \hline
-    \end{tabular}
-    \caption{Description of different axes in File specification distance}
-    \label{tab:distance_axes}
-\end{table}
-
 \subsection{Unique Consistency Requirements}
@@
 revisions.
-As the revision number if global, any change to the repository
+As the revision number is global, any change to the repository
 is reflected in a change to the revision number, which means 
-some software is compatible between revisions.
+some software is unchanged between revisions.
 This method leaves much to be desired when using common stable
@@
 CVMFS & & ls -lr & & & Shrinkwrap & \\
-Repository & Total Size (B) & inode Count & Time to walk & 
-  Data Files & Deduped Data Files & Dedup Bytes \\ \hline
+Repository & Total Size (MB) & inode Count & Time to walk & 
+  Data Files & Deduplicated Data Files & Deduplicated Data (MB) \\ \hline
 %alice-ocdb & 482,642,307,428 & 1,621,661 & 2:22 &
 % 1,621,282 & 1,595 & 24,433,131,203\\
-alice & 3,980,875,747,986 & 27,322,350 & 51:53 & 
-  797,404 & 23,996,541 & 2,098,509,015,172\\
+%alice & 3,980,875,747,986 & 27,322,350 & 51:53 & 
+%  797,404 & 23,996,541 & 2,098,509,015,172\\
+alice & 3,980,875 & 27,322,000 & 51:53 & 
+  797,000 & 23,997,000 & 2,098,509\\
 %atlas-condb & 719,548,941,240 & 11,781 & 0:04 &
 %  10,940 & 229 & 471,492,086 \\
-atlas & 6,195,013,689,056 & 120,598,601 & 3:48:07 &
- 8,823,179 & 85,554,444 & 3,760,712,059,548 \\
-boss & 120,048,527,315 & 2,418,025 & 4:18 &
- 256,061 & 1,868,192 & 72,301,955,666\\
-cms & 581,789,406,847 & 26,224,567 & 1:07:30 &
- 18,667,178 & 101,448,867 & *2,242,658,825,701\\
-cvmfs-config & 16,546 & 9 & 0:00 &
- 2 & 0 & 0 \\
-geant4 & 297,571,283,648 & 7,128,024 & 10:36 &
- 323,280 & 6,219,891 & 68,381,969,660 \\
-grid & 30,642,826,259 & 714,235 & 1:17 &
- 141,622 & 583,923 & 18,804,003,040 \\
-lhcb & 2,134,893,079,440 & 33,736,802 & 1:01:26 &
- 10,032,339 & 20,406,235 & 368,070,277,634 \\
-na61 & 5,222,967,734 & 25,602 & 0:02 &
- 8,609 & 15,001 & 2,425,507,431 \\
-sft & 3,996,527,275,237 & 128,608,792 & 4:17:58 &
- 6,821,699 & 63,210,139 & 2,016,609,810,573 \\ 
+%atlas & 6,195,013,689,056 & 120,598,601 & 3:48:07 &
+% 8,823,179 & 85,554,444 & 3,760,712,059,548 \\
+atlas & 6,195,013 & 120,599,000 & 3:48:07 &
+ 8,823,000 & 85,554,000 & 3,760,712 \\
+%boss & 120,048,527,315 & 2,418,025 & 4:18 &
+% 256,061 & 1,868,192 & 72,301,955,666\\
+boss & 120,048 & 2,418,000 & 4:18 &
+ 256,000 & 1,868,000 & 72,302\\
+%cms & 581,789,406,847 & 26,224,567 & 1:07:30 &
+% 18,667,178 & 101,448,867 & *2,242,658,825,701\\
+cms & 581,789 & 26,225,000 & 1:07:30 &
+ 18,667,000 & 101,449,000 & *2,242,658\\
+%geant4 & 297,571,283,648 & 7,128,024 & 10:36 &
+% 323,280 & 6,219,891 & 68,381,969,660 \\
+geant4 & 297,571 & 7,128,000 & 10:36 &
+ 323,000 & 6,220,000 & 68,382 \\
+%grid & 30,642,826,259 & 714,235 & 1:17 &
+% 141,622 & 583,923 & 18,804,003,040 \\
+grid & 30,643 & 714,000 & 1:17 &
+ 142,000 & 584,000 & 18,804 \\
+%lhcb & 2,134,893,079,440 & 33,736,802 & 1:01:26 &
+% 10,032,339 & 20,406,235 & 368,070,277,634 \\
+lhcb & 2,134,893 & 33,737,000 & 1:01:26 &
+ 10,032,000 & 20,406,000 & 368,070 \\
+%na61 & 5,222,967,734 & 25,602 & 0:02 &
+% 8,609 & 15,001 & 2,425,507,431 \\
+na61 & 5,222 & 26,000 & 0:02 &
+ 9,000 & 15,000 & 2,426 \\
+%sft & 3,996,527,275,237 & 128,608,792 & 4:17:58 &
+% 6,821,699 & 63,210,139 & 2,016,609,810,573 \\ 
+sft & 3,996,527 & 128,609,000 & 4:17:58 &
+ 6,822,000 & 63,210,000 & 2,016,610 \\ 
  \hline
@@
 CVMFS is able to flexibly move
-between these revision using the FUSE module
+between these revisions using the FUSE module
 to determine the structure and files and
@@
 projections.
-A projection is a concrete filesystem that can be
+A projection is a concrete filesystem image that can be
 mounted, loaded, or copied into an existing system.
@@
 was designed for general filesystem synchronization
-rsync flattens the files introducing duplicate data.
+rsync flattens the files, introducing duplicate data.
 This direct representation of the file tree also
@@
 
-GRAPH COMPARING BUILD TIME
-
-GRAPH COMPARING STORAGE SPACE
-
 When creating these images for analysis jobs its often the case that several
@@
 
-\begin{figure}[h]
-\includegraphics[width=\columnwidth]{plots/specification_size.png}
-\caption{Additive effect of host many images.}
-\label{figure:spec-size}
-\end{figure}
-
 \subsection{Use Cases of Shrinkwrap}
@@
 \subsubsection{Benchmarking Containers}
-For benchmarking purposes, the CERN IT-Department is interested in a ”Benchmark in a Container” solution which allows the distribution of standalone container images for system performance evaluations.
-For such benchmarks the IT-Department typically relies on simulation work- flows provided by CERN’s experiment collaborations.
+For benchmarking purposes, the CERN IT-Department is interested in a ``Benchmark in a Container'' solution which allows the distribution of standalone container images for system performance evaluations.
+For such benchmarks the IT-Department typically relies on simulation workflows provided by CERN's experiment collaborations.
 To execute these benchmarks in an easily portable and reproducible manner, it is of high interest to package these software workflows into containers.
-Since the workflows’ software stacks are usually stored inside CVMFS repositories, this is a perfect use case for the shrinkwrap utility and its Docker injection functionality.
+Since the workflows' software stacks are usually stored inside CVMFS repositories, this is a perfect use case for the shrinkwrap utility and its Docker injection functionality.
 
@@
 Unfortunately, this solution is very wasteful of computing time and storage space.
-Even a minimal image consisting of a precise set of required libraries and frameworks ranges in the tens to hundreds of gigabytes,
-depending on the applicaion.
+Even a minimal image consisting of a precise set of required libraries and frameworks ranges in the tens of gigabytes,
+depending on the application.
 The full-repo images being tested at NERSC are multiple terabytes in size and take the majority of a day to create.
@@
 \fbox{\includegraphics[width=0.95\textwidth]{drawings/system_architecture.pdf}}
-\caption{System Architecture}
+\caption{Architecture of a Site-wide Image Management System}
 \label{figure:sys-arch}
@@
 Next, we use MinHash to approximate the Jaccard distance between the new projection and all projections previously built.
+This computed distance can be cached to ensure that we estimate the distance between each pair of images only once.
 Since comparisons using MinHash take constant time for a given probability bound independent of the size of the sets involved,
@@
 This $\alpha$ parameter is something like the ``globbiness'' of the system.
-Very small choices of $\alpha$ requires that projections be extremely close before considering them for merging.
+Small choices of $\alpha$ requires that projections be extremely close before considering them for merging.
 In the extreme case with $\alpha = 0$,
@@
 we next check each in turn to find either a superset of the requested projection or one without version incompatibilities.
+We first sort the list of candidate images by Jaccard distance.
+This ensures that our system prioritizes candidates that are similar in size to the requested image.
 In the case of finding a superset,
@@
 Otherwise, on finding a compatible cached projection the next step is to augment it with the contents of the new request.
+We choose the first compatible image from the list of candidates for merging,
+again prioritizing images that are similar in size and contents.
 As previously discussed,
@@
 accumulating infrequently used dependencies and increasing overhead indefinitely for future tasks.
-We can take advantage of LRU here to ensure that bad merges are eventually removed.
-Since limited storage is available for container images,
-our original approach removed infrequently used images by LRU when necessary.
-We thus define the age of an image to be the maximum of the ages of its constituent images.
-In the case of the frequently used container image merged with a one-off job,
-the one-off job would define the age of the container image since it is never used again.
-Despite having frequently used components,
-this image would be one of the first candidates for removal.
-Removing old images in this way therefore helps to limit the increased overhead of bad merging decisions.
-For a merged image to remain in cache indefinitely,
-its components would need to all be used regularly.
+The Jaccard distance metric gives a natural way to capture and quantify this effect.
+As an image becomes bloated due to repeated merges,
+its distance from every individual request increases.
+After sufficient growth,
+the image will become too far from any request to even be considered.
+Without regular use,
+the bloated image will eventually be removed in LRU order.
+Choice of $\alpha$ allows an upper limit on the amount of useless bloat in images.
 
@@
 The first step in simulating system behavior over time was generating synthetic requests for images.
+We tested two schemes for generating image requests taking into account the properties of the software collections,
+as well as recorded user activity patterns.
+We also generated images via a uniform random scheme for comparison.
 We consulted with the developers of CVMFS,
-as well as HEP researchers at the University of Notre Dame collaborating with CERN.
+as well as HEP researchers at the University of Notre Dame collaborating with CERN to determine how current users interact with CVMFS.
 Since CVMFS is not currently widely deployed on HPC resources,
@@
 with one of the setups available publicly at \url{https://github.com/wyang007/atlas-fat-container}.
-Based on correspondence with researchers and existing applications running on the WLCG,
+
+Based on correspondence with researchers,
 we made some general observations about the behaviors and requirements of applications we need to support.
@@
 
-We use these observations about the structure and behavior of applications on CVMFS to a simplified simulation of dependency requirements for evaluating our container image management scheme.
+We use these observations about the structure and behavior of applications on CVMFS to build a simplified simulation taking into account dependency requirements for evaluating our container image management scheme.
 For the purpose of simulation,
@@
 We are intentionally somewhat vague here to allow for both explicitly defined software components and tracing outputs.
-We examined the software collections available in the ATLAS repo and determined that the median size of a software component around 150~MB,
-which we used as the baseline for simulated components.
-This is a significant simplification,
-as CVMFS hosts some very large components and files.
-Looking at the distribution of file sizes we measured for the major repos (figures M-N),
-we observed that these extreme outlier files are fairly rare.
-Our next assumption is that certain components are included in most container images.
+We examined the software collections available in the SFT repo and constructed a DAG of dependencies among the packages.
+The currently available software collection in the SFT repo consists of NNN packages,
+where a program or library typically provides packages for multiple versions and configurations.
+When building a projection,
+we recursively include dependencies of requested software.
+This more closely approximates the structures of actual container images,
+while still allowing for randomly selected package requests.
+For each simulated image request,
+we chose a random selection of packages (either uniform random selection or subject to the process below)
+and then added the closure of the package dependencies.
+This first image simulation scheme captures the structure inherent in the software collection,
+in that packages in addition to those requested are automatically included so as to ensure a functional image.
+The initial selection of packages,
+however, is simply uniformly random.
+This is not necessarily realistic, however,
+so we next took usage patterns into account.
+
+We assumed that certain components are included in most container images,
+while other components are relatively rarely used.
 While multiple versions and variations might exist,
@@
 For the purposes of simulation,
-we expect these components to be very widely duplicated in jobs across all users and experiments.
-Corresponding to the full software collections maintained by each experiment,
-we assume that there will be a larger set of components that is less uniformly included in images.
-Each user or group might have a set of tools that they use heavily,
-or jobs will request particular versions of libraries and components out of many avialable in the software collection.
-
-To simulate requests similar to these usage patterns,
-we first created two pools of components/packages: core and extra components.
-We surveyed the software collections available on CVMFS and chose pools of 200 core components and 3000 extra components to approximate the size and number of software components currently available on CVMFS.
-Of these core components,
-due to the availability of multiple versions and some duplication across experiments,
-we estimate that roughly half of these core components would be required to make a functional container image.
-Thus for the simulated requests,
-we included a random selection of 100 core components.
-This represents a fairly pathological case,
-as actual depencencies would show less random behavior and variation here.
-The simulation therefore tests near-worst case behavior.
-For the extra components,
-we surveyed available software collections,
-especially those of the ATLAS and CMS experiments,
-and chose a simulated pool size of 3000 extra components.
-For simulated image requests,
-we included a random selection of up to 10\% of the extra pool.
-This is again a worst case behavior,
-as actual requests would more ordered.
-We did not allow for the entire pool to be included as there is no compelling use case for this behavior.
-Even the multi-terabyte images being tested at NERSC are taken only from a single repo,
-the CMS repo.
-There is no analysis or simulation software that uses complete software stacks from multiple repos;
-each experiment maintains its own self-contained software collection.
-Thus aside from the core components that may be used in common across experiments,
-we do not expect to see images that require more than a fraction of the total available on CVMFS.
+we would like to see these components in a large proportion of jobs across all users and experiments.
+
+We obtained CVMFS usage logs from frontend nodes at CERN,
+which give rough usage frequencies of the packages in the SFT repo.
+To generate simulated container requests,
+we made random package selections subject to the observed distribution of package use.
+Many packages in the repo were not observed being used at all,
+so we added a small, non-zero probability of selecting packages lacking usage data.
+We applied this technique in combination with the package dependency graph,
+so that images contained a weighted random core selection along with any dependent packages.
+This second image simulation scheme takes into account the dependency structure of the repo as well as usage patterns observed in the wild.
+
+As a base for comparison,
+we also generated simulated images consisting of packages chosen in a uniform random way.
+To ensure that total size (or at least total number of packages)
+is comparable to images generated by the previous two methods,
+for this approach we started with an image request generated via the first scheme discussed
+(uniform random core selection with dependencies added).
+We considered only the number of software packages in the resulting image,
+and then chose the same number of packages uniformly randomly from the entire repo,
+ignoring usage information and package dependencies.
+while images generated in this way are not particularly realistic,
+they will be useful for evaluating image management schemes.
+By comparing results with random images to those with the previous two image generation schemes,
+we can compare the general case of containers as collections of arbitrary data to the specific focus of this work,
+i.e.\ containers with selections of experimental software.
 
@@
 It is the responsibility of the image management system to provide a container image satisfying the requirements of each request.
-Let the set $C$ be the union of all components in all requests.
+Let the set $C$ be the union of all components in one or more requests.
 $C$ provides a lower bound on the amount of space required to serve all requests at the same time.
 If the storage space required to hold all of $C$ is greater than the total available storage,
-then new requests will requre removal of container image(s) for previous requests.
+then images serving new requests will requre removal of container image(s) for previous requests.
 Since we have been considering image management as an online problem,
@@
 The image size is very large and any changes/updates are expensive both in terms of compute and bandwidth.
-This approach does, however,
+NERSC's approach does, however,
 avoid duplication of data as there is a single image that serves all users and applications.
-We next define the multiset $A$ as the union (with repetition) of all components in all requested images.
+
+We thus define the multiset $A$ as the union (with repetition) of all components in a set of requested images.
 If multiple images request the same component,
@@
 is to achieve storage cost as close to $C$ as possible without excessive overhead in terms of compute and network bandwidth to keep the image collection up to date.
-We will quantify this duplication ratio $d$ as $|A| / |C|$.
-For a system with no duplicated data in its container image(s),
-$d=1$.
-Larger values of $d$ indicate duplicated data.
+We define \emph{cache efficiency} as $|C| / |A|$ to quantify the amount of duplication in the system cache.
+With poor cache efficiency,
+the available storage for the system cache is wasted on repeated copies of the same data.
+Improving cache efficiency allows for more more unique image data within the same storage space,
+or equivalently allows the same set of job requests to be served efficiently from a smaller cache.
 
@@
 
-\bibliographystyle{IEEEtran}
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{otherpapers,cclpapers}

### 2019-01-23 paper.tex
Update paper build and add curated dir

@@
 
-\begin{figure}
-\includegraphics[width=\linewidth]{plots/file-hist/alice-ocdb.pdf}
-\caption{Distribution of file sizes in the \texttt{alice-ocdb.cern.ch} repo.
-The vertical line at $2^{20}$~B marks a file size of 1~MB.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{plots/file-hist/alice.pdf}
-\caption{Distribution of file sizes in the \texttt{alice.cern.ch} repo.
-The vertical line at $2^{20}$~B marks a file size of 1~MB.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{plots/file-hist/atlas-condb.pdf}
-\caption{Distribution of file sizes in the \texttt{atlas-condb.cern.ch} repo.
-The vertical line at $2^{20}$~B marks a file size of 1~MB.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{plots/file-hist/atlas.pdf}
-\caption{Distribution of file sizes in the \texttt{atlas.cern.ch} repo.
-The vertical line at $2^{20}$~B marks a file size of 1~MB.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{plots/file-hist/boss.pdf}
-\caption{Distribution of file sizes in the \texttt{boss.cern.ch} repo.
-The vertical line at $2^{20}$~B marks a file size of 1~MB.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{plots/file-hist/cms.pdf}
-\caption{Distribution of file sizes in the \texttt{cms.cern.ch} repo.
-The vertical line at $2^{20}$~B marks a file size of 1~MB.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{plots/file-hist/cvmfs-config.pdf}
-\caption{Distribution of file sizes in the \texttt{cvmfs-config.cern.ch} repo.
-The vertical line at $2^{20}$~B marks a file size of 1~MB.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{plots/file-hist/geant4.pdf}
-\caption{Distribution of file sizes in the \texttt{geant4.cern.ch} repo.
-The vertical line at $2^{20}$~B marks a file size of 1~MB.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{plots/file-hist/grid.pdf}
-\caption{Distribution of file sizes in the \texttt{grid.cern.ch} repo.
-The vertical line at $2^{20}$~B marks a file size of 1~MB.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{plots/file-hist/lhcb.pdf}
-\caption{Distribution of file sizes in the \texttt{lhcb.cern.ch} repo.
-The vertical line at $2^{20}$~B marks a file size of 1~MB.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{plots/file-hist/na61.pdf}
-\caption{Distribution of file sizes in the \texttt{na61.cern.ch} repo.
-The vertical line at $2^{20}$~B marks a file size of 1~MB.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{plots/file-hist/sft.pdf}
-\caption{Distribution of file sizes in the \texttt{sft.cern.ch} repo.
-The vertical line at $2^{20}$~B marks a file size of 1~MB.}
-\end{figure}
+\begin{figure*}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_cache_percent_plt.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_cache_percent_plt.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_request_percent_plt.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_request_percent_plt.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_uses_percent_plt.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_uses_percent_plt.pdf}
+\caption{}
+\end{figure*}
 

### 2019-01-23 curated/cdf/clean.py

@@ -1,0 +1,8 @@
+#!/usr/bin/env python
+import fileinput
+
+for a in fileinput.input():
+    log, count = a.split()
+    log = int(log)
+    count = int(count)
+    print("{} {}".format(2**log, count/1e6))

### 2019-01-23 curated/closure/closure.py

@@ -1,0 +1,69 @@
+#!/usr/bin/env python
+
+import sys
+import json
+import random
+
+SAMPLES = 100
+PICKS = 1000
+
+def median(lst):
+    # lst must already be sorted!
+    if len(lst) == 0: return 0
+    if len(lst) == 1: return lst[0]
+    i = len(lst) // 2
+    if len(lst) % 2 == 0:
+        return (lst[i] + lst[i - 1]) / 2
+    else:
+        return lst[i]
+
+def smedian(lst):
+    lst.sort()
+    return median(lst)
+
+def size(deps, img):
+    accum = 0
+    for pkg in img:
+        accum += median(deps[pkg]['size'])
+    return accum
+
+def closure(deps, pkg, res):
+    if pkg in res: return
+    res.add(pkg)
+    for p in deps[pkg]['deps']:
+        closure(deps, p, res)
+
+def blind(deps, img):
+    return set(random.sample(deps.keys(), len(img)))
+
+choice_sizes = [list() for i in range(PICKS)]
+closure_sizes = [list() for i in range(PICKS)]
+choice_counts = [list() for i in range(PICKS)]
+closure_counts = [list() for i in range(PICKS)]
+
+deps = json.load(sys.stdin)
+
+for i in range(SAMPLES):
+    choice = set()
+    out = set() 
+    sys.stderr.write('{} '.format(i))
+    for j in range(PICKS):
+        new_pkg = random.choice(deps.keys())
+        choice.add(new_pkg)
+        closure(deps, new_pkg, out)
+        choice_counts[j].append(len(choice))
+        closure_counts[j].append(len(out))
+        choice_sizes[j].append(size(deps, choice))
+        closure_sizes[j].append(size(deps, out))
+
+print('picks choicecount closurecount choicesize closuresize')
+for i in range(PICKS):
+    print('{} {} {} {} {}'.format(
+        i,
+        smedian(choice_counts[i]),
+        smedian(closure_counts[i]),
+        smedian(choice_sizes[i]),
+        smedian(closure_sizes[i]),
+    ))
+
+sys.stderr.write('\n')

### 2019-01-23 curated/comparative/dat.py

@@ -1,0 +1,60 @@
+#!/usr/bin/env python
+
+import sys
+import json
+
+INST = ['overhead', 'efficiency', 'images', 'imagesizes', 'pkgs',
+        'realsize', 'reqsize', 'size', 'upkgs', 'usize']
+CUM = ['byteswritten', 'deletes', 'inserts', 'merges', 'hits', 'hitrate',
+       'peaksize', 'reqwritten']
+
+def smedian(lst):
+    lst.sort()
+    return median(lst)
+
+
+def median(lst):
+    # lst must already be sorted!
+    if len(lst) == 0: return 0
+    if len(lst) == 1: return lst[0]
+    i = len(lst) // 2
+    if len(lst) % 2 == 0:
+        return (lst[i] + lst[i - 1]) / 2
+    else:
+        return lst[i]
+
+def sample_final(samples, field):
+    return samples[-1][field]
+
+def sample_median(samples, field):
+    values = [s[field] for s in samples]
+    values.sort()
+    return median(values)
+
+
+raw = json.load(sys.stdin)[sys.argv[1]]
+dat = dict()
+
+for alpha, runs in raw.items():
+    dat[float(alpha)] = runs
+    for stream in runs:
+        p = 0
+        rw = 0
+        for sample in stream:
+            sample['overhead'] = sample['realsize'] - sample['reqsize']
+            sample['efficiency'] = float(sample['reqsize'])/sample['realsize']
+            rw += sample['reqsize']
+            sample['reqwritten'] = rw
+            p = max(sample['size'], p)
+            sample['peaksize'] = p
+            sample['hitrate'] = float(sample['hits'])/(sample['hits']+sample['inserts']+sample['merges'])
+
+print(' '.join(['alpha'] + INST + CUM))
+
+alphas = dat.keys()
+alphas.sort()
+
+for i in alphas:
+    cc = [str(smedian([sample_final(stream, f) for stream in dat[i]])) for f in CUM]
+    ci = [str(smedian([sample_median(stream, f) for stream in dat[i]])) for f in INST]
+    print(' '.join([str(i)] + ci + cc))

### 2019-01-23 curated/time-series/ser.py

@@ -1,0 +1,19 @@
+#!/usr/bin/env python
+
+import sys
+import json
+
+def extract(samples, field, step):
+    values = [s[step][field] for s in samples]
+    values.sort()
+    return sim.median(values)
+
+dat = json.load(sys.stdin)[sys.argv[1]][sys.argv[2]][int(sys.argv[3])]
+
+headers = dat[0].keys()
+headers.sort()
+print(' '.join(['step'] + headers))
+
+for i in range(len(dat)):
+    cols = [str(dat[i][c]) for c in headers]
+    print(' '.join([str(i)] + cols))

### 2019-01-23 simulation/dat.py

@@ -5,10 +5,10 @@ import json
 
 import sim
 
-INST = ['overhead', 'efficiency', 'images', 'imagesizes', 'pkgs',
+INST = ['overhead', 'container_efficiency', 'cache_efficiency', 'images', 'imagesizes', 'pkgs',
         'realsize', 'reqsize', 'size', 'upkgs', 'usize']
-CUM = ['byteswritten', 'deletes', 'inserts', 'merges', 'hits',
-       'peaksize']
+CUM = ['byteswritten', 'deletes', 'inserts', 'merges', 'hits', 'hitrate',
+       'peaksize', 'reqwritten']
 
 def sample_final(samples, field):
     return samples[-1][field]
@@ -26,11 +26,16 @@ for alpha, runs in raw.items():
     dat[float(alpha)] = runs
     for stream in runs:
         p = 0
+        rw = 0
         for sample in stream:
             sample['overhead'] = sample['realsize'] - sample['reqsize']
-            sample['efficiency'] = float(sample['reqsize'])/sample['realsize']
+            sample['container_efficiency'] = float(sample['reqsize'])/sample['realsize']
+            sample['cache_efficiency'] = float(sample['usize'])/sample['size']
             p = max(sample['size'], p)
             sample['peaksize'] = p
+            rw += sample['reqsize']
+            sample['reqwritten'] = rw
+            sample['hitrate'] = float(sample['hits'])/(sample['hits']+sample['inserts']+sample['merges'])
 
 print(' '.join(['alpha'] + INST + CUM))
 

### 2019-01-23 simulation/sim.py

@@ -81,9 +81,10 @@ class BlindStream:
         return frozenset(blind(self.deps, self.stream.next()))
 
 class Cache:
-    def __init__(self, deps, alpha, reuse, capacity):
+    def __init__(self, deps, alpha, reuse, capacity, container_limit):
         self.reuse = reuse
         self.capacity = capacity
+        self.container_limit = container_limit
         self.deps = deps
         self.size = 0
         self.alpha = alpha
@@ -149,10 +150,13 @@ class Cache:
         }
 
     def merge(self, existing, img):
-        self.size -= self.contents.pop(existing)
         new_img = existing | img
-        self.contents[new_img] = size(self.deps, new_img)
+        img_size = size(self.deps, new_img)
+        if self.container_limit > 0 and img_size > self.container_limit:
+            return
+        self.contents[new_img] = img_size
         self.size += self.contents[new_img]
+        self.size -= self.contents.pop(existing)
         self.bytes_written += self.contents[new_img]
         self.merges += 1
         return size(self.deps, img), self.contents[new_img], new_img
@@ -189,8 +193,10 @@ class Cache:
             if self.issubset(img, a[1]):
                 return self.hit(a[1], img)
 
-        if len(candidates) > 0:
-            return self.merge(candidates[0][1], img)
+        for a in candidates:
+            merge_result = self.merge(a[1], img)
+            if merge_result is not None:
+                return merge_result
 
         return self.insert(img)
 
@@ -220,18 +226,20 @@ class Cache:
         return self.log
 
 def run(params):
-    alpha, deps, deps_freq, reuse, jobs, capacity = params
+    alpha, deps, deps_freq, reuse, jobs, capacity, container_limit, tree, blind = params
     out = {}
-    for i in range(10):
-        b = Cache(deps, alpha, reuse, capacity)
-        #c = Cache(deps, alpha, reuse, capacity)
-        #d = Cache(deps, alpha, reuse, capacity)
+    for i in range(20):
+        b = Cache(deps, alpha, reuse, capacity, container_limit)
         b.process(itertools.islice(DistStream(deps, deps_freq), jobs))
-        #c.process(itertools.islice(Stream(deps), jobs))
-        #d.process(itertools.islice(BlindStream(deps), jobs))
         out['dist'] = b.log
-        #out['tree'] = c.log
-        #out['blind'] = d.log
[…]

### 2019-01-23 curated/cdf/clean.py
Updates from ShareLaTeX

@@ -1,8 +1,0 @@
-#!/usr/bin/env python
-import fileinput
-
-for a in fileinput.input():
-    log, count = a.split()
-    log = int(log)
-    count = int(count)
-    print("{} {}".format(2**log, count/1e6))

### 2019-01-23 curated/closure/closure.py
delete stuff

@@ -1,69 +1,0 @@
-#!/usr/bin/env python
-
-import sys
-import json
-import random
-
-SAMPLES = 100
-PICKS = 1000
-
-def median(lst):
-    # lst must already be sorted!
-    if len(lst) == 0: return 0
-    if len(lst) == 1: return lst[0]
-    i = len(lst) // 2
-    if len(lst) % 2 == 0:
-        return (lst[i] + lst[i - 1]) / 2
-    else:
-        return lst[i]
-
-def smedian(lst):
-    lst.sort()
-    return median(lst)
-
-def size(deps, img):
-    accum = 0
-    for pkg in img:
-        accum += median(deps[pkg]['size'])
-    return accum
-
-def closure(deps, pkg, res):
-    if pkg in res: return
-    res.add(pkg)
-    for p in deps[pkg]['deps']:
-        closure(deps, p, res)
-
-def blind(deps, img):
-    return set(random.sample(deps.keys(), len(img)))
-
-choice_sizes = [list() for i in range(PICKS)]
-closure_sizes = [list() for i in range(PICKS)]
-choice_counts = [list() for i in range(PICKS)]
-closure_counts = [list() for i in range(PICKS)]
-
-deps = json.load(sys.stdin)
-
-for i in range(SAMPLES):
-    choice = set()
-    out = set() 
-    sys.stderr.write('{} '.format(i))
-    for j in range(PICKS):
-        new_pkg = random.choice(deps.keys())
-        choice.add(new_pkg)
-        closure(deps, new_pkg, out)
-        choice_counts[j].append(len(choice))
-        closure_counts[j].append(len(out))
-        choice_sizes[j].append(size(deps, choice))
-        closure_sizes[j].append(size(deps, out))
-
-print('picks choicecount closurecount choicesize closuresize')
-for i in range(PICKS):
-    print('{} {} {} {} {}'.format(
-        i,
-        smedian(choice_counts[i]),
-        smedian(closure_counts[i]),
-        smedian(choice_sizes[i]),
-        smedian(closure_sizes[i]),
-    ))
-
-sys.stderr.write('\n')

### 2019-01-23 curated/comparative/dat.py

@@ -1,60 +1,0 @@
-#!/usr/bin/env python
-
-import sys
-import json
-
-INST = ['overhead', 'efficiency', 'images', 'imagesizes', 'pkgs',
-        'realsize', 'reqsize', 'size', 'upkgs', 'usize']
-CUM = ['byteswritten', 'deletes', 'inserts', 'merges', 'hits', 'hitrate',
-       'peaksize', 'reqwritten']
-
-def smedian(lst):
-    lst.sort()
-    return median(lst)
-
-
-def median(lst):
-    # lst must already be sorted!
-    if len(lst) == 0: return 0
-    if len(lst) == 1: return lst[0]
-    i = len(lst) // 2
-    if len(lst) % 2 == 0:
-        return (lst[i] + lst[i - 1]) / 2
-    else:
-        return lst[i]
-
-def sample_final(samples, field):
-    return samples[-1][field]
-
-def sample_median(samples, field):
-    values = [s[field] for s in samples]
-    values.sort()
-    return median(values)
-
-
-raw = json.load(sys.stdin)[sys.argv[1]]
-dat = dict()
-
-for alpha, runs in raw.items():
-    dat[float(alpha)] = runs
-    for stream in runs:
-        p = 0
-        rw = 0
-        for sample in stream:
-            sample['overhead'] = sample['realsize'] - sample['reqsize']
-            sample['efficiency'] = float(sample['reqsize'])/sample['realsize']
-            rw += sample['reqsize']
-            sample['reqwritten'] = rw
-            p = max(sample['size'], p)
-            sample['peaksize'] = p
-            sample['hitrate'] = float(sample['hits'])/(sample['hits']+sample['inserts']+sample['merges'])
-
-print(' '.join(['alpha'] + INST + CUM))
-
-alphas = dat.keys()
-alphas.sort()
-
-for i in alphas:
-    cc = [str(smedian([sample_final(stream, f) for stream in dat[i]])) for f in CUM]
-    ci = [str(smedian([sample_median(stream, f) for stream in dat[i]])) for f in INST]
-    print(' '.join([str(i)] + ci + cc))

### 2019-01-23 curated/time-series/ser.py
Updates from ShareLaTeX

@@ -1,19 +1,0 @@
-#!/usr/bin/env python
-
-import sys
-import json
-
-def extract(samples, field, step):
-    values = [s[step][field] for s in samples]
-    values.sort()
-    return sim.median(values)
-
-dat = json.load(sys.stdin)[sys.argv[1]][sys.argv[2]][int(sys.argv[3])]
-
-headers = dat[0].keys()
-headers.sort()
-print(' '.join(['step'] + headers))
-
-for i in range(len(dat)):
-    cols = [str(dat[i][c]) for c in headers]
-    print(' '.join([str(i)] + cols))

### 2019-01-23 paper.tex
Updates from ShareLaTeX
Added curated
Update to paper state

@@
 
-Shrinkwrap provides a mechinsm for projecting parts of CVMFS,
+Shrinkwrap provides a mechanism for projecting parts of CVMFS,
 but is not by itself a complete solution as it does not integrate with job submission or manage images after creation.
@@
 
+\subsection{Simulating Job Submissions}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/cdf/cdf.pdf}
+\label{fig:cdf}
+\if 0
+lots of small files
+table shows duplication
+long tail of large files
+different scales, similar usage across repos
+
+\fi
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/closure/closure.pdf}
+\label{fig:picks}
+\if 0
+you get more than you ask for
+smaller requirements have relatively large responses
+non-linear, so bigger ones have more duplicates
+appraoches repo size in limit
+merging is subject to drop-off(more duplication)
+\fi
+\end{figure}
+
+To evaluate the effect of including dependencies in container images,
+we made a large number of random selections of packages from the SFT repo and then compared with the set of dependencies required by each selection.
+Figure~\ref{fig:picks} shows the results over a large number of selection sizes.
+We observe of course that selection size increases uniformly along the $x$ axis.
+The on-disk size of the selections appears to increase proportionally to selection size.
+Recursively including the dependencies of these same selections,
+however, results in a significant increase in number of packages and size on disk.
+For small selections (less than 100~packages),
+the complete images with dependencies might contain 5x as many packages as requested,
+with storage increasing proportionally.
+With larger selections, however,
+this increase becomes less prominent.
+This non-linear change is the result of the tree structure of the software dependencies.
+There are a number of core components that are transitive dependencies of a large number of packages.
+These common dependencies are only included once,
+so subsequent selections that depend on them do not add them again.
+As the number of selections increases,
+this curve will approach the total repo size.
+Also note that the non-linear increase in size with increasing selection size means that the size of combining two selections of 100 packages (and their dependencies) is not as simple as adding them together.
+Due to duplication, we would expect the resulting combined size to be lower,
+as if choosing 200 packages from the start.
+
+\if 0
+spread across everyone, lot of commonality/overlap
+few with very high usage, very large number of rarely used
+lot not observed
+\fi
+
+Since there are currently few regular CVMFS on HPC users,
+we decided to simulate the activity of users requesting container images.
+We decided to simulate at the granularity of packages rather than individual files,
+because this results in more robust container images.
+As mentioned previously,
+fine-grained tracing runs the risk of leaving out important dependencies.
+In addition, working only based on the properties of individual files ignores the structure inherent in software collections.
+Looking only at the properties of individual files,
+Figure~\ref{fig:cdf} shows cumulative distributions for file sizes in several of the largest and most active CVMFS repos.
+The repos vary in scale,
+but all show similar usage patterns.
+We see that across all repos,
+the majority of files are small (between 1~KB and 1~MB).
+The repos do contain, however,
+a small number of large files.
+This fits our expectation that the CVMFS repos primarily contain software.
+Most individual files are small,
+but the repo size is defined by the large number of packages in the software collection.
+
+We tested two schemes for generating image requests taking into account the properties of the software collections,
+as well as recorded user activity patterns.
+We also generated images via a uniform random scheme for comparison.
+We consulted with the developers of CVMFS,
+as well as HEP researchers at the University of Notre Dame collaborating with CERN to determine how current users interact with CVMFS.
+Since CVMFS is not currently widely deployed on HPC resources,
+the prospective user base for our container management system is limited.
+We were, however, informed of some researchers and groups experimenting with static mirrors of CVMFS for container images,
+including projects at NERSC and Stanford University,
+with one of the setups available publicly at \url{https://github.com/wyang007/atlas-fat-container}.
+
+Based on correspondence with researchers,
+we made some general observations about the behaviors and requirements of applications we need to support.
+First, there are some components such as ROOT that serve as common, foundational components to almost every other simulation and analysis stack.
+These core components are often stored separately from the rest of the software collections available in CVMFS.
+Second, there are shared components such as setup scripts, configuration files, and calibration or condition data (required to interpret readings from detectors) that are also nearly universally required but are more particular to a given experiment or group.
+Finally, there is the collection of software maintained by each experiment.
+Software collections may contain multiple copies of the same software for different architectures and versions.
+A single analysis job does not use the majority of the available software.
+There is some overlap among these categories of software,
+for example some software components are widely enough used to be considered core components,
+and usage and availability of components varies over time.
+
+We use these observations about the structure and behavior of applications on CVMFS to build a simplified simulation taking into account dependency requirements for evaluating our container image management scheme.
+For the purpose of simulation,
+we assume that the requirements of a job are given as discrete components or collections of files.
+While Shrinkwrap can operate at the granularity of individual files,
+both static analysis and runtime tracing produce file collections to include.
+We are intentionally somewhat vague here to allow for both explicitly defined software components and tracing outputs.
+We examined the software collections available in the SFT repo and constructed a DAG of dependencies among the packages.
+The currently available software collection in the SFT repo consists of NNN packages,
+where a program or library typically provides packages for multiple versions and configurations.
+When building a projection,
+we recursively include dependencies of requested software.
+This more closely approximates the structures of actual container images,
+while still allowing for randomly selected package requests.
+For each simulated image request,
+we chose a random selection of packages (either uniform random selection or subject to the process below)
+and then added the closure of the package dependencies.
+This first image simulation scheme captures the structure inherent in the software collection,
+in that packages in addition to those requested are automatically included so as to ensure a functional image.
+The initial selection of packages,
+however, is simply uniformly random.
+This is not necessarily realistic, however,
+so we next took usage patterns into account.
+
+We assumed that certain components are included in most container images,
+while other components are relatively rarely used.
+While multiple versions and variations might exist,
+these components have a very high likelihood of appearing in every container image.
+These components correspond to the core components, setup scripts, calibration data, etc.\ described previously.
+For the purposes of simulation,
+we would like to see these components in a large proportion of jobs across all users and experiments.
+
+We obtained CVMFS usage logs from frontend nodes at CERN,
+which give rough usage frequencies of the packages in the SFT repo.
+To generate simulated container requests,
+we made random package selections subject to the observed distribution of package use.
+Many packages in the repo were not observed being used at all,
+so we added a small, non-zero probability of selecting packages lacking usage data.
+We applied this technique in combination with the package dependency graph,
+so that images contained a weighted random core selection along with any dependent packages.
+This second image simulation scheme takes into account the dependency structure of the repo as well as usage patterns observed in the wild.
+
+As a base for comparison,
+we also generated simulated images consisting of packages chosen in a uniform random way.
+To ensure that total size (or at least total number of packages)
+is comparable to images generated by the previous two methods,
+for this approach we started with an image request generated via the first scheme discussed
+(uniform random core selection with dependencies added).
+We considered only the number of software packages in the resulting image,
+and then chose the same number of packages uniformly randomly from the entire repo,
+ignoring usage information and package dependencies.
+while images generated in this way are not particularly realistic,
+they will be useful for evaluating image management schemes.
+By comparing results with random images to those with the previous two image generation schemes,
+we can compare the general case of containers as collections of arbitrary data to the specific focus of this work,
+i.e.\ containers with selections of experimental software.
+
 \subsection{An initial approach}
@@
 The simplest approach for automatically projecting container images is to create a suitable image on each batch job submission.
+Under this design,
+on job submission the bridge machine first checks if the same projection had been requested before.
+If so, it arranges to transfer its cached container image to worker nodes that require it.
+If the bridge node does not have that projection built already,
+it uses Shrinkwrap to create the requested projection,
+then invokes mksquashfs to create the container image.
+Once the bridge node places the newly created image in its cache,
+the process proceeds as in the previous case.
+The architecture of this initial approach is shown in figure ???.
+
 Unfortunately, this solution is very wasteful of computing time and storage space.
@@
 this method could be sufficient.
-
-After creating a set of static images,
-the next step is to transfer them to local caches on the workers.
-Worker nodes are assumed to have some fixed amount of storage available for caching images.
-If the size of the ``working set'' of images is smaller than the node-local storage,
-then each worker can keep a complete set of images.
-While ideal, this case is not necessarily realistic beyond a few images.
-With small/medium images in the tens-to-hundreds of gigabytes range,
-maintining a complete set of projected CVMFS images at each worker requires substantial resource provisioning.
-Thus it is important for workers be able to to cache only a subset of the available projected images.
-There are numerous caching policies and options available,
-but choice of worker cache strategy is not critically important for this application.
-Since projections and images of CVMFS can be re-created on demand,
-cache evictions might result in computation overhead but would not otherwise affect the functionality or correctness of the system.
-In the absence of a compelling and site-specific reason,
-a simple caching policy such as LRU would be sufficient.
-As we will see in section ??,
-LRU will also be an important piece of our final proposed approach.
-
-Under this design,
-on job submission the bridge machine first checks if the same projection had been requested before.
-If so, it arranges to transfer its cached container image to worker nodes that lack that image locally.
-If the bridge node does not have that projection built already,
-it uses Shrinkwrap to create the requested projection,
-then invokes mksquashfs to create the container image.
-Once the bridge node places the newly created image in its cache,
-the process proceeds as in the previous case.
-The architecture of this initial approach is shown in figure ???.
-
-%Node-local scratch storage is sufficient for temporarily caching these images.
-%We also assumed that durable storage is available on the bridge machine for storing the data cache and projections associated with Shrinkwrap,
-%as well as a cache of container images derived from these projections.
-%Again, we chose LRU as a caching policy for our evaluation,
-%though administrators would be free to experiment with different policies.
-
-%Technologies like Docker currently operate this way (often to the chagrin of users and administrators doing scientific computing).
 
@@
 
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/time-series/time-series.pdf}
+\label{fig:series}
+\if 0
+zoomed out?
+cache limit, deletes start
+merge tracks writes, it's the dominant factor
+hits grow gradually as cache merges and fills
+deletes help prune useless stuff, remaining should be frequent hits
+far ends are either insert or merge, middle alphas have both
+\fi
+\end{figure}
+
+Figure~\ref{fig:series} shows a single simulation of the image management system,
+with several notable details.
+First, we see that most of the operations are merges.
+This makes sense, as this simulation was at fairly high $\alpha$.
+The total bytes written also closely tracks merges,
+indicating that merging is the dominant source of I/O.
+We can still see inserts over the course of the simulation.
+At more extreme $\alpha$ values,
+we would expect to see one of these operations almost exclusively.
+As the data in cache continues to rise,
+we eventually see the cache limit,
+after which the delete count starts increasing.
+Over the course of the simulation,
+inserts and deletes are filling and emptying the cache such that it remains close to its storage limit.
+We also observe the number of cache hits continue to rise despite deletions.
+As we will see, merging allows for a greater proportion of hits even if the amount of data remains constant.
+This is because frequently used data can be merged,
+reducing duplication.
+The cache limit then ensures that infrequently used parts are eventually removed.
+
+% explain alpha plot
+
+Sweeping over the range of $\alpha$ values in this way,
+we can immediately see differences in the frequency of simulated operations.
+Figure~\ref{fig:opcounts} shows the upper range of $\alpha$ values where behavior differences appear.
+From the lower values on the left,
+the insert and delete counts are the primary (or only) operations,
+with number of hits relatively constant.
+This corresponds to a simple LRU-based cache.
+The insert count is higher due to cache filling,
+but the two tend to increase in lockstep.
+As $\alpha$ increases,
+image merges become more probable.
+The merge count in the figure steadily increases throughout most of the upper range,
+while inserts and deletes decrease.
+This suggests that at high $\alpha$ values,
+the cache space would be more efficiently used,
+with some of the duplication merged out.
+In the extreme case with $\alpha=1$,
+every request is merged into a single image in cache,
+hence the sudden increase in hit rate and decrease in total merges at the far right of Figure~\ref{fig:opcount}.
+Note that as merging increases,
+the proportion of hits generally also increases.
+This indicates that more of the requests can be satisfied by previously merged images in the cache.
+From the perspective of the system,
+this is beneficial as there is no extra time or compute cost for handling a larger proportion of images.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/comparative/operation_count.pdf}
+\label{fig:opcounts}
+\if 0
+inserts and merges switch as alpha goes up
+more hits with merging
+allows better responsiveness? image is immediately ready
+inserts and deletes go to zero, so storage use will taper off
+drop off: we keep everything, so 500 requests get merged; hit count goes up correspondingly since everything is in cache
+not gonna be close to cache limit
+inserts and deletes track: corresponds to LRU, inserts while filling then balanced
+\fi
+\end{figure}
+
 To decide if two projections are ``close'',
@@
 
-\subsection{Simulating Job Submissions}
-
-The first step in simulating system behavior over time was generating synthetic requests for images.
-We tested two schemes for generating image requests taking into account the properties of the software collections,
-as well as recorded user activity patterns.
-We also generated images via a uniform random scheme for comparison.
-We consulted with the developers of CVMFS,
-as well as HEP researchers at the University of Notre Dame collaborating with CERN to determine how current users interact with CVMFS.
-Since CVMFS is not currently widely deployed on HPC resources,
-the prospective user base for our container management system is limited.
-We were, however, informed of some researchers and groups experimenting with static mirrors of CVMFS for container images,
-including projects at NERSC and Stanford University,
-with one of the setups available publicly at \url{https://github.com/wyang007/atlas-fat-container}.
-
-Based on correspondence with researchers,
-we made some general observations about the behaviors and requirements of applications we need to support.
-First, there are some components such as ROOT that serve as common, foundational components to almost every other simulation and analysis stack.
-These core components are often stored separately from the rest of the software collections available in CVMFS.
-Second, there are shared components such as setup scripts, configuration files, and calibration or condition data (required to interpret readings from detectors) that are also nearly universally required but are more particular to a given experiment or group.
-Finally, there is the collection of software maintained by each experiment.
-Software collections may contain multiple copies of the same software for different architectures and versions.
-A single analysis job does not use the majority of the available software.
-There is some overlap among these categories of software,
-for example some software components are widely enough used to be considered core components,
-and usage and availability of components varies over time.
-
-We use these observations about the structure and behavior of applications on CVMFS to build a simplified simulation taking into account dependency requirements for evaluating our container image management scheme.
-For the purpose of simulation,
-we assume that the requirements of a job are given as discrete components or collections of files.
-While Shrinkwrap can operate at the granularity of individual files,
-both static analysis and runtime tracing produce file collections to include.
-We are intentionally somewhat vague here to allow for both explicitly defined software components and tracing outputs.
-We examined the software collections available in the SFT repo and constructed a DAG of dependencies among the packages.
-The currently available software collection in the SFT repo consists of NNN packages,
-where a program or library typically provides packages for multiple versions and configurations.
-When building a projection,
-we recursively include dependencies of requested software.
-This more closely approximates the structures of actual container images,
-while still allowing for randomly selected package requests.
-For each simulated image request,
-we chose a random selection of packages (either uniform random selection or subject to the process below)
-and then added the closure of the package dependencies.
-This first image simulation scheme captures the structure inherent in the software collection,
-in that packages in addition to those requested are automatically included so as to ensure a functional image.
-The initial selection of packages,
-however, is simply uniformly random.
-This is not necessarily realistic, however,
-so we next took usage patterns into account.
-
-We assumed that certain components are included in most container images,
-while other components are relatively rarely used.
-While multiple versions and variations might exist,
-these components have a very high likelihood of appearing in every container image.
-These components correspond to the core components, setup scripts, calibration data, etc.\ described previously.
-For the purposes of simulation,
-we would like to see these components in a large proportion of jobs across all users and experiments.
-
-We obtained CVMFS usage logs from frontend nodes at CERN,
-which give rough usage frequencies of the packages in the SFT repo.
-To generate simulated container requests,
-we made random package selections subject to the observed distribution of package use.
-Many packages in the repo were not observed being used at all,
-so we added a small, non-zero probability of selecting packages lacking usage data.
-We applied this technique in combination with the package dependency graph,
-so that images contained a weighted random core selection along with any dependent packages.
-This second image simulation scheme takes into account the dependency structure of the repo as well as usage patterns observed in the wild.
-
-As a base for comparison,
-we also generated simulated images consisting of packages chosen in a uniform random way.
-To ensure that total size (or at least total number of packages)
-is comparable to images generated by the previous two methods,
-for this approach we started with an image request generated via the first scheme discussed
-(uniform random core selection with dependencies added).
-We considered only the number of software packages in the resulting image,
-and then chose the same number of packages uniformly randomly from the entire repo,
-ignoring usage information and package dependencies.
-while images generated in this way are not particularly realistic,
-they will be useful for evaluating image management schemes.
-By comparing results with random images to those with the previous two image generation schemes,
-we can compare the general case of containers as collections of arbitrary data to the specific focus of this work,
-i.e.\ containers with selections of experimental software.
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/comparative/distribution_efficiency.pdf}
+\label{fig:dist-eff}
+\if 0
+explain what happened
+metrics: why we care about efficiency?
+cache efficiency: gauge duplication (waste), keeping working set costs lots of storage
+container efficiency: unexpected overhead for each request; compute, network, etc.
+place limits based on site, conditions
+limits exist, we won't tell you
+maybe just buy more cache?
+large scale (10-100 TB), no can do...so need to merge sometimes
+why 100 and 20 on the left
+why they switch places: extreme cases
+strike a balance
+one curve might be more important for you
+effects are not uniform/balanced gentle drop, massive spike in overhead on right
+\fi
+\end{figure}
+
+When sweeping over the range of $\alpha$ values,
+there are a large number of metrics available to summarize each simulation run.
+Many, however, are highly coupled with the particular system configuration and difficult to compare as we vary the paramaters of the simulations.
+
+We therefore chose to define two metrics, cache efficiency and container efficiency,
+to indicate the relative utilization of the system.
+Figure~\ref{fig:dist-eff} plots these two metrics for a representative simulation.
+We defined cache efficiency as the ratio of unique data to total data in the cache.
+In our case, this is equivalent to the ratio of the size of the unique packages to the total cache size.
+If many images contain copies of the same packages,
+the cache efficiency decreases.
+This metric captures the extent of duplication within the cache across all images.
+With no merging there is a high degree of duplication,
+so the cache efficiency is low.
+On the other end of the spectrum,
+maintaining a single, large image containing all data results in cache efficiency of 100\%,
+as the entire cache resides in that single image with no duplication
+
+We defined container efficiency as the ratio of the size of the requested container
+(in our case a set of requested packages plus all dependencies)
+to the size of the container the system actually used for the job.
+In the absence of merging,
+these two are equal so the container efficiency is 100\%;
+jobs are run with exactly what was requested.
+By merging to allow for image reuse,
+we include additional, unrequested data in container images.
+The container efficiency measures this difference between requests and containers.
+In the extreme case of $\alpha=1$ with a single large image, for example,
+the container efficiency is poor because the entire repo is used for every request,
+regardless of size.
+
+These two extreme cases, no merging among many images and a single merged image,
+can both be useful in some situations.
+If the requirements of all jobs are known in advance and worker storage and bandwidth are sufficient,
+maintaining a single, large image for all jobs on all workers is an useful strategy.
+After the initial (costly) image creation and transfer,
+there are no continuing costs or delays.
+When resources are limited or requirements change regularly,
+however, this approach becomes prohibitively expensive.
+The high compute and bandwidth overhead apply for every image update,
+which in the worst case could be every job.
+
+The other extreme situation,
+simply caching requests with no merging,
+can also be viable.
+This approach is simpler and does not come with extreme per-job overhead.
+In addition, frequent changes to the jobs and requirements do not degrade the efficiency of each job.
+At large scale, however,
+the overall system efficiency suffers.
+Due to duplication among images,
+larger caches contain less and less unique data.
+Thus to support a given repo,
+it can become necessary to provision a cache much larger than the size of the repo.
+With common CVMFS repos consuming several terabytes of storage each,
+the amount of cache space used can quickly grow.
+
+These two limiting factors,
+the compute and transfer cost in the highly merged case and the cache efficiency in the unmerged case,
+serve as limits on the viable range of $\alpha$ values for the system.
+Figure~\ref{fig:dist-eff} shows two vertical lines indicating these limits on a plot of efficiencies.
+The left line shows a lower limit of around 30\% on the cache efficiency.
+Choosing $\alpha$ too low results in cache efficiency reaching below 20\%.
+On the right, the line gives an upper limit on the likelihood of merging.
+As shown in Figure~\ref{fig:opcounts},
+the amount of I/O and compute to update images becomes much larger if $\alpha$ is set too high.
+These two limits define a range of viable values for $\alpha$.
+There is no general rule for the placement of these limits,
+wh depends strongly on the performance characteristics of the execution environment,
+as well as the priorities of the administrators in optimizing the system.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/comparative/cache_efficiency.pdf}
+justification for cache efficiency
+unique vs total
+full x range?
+merging can give more unique data in less cache space
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/comparative/container_efficiency.pdf}
+\if 0
+related to container efficiency: not exact but shows it
+too much jitter from individual requests, so cumulative
+crossover: hit rate allows reuse, so slightly less bytes written than requested
+requested is based on random stream, constant
+actual written goes up fast with alpha, so more overhead
+proxy for compute/network/etc.
+really steep at high alpha
+\fi
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/comparative/distribution_efficiency_comp.pdf}
+\if 0
+different data patterns cause different effects
+random has high utilization
+pushed right, limited opportunity for merge, since no overlap
+merging doesn't help
+random is not realistic though
+merging is not general solution, but for trees of software, look at this
+
+tree has more duplication, starts merging much sooner
+closure is very non-controvertial, but results in serious changes in efficiency
+common to any software, so opportunity for optimization
+common core that can merge to avoid dups
+including usage stats (dist) amplifies these somewhat
+real usage should be stronger than dist
+
+use dist from now on because reasonable approximation to actual usage
+similar to tree anyway
+mild but noticable difference from tree, more natural approximation to real usage
+\fi
+\end{figure}
 
@@
 
+\begin{figure*}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_cache_percent_plt.pdf}
+\hfill
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_cache_percent_plt.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_request_percent_plt.pdf}
+\hfill
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_request_percent_plt.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_uses_percent_plt.pdf}
+\hfill
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_uses_percent_plt.pdf}
+\if 0
+[t] "just buy more cache"
+efficiency is worse, since cache holds on to lots of questionable data
+loads of opportunities to try merges, so we see a lot of unnecessary merges
+less force removing bad choices, so more cruft can accumulate in many images
+better cache utilization with tiny cache, but of course more thrashing
+at scale, bad idea
+large cache is too pricey at scale
+[m] varying reuse doesn't affect too much
+below 100 doesn't seem full/steady state yet
+cache isn't settled? so haven't dropped bad merges; more unique data in cache boosts efficiency
+[b] high number of requests, repeats don't matter
+unique data dominates
+hard to tell difference between [m] and [b], so not too sensitive to these
+once you hit steady state, jobs count/repeats has little effect on efficiency
+general trends of merging, not interesting by themselves
+\fi
+\end{figure*}
+
 Using these metrics for cost and overhead,
@@
 
-\begin{figure*}
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_cache_percent_plt.pdf}
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_cache_percent_plt.pdf}
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_request_percent_plt.pdf}
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_request_percent_plt.pdf}
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_uses_percent_plt.pdf}
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_uses_percent_plt.pdf}
-\caption{}
-\end{figure*}
-
 \section{Reproducibility Data}

### 2019-01-23 paper.tex
Updated figure layout

@@
 Nearly all research software used at CERN assumes the availability of the CernVM File System (CVMFS),
-a distributed global filesystem built for distributing software and configuration data.
+a global filesystem built for distributing software and configuration data.
 CMVFS provides a read-only, distributed file system using HTTP for network transport and extensive local
@@
 on the other hand, these requirements present a serious barrier.
+Solutions exist such as
 Parrot, a debugging tool and personal filesystem developed at the University of Notre Dame,
-can be used in cases where privilege on the worker nodes is limited and FUSE is not available but still requires unrestricted network access.
-For computing at scale, it is important to provide a caching proxy as well,
-which is likely to require assistance from administrators.
+which sidesteps FUSE but still requires unrestricted network access and
+UnCVMFS which locally store all of CVMFS to avoid network access, 
+but relies on large storage and permission mounting.
+For computing at scale, it is important to provide a caching proxy and concise storage,
+which currently require assistance from administrators.
 
@@
 HPC resources are an appealing source of computing power to supplement the WLCG.
-Leadership class machines such as Summit (?) will by themselves be able to perform a substantial portion of the work of the entire WLCG today.
-Use of national-scale HPC resources across the United States and Europe such as XXX, YYY, Piz Daint at the Swiss CSCS HPC center could be key in meeting the computational demands following the LHC's high-luminosity upgrade.
+Leadership class machines such as Summit, at Oak Ridge National Laboratory, will by themselves be able to perform a substantial portion of the work of the entire WLCG today.
+Use of national-scale HPC resources across the United States and Europe such as Cori at NERSC, Sierra at Lawrence Livermore National Laboratory, Piz Diant at Swiss National Supercomputing Centre could be key in meeting the computational demands following the LHC's high-luminosity upgrade.
 
@@
 
+\if 0
 % https://indico.cern.ch/event/587955/contributions/2937411/attachments/1683156/2705423/WahidBhimji-CHEP18-NERSC.pdf
@@
 DVMFS is nonetheless a viable solution as long as the costs of developing/maintaining the service and allocating infrastructure do not grow too high.
+\fi
 
@@
 specifications.
+
+\subsubsection{Tracing}
 The first method is to create a sample trace for your
@@
 or building the tracing in to the application.
-As we will see later, we chose application
-level tracing as it provides more context
-for the use.
-
+
+Tracing is attractive as it provides a consistent 
+behavior between experiments and use cases.
+In addition, it is easier to automate and sample arbitrary applications without assistance from each user.
+This allows the solution to port between applications
+quickly without having to delve in to the applications
+specific configuration.
+As previously mentioned, there are several methods for
+obtaining a trace from a file.
+
+Despite assumption that both CVMFS and FUSE will be limited
+on the HPC site, this methods was used as it is reasonable to
+allow for sampling of these jobs on WLCG. 
+WLCG provides the flexibility of using CVMFS and FUSE, without
+having to implement or provision a separate tracing infrastructure
+at each site.
+Additionally, as many of these jobs would have originally been 
+destine to run on WLCG, the sampling sizes are unlikely to 
+create significant overhead.
+Using CVMFS for the tracing allows CVMFS's just-in-time deployment
+of files to patch the environment and give an accurate view without
+the risk of missing files.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/cdf/cdf.pdf}
+\caption{The distribution of file size for each repository.
+From this we can see that the large majority of files
+accumulate in the kilobyte to megabyte range.}
+\label{fig:cdf}
+\if 0
+lots of small files
+table shows duplication
+long tail of large files
+different scales, similar usage across repos
+
+\fi
+\end{figure}
+
+\subsubsection{Packages}
 The second automated method is to use static
@@
 to determine the configuration.
+In the SFT CVMFS repository, where many core research packages are located,
+dependencies are mapped for many packages using buildinfo files.
+Chaining these buildinfo files together allow for a rough dependency
+tree to be create, which was used to sample for \Cref{fig:picks}.
 With either of these methods the issues
@@
 
-For the analysis and creation pipeline discussed
-later in the paper, tracing was used to provide
-the file specifications. 
-Tracings flexibility between repositories
-and more precises file inclusion allow
-for a lean package that provides coverage for
-an experiment.
-
-\subsection{Trace Sampling}
-Tracing is attractive as it provides a consistent 
-behavior between experiments and use cases.
-In addition, it is easier to automate and sample arbitrary applications without assistance from each user.
-This allows the solution to port between applications
-quickly without having to delve in to the applications
-specific configuration.
-As previously mentioned, there are several methods for
-obtaining a trace from a file.
-
-Tracing was implemented in the CVMFS software and
-interacts via the FUSE interface.
-The tracing utility has been implemented in a way that 
-minimizes the overhead if the tracer is deactivated and 
-also minimizes the measurable overhead in the FUSE calls for an activated tracer 
-by delegating the actual log writing to a separate thread to avoid blocks 
-during the accesses to CVMFS. 
-In fact, for a deactivated tracer the entire overhead consists of 
-a single inline if for a subset of the FUSE calls.
-This allows the actual observed CVMFS usage to  be tracked
-and then used for determining the file specification.
-
-Despite assumption that both CVMFS and FUSE will be limited
-on the HPC site, this methods was used as it is reasonable to
-allow for sampling of these jobs on WLCG. 
-WLCG provides the flexibility of using CVMFS and FUSE, without
-having to implement or provision a separate tracing infrastructure
-at each site.
-Additionally, as many of these jobs would have originally been 
-destine to run on WLCG, the sampling sizes are unlikely to 
-create significant overhead.
-Using CVMFS for the tracing allows CVMFS's just-in-time deployment
-of files to patch the environment and give an accurate view without
-the risk of missing files.
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/closure/closure.pdf}
+\caption{Difference in number of packages and size when accounting for dependency based closure in selection.}
+\label{fig:picks}
+\if 0
+you get more than you ask for
+smaller requirements have relatively large responses
+non-linear, so bigger ones have more duplicates
+appraoches repo size in limit
+merging is subject to drop-off(more duplication)
+\fi
+\end{figure}
 
@@
 and providing this information for verification.
-
-The use of a global revision number provides a quick reference
-on the state of the repository at the highest level.
-Global revision numbers can instantly verify the expected
-contents, but does not give evidence of compatibility between
-revisions.
-As the revision number is global, any change to the repository
-is reflected in a change to the revision number, which means 
-some software is unchanged between revisions.
-This method leaves much to be desired when using common stable
-packages that change irregularly.
-
-Aside from computing and checking the contents of the
-repository or its files, it is possible to verify the
-software using experiment computation.
-This is a common method used to testing software
-and works equally when the target software provides
-coverage of the features used by the application.
-The difficulty here is in providing tests that
-behave consistently, verify if computation is
-incorrect, and cover the desired features.
-Though it is possible for the submitting user
-to provide test with coverage, application
-developers have a better grasp of it usage
-and how to provide coverage in testing.
-An additional method is to run the application
-on a determined site for initial results and
-compare in the future for verification, but
-this relies on deterministic code which is 
-not always the assumption in HEP computing.
 
@@
 \label{tab:repo-sizes}
-This table shows the size, number of inodes, and the amount
+This table shows the size, number of inodes (rounded to nearest thousand), and the amount
 of duplicate data that is eliminated using CVMFS.
@@
 \end{table*}
-
-\if 0
-I AM NOT SURE WHICH WE WILL USE. WRAP THIS UP LATER.
-\fi
 
@@
 
-These methods were compared to showcase Shrinkwraps
-behavior and evaluate usage performance.
-The first analysis compares all three methods pulling 
-an entire repository, to allow comparision between the
-three.
-GRAPH OF DEDUPLICATED DATA/FULL REPO SIZE/TIME TO CREATE.
-The second looks at the time to update existing repositories 
-between verison.
-GRAPH UPDATE TIME ON FULL REPO AND MAYBE SUB REPO(SUB REPO ONLY COMPARES RSYNC AND SHRINKWRAP)
-The third looks at hosting several projection at once and
-switching between them.
-COMPARES RSYNC AND SHRINKWRAP. UNCVMFS CANT COMPETE.
-
-\begin{figure}[h]
-\caption{COMPARISON OF SHRINKWRAP unCMVFS and RSYNC}
-\label{figure:projection-comp}
-\end{figure}
-
 \subsection{Packaging Projections}
 
-All of the above methods describe utilize CVMFS
+All of the above methods described utilize CVMFS
 and rely on the internet connection to create
 the projections.
-A core assumption of this paper is that internet
+A core assumption is that external network
 access is limited at the compute nodes, so a method
@@
 for computation.
-For this we looked at several methods for providing
-these projections and compared using 
+For this several methods for providing
+these projections using 
 Container images (images specific to Docker, Singularity, ect.)
-versus SquashFS images.
+or SquashFS images can be used.
 The goal is to find a flexible, compatible intermediate
@@
 For this purpose, Docker images, Singularity images, and 
-SqaushFS are compared on their time to create,
-storage requirements, and most importantly compatibility
-with several used technologies, namely 
-Docker, Singularity, and Shifter.
+SqaushFS are interchangeable as outputs of the Shrinkwrap process
+and changed as needed for each use case as seen below.
+
+%are compared on their time to create,
+%storage requirements, and most importantly compatibility
+%with several used technologies, namely 
+%Docker, Singularity, and Shifter.
 
@@
 
-When creating these images for analysis jobs its often the case that several
-of these images may be created and used in the system.
-Using Shrinkwrap to organize the repository data and projections,
-the impact of an increasing number of images can be see in 
-\Cref{figure:spec-size}.
-This graph illustrates a new challenge that needs to be addressed.
-As the number of experiments run and projections hosted by Shrinkwrap
-increase. 
-The raw number of images that need to be host increases to an unmanageable
-size.
-A solution is needed for how to manage these images and still
-provide responsive execution of HEP analysis on HPC.
-
-\subsection{Use Cases of Shrinkwrap}
-%Here we illustrate several real world use cases.
-%These use cases are being used as the existing solutions
-%were not feasible and a new approach was needed.
-%In analyzing these use cases we can see the beginnings of
-%the problem outlined in the previous subsection.
-
 \subsubsection{Atlas at NERSC}
 
-We are currently discussing options to integrate our shrinkwrap utility into the HPC deployment processes of ATLAS as we believe that the proposed software solution can help to reduce the overhead of HPC software delivery significantly.
-On the one hand the actual uncvmfs export process can be optimized as the current export workflow loads the entire CVMFS repository in a first step and later on removes certain files in the exported image which were not in fact needed.
-This is something that can be entirely avoided through the definition of exclusion rules passed to the shrinkwrap utility.
-
-On the other hand, we believe that an export client which relies on the libcvmfs library might reduce the total overhead and therefore improve the export performance in comparison to both uncvmfs and rsync (for uncvmfs this is something that we haven’t had yet the chance to evaluate in an experiment).
-As a longer-term goal, it might be useful to start tracing ATLAS simulation workflows on a few machines to collect data that might allow further export optimizations in a later stage through automated specification building.
+As previously described, NERSC is working in collaboration the LHC ATLAS project
+to provide HPC resources for simulation and experimentation.
+NERSC is currently deploying CVMFS images as monolithic container 
+built using UnCVMFS, packaged in Squashfs images, and then containerized 
+into Shifter or Singularity format, whose final format
+relies on the target site.
+This entire process can take up to a day to complete with the majority
+of compute used to pull down and store the full CVMFS copy and storage
+needs in excess of terabytes for images.
+By combining either traced samples or statically defined requirements for 
+jobs, Shrinkwrap can reduce these images orders of magnitudes.
+Shrinkwrap which relies on the libcvmfs library reduces 
+the total overhead and eliminates unnecessary files transfers
+over UnCVMFS by allowing selections and rsync using data deduplication.
 
@@
 
-\begin{figure}
-\includegraphics[width=\linewidth]{curated/cdf/cdf.pdf}
-\label{fig:cdf}
-\if 0
-lots of small files
-table shows duplication
-long tail of large files
-different scales, similar usage across repos
-
-\fi
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{curated/closure/closure.pdf}
-\label{fig:picks}
-\if 0
-you get more than you ask for
-smaller requirements have relatively large responses
-non-linear, so bigger ones have more duplicates
-appraoches repo size in limit
-merging is subject to drop-off(more duplication)
-\fi
-\end{figure}
+Since there are currently few regular CVMFS on HPC users,
+we decided to simulate the activity of users requesting container images.
+We decided to simulate at the granularity of packages rather than individual files,
+because this results in more robust container images.
+As mentioned previously,
+fine-grained tracing runs the risk of leaving out important dependencies.
+In addition, working only based on the properties of individual files ignores the structure inherent in software collections.
+Looking only at the properties of individual files,
+Figure~\ref{fig:cdf} shows cumulative distributions for file sizes in several of the largest and most active CVMFS repos.
+The repos vary in scale,
+but all show similar usage patterns.
+We see that across all repos,
+the majority of files are small (between 1~KB and 1~MB).
+The repos do contain, however,
+a small number of large files.
+This fits our expectation that the CVMFS repos primarily contain software.
+Most individual files are small,
+but the repo size is defined by the large number of packages in the software collection.
+
+We tested two schemes for generating image requests taking into account the properties of the software collections,
+as well as recorded user activity patterns.
+We also generated images via a uniform random scheme for comparison.
+We consulted with the developers of CVMFS,
+as well as HEP researchers at the University of Notre Dame collaborating with CERN to determine how current users interact with CVMFS.
+Since CVMFS is not currently widely deployed on HPC resources,
+the prospective user base for our container management system is limited.
+We were, however, informed of some researchers and groups experimenting with static mirrors of CVMFS for container images,
+including projects at NERSC and Stanford University,
+with one of the setups available publicly at \url{https://github.com/wyang007/atlas-fat-container}.
 
@@
 \fi
-
-Since there are currently few regular CVMFS on HPC users,
-we decided to simulate the activity of users requesting container images.
-We decided to simulate at the granularity of packages rather than individual files,
-because this results in more robust container images.
-As mentioned previously,
-fine-grained tracing runs the risk of leaving out important dependencies.
-In addition, working only based on the properties of individual files ignores the structure inherent in software collections.
-Looking only at the properties of individual files,
-Figure~\ref{fig:cdf} shows cumulative distributions for file sizes in several of the largest and most active CVMFS repos.
-The repos vary in scale,
-but all show similar usage patterns.
-We see that across all repos,
-the majority of files are small (between 1~KB and 1~MB).
-The repos do contain, however,
-a small number of large files.
-This fits our expectation that the CVMFS repos primarily contain software.
-Most individual files are small,
-but the repo size is defined by the large number of packages in the software collection.
-
-We tested two schemes for generating image requests taking into account the properties of the software collections,
-as well as recorded user activity patterns.
-We also generated images via a uniform random scheme for comparison.
-We consulted with the developers of CVMFS,
-as well as HEP researchers at the University of Notre Dame collaborating with CERN to determine how current users interact with CVMFS.
-Since CVMFS is not currently widely deployed on HPC resources,
-the prospective user base for our container management system is limited.
-We were, however, informed of some researchers and groups experimenting with static mirrors of CVMFS for container images,
-including projects at NERSC and Stanford University,
-with one of the setups available publicly at \url{https://github.com/wyang007/atlas-fat-container}.
 
@@
 
+To decide if two projections are ``close'',
+we define the parameter $\alpha$ as the maximal Jaccard distance between closely related projections.
+Since Jaccard distance is by definition between zero and one,
+$\alpha$ must be in the same range.
+This $\alpha$ parameter is something like the ``globbiness'' of the system.
+Small choices of $\alpha$ requires that projections be extremely close before considering them for merging.
+In the extreme case with $\alpha = 0$,
+only \emph{identical} images will be considered close so no images will be merged.
+This results in a larger number of independent images.
+Choosing $\alpha$ to be larger makes it more likely for images to be considered similar and merged.
+This results in more augmented images that serve multiple tasks.
+In the extreme case of $\alpha = 1$,
+\emph{every} pair of images is considered close and merged if possible.
+This results in large container images that have accumulated many projections.
+We explore tradeoffs when choosing $\alpha$ and give recommendations based on observed patterns of activity in section ??.
+
+Having selected a set of candidate projections that are close enough to the new request,
+we next check each in turn to find either a superset of the requested projection or one without version incompatibilities.
+We first sort the list of candidate images by Jaccard distance.
+This ensures that our system prioritizes candidates that are similar in size to the requested image.
+In the case of finding a superset,
+we can simply use the corresponding cached container image.
+Presumably this superset was created by the merging process below,
+so we are done.
+Otherwise, we continue to look for a compatible projection
+As discussed in section ??,
+two projections are incompatible if they both include the same path but point to different data.
+This situation can arise if for example a library in CVMFS is updated in place.
+Old and new projections would include an identical filesystem path,
+but the new version would have different contents.
+This situation can arise in more subtle ways,
+for example with symlinks controlled by environment variables.
+Since projections of CVMFS must statically resolve such links,
+incompatibilities can arise despite paths appearing compatible under CVMFS.
+Any candidate projections that are incompatible with the new request are removed from further consideration.
+
+If no compatible projections were found among the candidates,
+we create a container image for the projection as requested and add it to the cache.
+Otherwise, on finding a compatible cached projection the next step is to augment it with the contents of the new request.
+We choose the first compatible image from the list of candidates for merging,
+again prioritizing images that are similar in size and contents.
+As previously discussed,
+Shrinkwrap can update a projection that was previously created;
+since we keep the projection used to generate each container image in the cache,
+we can simply augment the selected projection with the new files from the request.
+Next, we build a container image based on the newly updated projection.
+The computational cost of this step could become high since we are creating larger combined projections.
+Appropriate choice of $\alpha$, however,
+bounds the distance between the combined projections such that the cost of creating the augmented container is arbitrarily close to cost of simply fulfilling the original request.
+Since the augmented projection is the union of two different projections,
+we can substitute it for both.
+The container image for the augmented projection can replace the previous image,
+so that the net increase in storage space comes only from the newly added files.
+Since a single augmented image now serves two different batch job submissions,
+we avoid making two copies of the common files and data.
+
+One potential issue with this merging strategy is that there is no way to prevent poor merging decisions.
+In the case of a one-off submission,
+the system might merge an existing, frequently used image.
+This is not problematic for the one-off job itself,
+but increases overhead for subsequent jobs.
+The original, frequently used image would increase in size with no additional benefit to other tasks.
+Images might become ``bloated'' in this way,
+accumulating infrequently used dependencies and increasing overhead indefinitely for future tasks.
+The Jaccard distance metric gives a natural way to capture and quantify this effect.
+As an image becomes bloated due to repeated merges,
+its distance from every individual request increases.
+After sufficient growth,
+the image will become too far from any request to even be considered.
+Without regular use,
+the bloated image will eventually be removed in LRU order.
+Choice of $\alpha$ allows an upper limit on the amount of useless bloat in images.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/time-series/time-series.pdf}
+\label{fig:series}
+\if 0
+zoomed out?
+cache limit, deletes start
+merge tracks writes, it's the dominant factor
+hits grow gradually as cache merges and fills
+deletes help prune useless stuff, remaining should be frequent hits
+far ends are either insert or merge, middle alphas have both
+\fi
+\end{figure}
+
+Figure~\ref{fig:series} shows a single simulation of the image management system,
+with $\alpha=0.0.75$, cache size XXX, ...
+First, we note that most of the operations are merges.
+This makes sense, as this simulation was at fairly high $\alpha$.
+The total bytes written also closely tracks merges,
+indicating that merging is the dominant source of I/O.
+We can still see inserts over the course of the simulation.
+At more extreme $\alpha$ values,
+we would expect to see one of these operations almost exclusively.
+As the data in cache continues to rise,
+we eventually see the cache limit,
+after which the delete count starts increasing.
+Over the course of the simulation,
+inserts and deletes are filling and emptying the cache such that it remains close to its storage limit.
+We also observe the number of cache hits continue to rise despite deletions.
+As we will see, merging allows for a greater proportion of hits even if the amount of data remains constant.
+This is because frequently used data can be merged,
+reducing duplication.
+The cache limit then ensures that infrequently used parts are eventually removed.
+
 \begin{figure}
@@
 
-% explain alpha plot
+\section{Evaluation}
+
+To evaluate the viability of our approach in automatically deduplicating a container image store,
+we simulated the behavior of the system over time for a range of $\alpha$ values.
+Choice of $\alpha$ is important here to ensure that common components can be detected and merged in container images,
+and that old images or poor decisions are eventually removed from the system.
+Choosing $\alpha$ too far in either direction will result in excessive storage overhead due to duplicated components,
+or excessive compute and network overhead repeatedly merging largely unrelated images.
+Our goal in this evaluation is therefore to choose $\alpha$ so as to minimize the storage and compute costs associated with maintaining a collection of images.
+
+Since our simulation uses random simulated requests,
+there is noticable variability between individual simulations.
+Thus for a given choice of cache size, job count, etc.
+we repeated the simulation 20 times and reported the median behavior over the runs.
+At each choice of $\alpha$ (in steps of 0.05)
+we performed a set of 20 simulated runs,
+allowing us to plot various measurements of the system versus $\alpha$.
 
@@
 
-To decide if two projections are ``close'',
-we define the parameter $\alpha$ as the maximal Jaccard distance between closely related projections.
-Since Jaccard distance is by definition between zero and one,
-$\alpha$ must be in the same range.
-This $\alpha$ parameter is something like the ``globbiness'' of the system.
-Small choices of $\alpha$ requires that projections be extremely close before considering them for merging.
-In the extreme case with $\alpha = 0$,
-only \emph{identical} images will be considered close so no images will be merged.
-This results in a larger number of independent images.
-Choosing $\alpha$ to be larger makes it more likely for images to be considered similar and merged.
-This results in more augmented images that serve multiple tasks.
-In the extreme case of $\alpha = 1$,
-\emph{every} pair of images is considered close and merged if possible.
-This results in large container images that have accumulated many projections.
-We explore tradeoffs when choosing $\alpha$ and give recommendations based on observed patterns of activity in section ??.
-
-Having selected a set of candidate projections that are close enough to the new request,
-we next check each in turn to find either a superset of the requested projection or one without version incompatibilities.
-We first sort the list of candidate images by Jaccard distance.
-This ensures that our system prioritizes candidates that are similar in size to the requested image.
-In the case of finding a superset,
-we can simply use the corresponding cached container image.
-Presumably this superset was created by the merging process below,
-so we are done.
-Otherwise, we continue to look for a compatible projection
-As discussed in section ??,
-two projections are incompatible if they both include the same path but point to different data.
-This situation can arise if for example a library in CVMFS is updated in place.
-Old and new projections would include an identical filesystem path,
-but the new version would have different contents.
-This situation can arise in more subtle ways,
-for example with symlinks controlled by environment variables.
-Since projections of CVMFS must statically resolve such links,
-incompatibilities can arise despite paths appearing compatible under CVMFS.
-Any candidate projections that are incompatible with the new request are removed from further consideration.
-
-If no compatible projections were found among the candidates,
-we create a container image for the projection as requested and add it to the cache.
-Otherwise, on finding a compatible cached projection the next step is to augment it with the contents of the new request.
-We choose the first compatible image from the list of candidates for merging,
-again prioritizing images that are similar in size and contents.
-As previously discussed,
-Shrinkwrap can update a projection that was previously created;
-since we keep the projection used to generate each container image in the cache,
-we can simply augment the selected projection with the new files from the request.
-Next, we build a container image based on the newly updated projection.
-The computational cost of this step could become high since we are creating larger combined projections.
-Appropriate choice of $\alpha$, however,
-bounds the distance between the combined projections such that the cost of creating the augmented container is arbitrarily close to cost of simply fulfilling the original request.
-Since the augmented projection is the union of two different projections,
-we can substitute it for both.
-The container image for the augmented projection can replace the previous image,
-so that the net increase in storage space comes only from the newly added files.
-Since a single augmented image now serves two different batch job submissions,
-we avoid making two copies of the common files and data.
-
-One potential issue with this merging strategy is that there is no way to prevent poor merging decisions.
-In the case of a one-off submission,
-the system might merge an existing, frequently used image.
-This is not problematic for the one-off job itself,
-but increases overhead for subsequent jobs.
-The original, frequently used image would increase in size with no additional benefit to other tasks.
-Images might become ``bloated'' in this way,
-accumulating infrequently used dependencies and increasing overhead indefinitely for future tasks.
-The Jaccard distance metric gives a natural way to capture and quantify this effect.
-As an image becomes bloated due to repeated merges,
-its distance from every individual request increases.
-After sufficient growth,
-the image will become too far from any request to even be considered.
-Without regular use,
-the bloated image will eventually be removed in LRU order.
-Choice of $\alpha$ allows an upper limit on the amount of useless bloat in images.
-
-\section{Evaluation}
-
-To evaluate the viability of our approach in automatically deduplicating a container image store,
-we simulated the behavior of the system over time for a range of $\alpha$ values.
-Choice of $\alpha$ is important here to ensure that common components can be detected and merged in container images,
-and that old images or poor decisions are eventually removed from the system.
-Choosing $\alpha$ too far in either direction will result in excessive storage overhead due to duplicated components,
-or excessive compute and network overhead repeatedly merging largely unrelated images.
-Our goal in this evaluation is therefore to choose $\alpha$ so as to minimize the storage and compute costs associated with maintaining a collection of images.
-
 \begin{figure}
@@
 
+\subsection{Determining a Viable Range for $\alpha$}
+
 When sweeping over the range of $\alpha$ values,
@@
 There is no general rule for the placement of these limits,
-wh depends strongly on the performance characteristics of the execution environment,
+which depends strongly on the performance characteristics of the execution environment,
 as well as the priorities of the administrators in optimizing the system.
+Rather than defining where these limits fall,
+we give per-site considerations in choosing limits in Section ??.
 
@@
 \includegraphics[width=\linewidth]{curated/comparative/cache_efficiency.pdf}
+\label{fig:cache-eff}
 justification for cache efficiency
@@
 
+Figure~\ref{fig:cache-eff} shows the actual data sizes used to calculate the cache efficiency.
+Without merging, the unique data makes up a small proportion of the cache.
+With increased merging, the amount of unique data increases.
+For sufficiently high $\alpha$,
+merges occur more frequently than cache inserts,
+resulting a drop in total storage size.
+On the far right of Figure~\ref{fig:cache-eff},
+the entire cache contents are merged into a single large image,
+so that the unique and total data sizes become equal.
+The cache efficiency metric is affected both by the increase in unique data and by the decrease in total cache size.
+
 \begin{figure}
 \includegraphics[width=\linewidth]{curated/comparative/container_efficiency.pdf}
+\label{fig:container-eff}
 \if 0
@@
 
+Figure~\ref{fig:container-eff} shows the cumulative costs of creating and updating images relative to the cost of simply providing the requested images.
+These costs illustrate the effect of container efficiency over the course of a simulated run.
+Without merging, the actual I/O and compute costs in the system closely follow the requests.
+In Figure~\ref{fig:container-eff} the actual I/O is slightly smaller than the requested amount due to caching:
+the system can sometimes reuse images without performing any I/O.
+As $\alpha$ increases, the cost of updating and merging images comes to dominate the total cost.
+Administrators migh decide, for example,
+to spend at most twice the compute and I/O time relative to the the actual requested operations.
+This would correspond to the upper compute limit shown on \label{fig:dist-eff}.
+Depending on the priorities at a particular execution site,
+either the storage utilization or compute and I/O time could be the critical factor in optimizing performance.
+
 \begin{figure}
 \includegraphics[width=\linewidth]{curated/comparative/distribution_efficiency_comp.pdf}
+\label{fig:comp-eff}
 \if 0
@@
 
+We also compared the efficiencies of several types simulated images under a range of $\alpha$ values.
+For the previous discussion, synthetic images were generated taking into account both the tree of package dependencies and usage statistics.
+Figure~\ref{fig:comp-eff} shows a representative simulation with two other synthetic image types included.
+First, we measured images considering only the dependency relationships but not usage.
+These images show similar behavior to those previously discussed,
+albeit with slightly lower efficiencies.
+More realistic images primarily make the same trends more pronounced.
+This suggests that even with random, non-realistic requests,
+the tree structure of package dependencies is sufficient to produce duplication in the cache.
+The other type of synthetic image shown in Figure~\ref{fig:comp-eff} is purely randomly generated.
+The set of packages in each image was chosen uniformly randomly from the full SFT repo.
+These random images do not take into account usage statistics or package dependencies,
+making the software in them entirely nonfunctional.
+For the random images,
+we observe markedly different cache behavior.
+The cache efficiency is much higher,
+indicating a greater proportion of unique data in the cache.
+Since the image contents were selected purely at random,
+this is expected.
+The container efficiency is also high over a much larger range of $\alpha$ values.
+This is because there is effectively no merging until extremely high $\alpha$.
+Since image contents are purely random,
+there is no correlation between different images.
+Thus, it is much more difficult to find images similar enough to merge until the $\alpha$ value is very lax.
+This would indicate that our merging strategy is not applicable to arbitrary collections of data.
+Random images show little to no effect for most $\alpha$ values.
+Our merging strategy, which takes advantage of duplicated content included as a result of dependencies in software,
+would be ill advised for situations that are not known to follow similar patterns of duplication.
+
+\if 0
 \subsection{Quantifying Image Management Costs}
@@
 or equivalently allows the same set of job requests to be served efficiently from a smaller cache.
-
-\subsection{Effect of $\alpha$ Threshold}
+\fi
+
+\subsection{System-specific Effects of $\alpha$ Threshold}
 
@@
 \includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_uses_percent_plt.pdf}
+\label{fig:sensitivity}
 \if 0
@@
 
+In Figure~\ref{fig:sensitivity}, we plot efficiency curves for a range of simulation conditions.
+The left column shows container efficiency,
+while the right column shows cache efficiency.
+In the first row, the number of jobs and the amount of repetition are constant while the cache size is varied.
+In the second row, the number of unique requests is varied while holding the other parameters constant,
+and the third row varies the number of times each job is submitted while holding the rest constant.
+
+The size of the cache has an inverse relationship with both the container and cache efficiency.
+A larger cache can of course hold a larger number of images,
+but since each image contains significant duplicated portions,
+increasing cache size tends to decrease cache efficiency.
+Conversely, small caches more quickly evict images so that ineffective merges and similar images tend not to remain in cache too long.
+A larger cache also allows for more opportunities to merge images,
+leading to decreased container efficiency.
+When deciding how to handle a request,
+a large cache full of images is much more likely to contain an image suitable for merging.
+With a small cache, opportunities to merge are much more dependent on the order in which requests happen to come.
+
+The effect of varying the number of unique jobs is less pronounced than the effect of cache size.
+Streams of 500 and 1000 unique jobs show nearly indistinguishable behavior,
+indicating that by 500 jobs the system has reached a steady state.
+Continuing with an arbitrarily long stream should not result in significant performance changes.
+It appears that 100 unique jobs were not sufficient to fill the cache and reach a steady state.
+In this case the container efficiency is somewhat decreased over a range of $\alpha$ values,
+suggesting that some ineffective merges had not made their way out of the cache yet.
+The cache efficiency in this case is slightly increased.
+This would suggest that before reaching a steady state,
+the contents of the cache are more assorted and some unnecessary data remain in cache.
+
+The effect of varying the number of times each unique job is submitted appear minimal at best.
+As discussed with the previous row,
+a stream of 500 unique images should be sufficient to reach a steady state.
+For a sufficiently large cache, therefore,
+it appears that the efficiency is determined by the amount of unique data and fairly insensitive to how much those pieces of data are used.
+This simulation is somewhat unrealistic in that each job is repeated a uniform number of times.
+In real systems, we expect more variation in the frequency of repeated jobs.
+We predict that a steady, non-uniform distribution of frequencies would result in the same behavior as shown in the bottom row of Figure~\ref{fig:sensitivity},
+while changes in the frequencies of jobs (e.g. a new user starting an experiment)
+would disturb the steady state and produce effects similar to those in the middle row.
+
+\if 0
 Using these metrics for cost and overhead,
@@
 The high cost of creation means that such large images can only be infrequently created or updated.
+\fi
+
+\subsection{Steps to Determining $\alpha$}
+TODO
 
 \section{Related Work}
+TODO
 
 \section{Conclusion}
+TODO
 

### 2019-01-23 paper.tex
Update problem section and graph sizes

@@
 \begin{document}
-\title[Projecting Globel Filesystems into Local Containers.]{Projecting Global Filesystems into Local Containers for High-throughput Computing on HPC Resources}
+\title[Projecting Global Filesystems into Local Containers.]{Projecting Global Filesystems into Local Containers for High-throughput Computing on HPC Resources}
 
@@
 Scientific applications often depend on large, complex software repositories containing libraries, tools, and frameworks in multiple versions and platforms to support many researchers.
-As a practical necessity, software repos are most commonly distributed via shared or global filesystems.
-In moving existing research to HPC environments, however,
-scientists must work around restrictions that can interfere with global software delivery methods.
-To illustrate these challenges,
-we examine CVMFS, the CernVM File System.
-CVMFS is a global, distributed filesystem that provides read-only access to experiment software and configurations for high energy physics research at CERN.
-CVMFS requires a number of network and OS configurations that are incompatible with a restrictive HPC environment.
-In this work, we introduce a set of tools and procedures for \emph{projecting} the global filesystem (or parts thereof) into local container images suitable for HPC sites.
-As part of our evaluation,
-we found that using local approximations of a global filesystem tends to introduce problems of its own in terms of compute and storage overhead.
-We thus examined several declarative and profiling-based approaches to creating more fine-grained projections to fit application requirements,
-and propose a strategy for efficiently managing these projected container images at an HPC site.
-Using our projection tools and image management strategy,
-we implemented a proof-of-concept submission system for running existing HEP analysis tasks in an HPC environment.
+Such repositories are distributed to hundreds of major computing
+facilities around the world using a global filesystem like CVMFS.
+However, not all HPC facilities permit execution nodes to
+mount and access global network resources, instead encouraging
+applications to make use of local filesystems and container
+technologies.  In this work, we present a technique for efficiently.
+\emph{projecting} a global filesystem into local resources.
+This consists of creating minimal local containers that meet
+the needs of specific jobs, balancing between the extremes
+of having many small single-purpose containers and having one
+large all-purpose containers.  We evaluate these techniques
+using data from production instances of the CVMFS global filesystem
+used by a wide variety of scientific collaborations.
 \end{abstract}
@@
 \section{Introduction}
+
+Global scientific experiment collaborations like ATLAS, CMS, LIGO,
+and LSST make use of hundreds of thousands of cores spread across
+many computing facilities around the world in order to achieve the massive scale needed simulate and analyze data from high bandwidth
+scientific instruments.  In order to ensure consistency of software
+running at many different sites, collaborations establish global
+software repositories that publish the tested and trusted versions
+of software to be used for production runs.  These repositories
+can be quite large, consisting of thousands of software packages
+consuming millions of files and terabytes of data.  Software
+such as CVMFS~\cite{globalfs-cise-2015} is used to distribute
+these complex software stacks around the world.
+
+However, not all computing facilities make it easy to access global filesystems at runtime.
+In particular, national HPC facilities that are not associated with any
+particular experiment often prevent network connections between
+individual worker nodes and the outside world, for reasons
+of both security and system efficiency.  Such facilities instead
+encourage the use of local high performance filesystems and container
+technologies to deliver software to individual worker nodes.
+To make use of such sites, some mapping between the global
+filesystem and the locally available technologies must be provided.
+
+In this paper, we describe the problem of {\bf projecting} a global
+filesystem into the resources available at a local HPC facility.
+There are two main components to this problem.
+
+First, for a given job, we must have a means of identifying the
+subset of the global filesystem that is relevant for the current
+execution.  This can be done either by dynamic tracing of the resources
+actually used, or by static analysis of its dependency tree.
+We describe Shrinkwrap, a tool that performs this analysis on
+a Unix application, and produces a minimal container
+that can be used to execute the application, or one closely related.
+We analyze the efficiency of this tool on a selection of high
+energy physics applications, and demonstrate storage efficiency of this technique, and measure the performance and storage efficiency.
+
+Second, for a large number of jobs, we must have a method for
+managing the large number of containers that will be created.
+If one small container is created for every single job,
+site storage will overflow with duplicated data.  On the other
+hand, if one large container is constructed to satisfy all
+possible jobs, it will overflow constraints at the worker nodes.
+To address this, we develop an online algorithm that manages
+a shared storage space with multiple containers, choosing to
+either create, delete, or merge containers as new requests arrive.
+The key parameter ($\alpha$) to this algorithm allows for a
+tradeoff between the efficiency of cache storage and container size.
+
+Both techniques are evaluated in light of the properties
+of large global filesystems used in production by high
+energy physics applications.  In particular, we exploit the
+property that many jobs share a small set of frequently-used
+dependencies, and have a long tail of infrequently-used dependencies.
+We found that our approaches can effectively manage the container images for a continuous stream of submissions with varying software dependencies and versions with less storage overhead than creating a new image per application.
+In addition, our results show that applications can still take advantage of precise dependency and version specifications to reduce image creation and transfer overhead compared to simple mirroring.
+
+\if 0
 
@@
 
-\section{Background and Challenges}
-
-\subsection{CVMFS}
-The CernVM File System (CVMFS) is a globally distributed filesystem designed for providing efficient,
-read-only access to scientific software.
-CVMFS is primarily used for distribution of High Energy Physics (HEP) software used as part of the Large Hadron Collider (LHC) at CERN.
-Researchers at CERN use CVMFS as the primary means of distributing the analysis and simulation software they develop to the Worldwide LHC Computing Grid (WLCG).
-
-Each WLCG site commits to providing some fixed amount of computational resources for analyzing particle collision events and running simulations.
-Worker nodes at each site must be configured to mount CVMFS as a prerequisite to performing any computational tasks.
-Since CVMFS is implemented as a FUSE module,
-WLCG sites are expected to support FUSE on their worker nodes.
-CVMFS also relies on global configuration files,
-requiring administrative involvement in preparing worker nodes for WLCG jobs.
-In addition, sites generally provision some infrastructure to support CVMFS,
-such as caching proxies to reduce load created by local workers on publicly accessible filesystem servers.
-
-As a result, it is difficult to contribute computational resources toward the LHC experiments without making some site-wide provisioning and configuration changes.
-This is not an issue for the WLCG sites,
-as they provide dedicated computing resources and can customize their setups to meet the requirements of CVMFS.
-For a professor attempting to harness a campus cluster,
-on the other hand, these requirements present a serious barrier.
-Solutions exist such as
-Parrot, a debugging tool and personal filesystem developed at the University of Notre Dame,
-which sidesteps FUSE but still requires unrestricted network access and
-UnCVMFS which locally store all of CVMFS to avoid network access, 
-but relies on large storage and permission mounting.
-For computing at scale, it is important to provide a caching proxy and concise storage,
-which currently require assistance from administrators.
-
-\subsection{HPC}
-
-With the rapidly increasing demands of the LHC experiments at CERN,
-HPC resources are an appealing source of computing power to supplement the WLCG.
-Leadership class machines such as Summit, at Oak Ridge National Laboratory, will by themselves be able to perform a substantial portion of the work of the entire WLCG today.
-Use of national-scale HPC resources across the United States and Europe such as Cori at NERSC, Sierra at Lawrence Livermore National Laboratory, Piz Diant at Swiss National Supercomputing Centre could be key in meeting the computational demands following the LHC's high-luminosity upgrade.
-
-Unfortunately, HPC sites are generally more restrictive than the previously mentioned campus resources.
-In addition to limited privilege on workers and lack of FUSE support,
-HPC sites often impose restrictions on network activity,
-including outbound firewalls preventing access to CVMFS.
-
-The US collaboration of the ATLAS project is currently taking advantage of computing resources at various supercomputers in the United States including Cori at NERSC.
-For the deployment of their software the \texttt{uncvmfs} utility is used to unpack entire CVMFS repositories into standalone images (usually in ext4 or squashfs format) which in turn can be used to build Shifter or Singularity images.
-While these images can very well be scaled out onto a large number of nodes inside the NERSC infrastructure,
-the image build process takes around 24 hours which makes it difficult to deploy up-to-date versions of the software on a regular basis.
-As these images contain complete copies of everything in CVMFS,
-each takes on the order of terabytes,
-necessitating careful storage management.
-In addition, the process requires administrators for image creation, deployment, and cleanup.
-Currently, this approach is used in production on the ATLAS and CMS experiments.
-As additional experiments want to take advantage of the resources at NERSC,
-the administrative burden of managing multiple CVMFS images on multiple software versions increases accordingly.
-For the experiments to use the same images (saving time and cost at NERSC),
-multiple groups would have to come to agreement about a single common set of software versions to use and adjust their pipelines elsewhere.
-Both solutions introduce new inconveniences and administrative burden.
-
-More broadly, the use of static, manually managed container images breaks user assumptions about the operation of CVMFS.
-When running jobs using CVMFS,
-researchers can specify the version of each repo to use,
-allowing precise control over software and dependencies.
-Each experiment has its own policies and procedures for choosing a software version;
-some may use bleeding-edge analysis software or another experiment may prefer older, stable versions.
-CVMFS is able to provide multiple customized views of the global filesystem without incurring much additional cost.
-CVMFS also supports some more exotic features,
-such as symlink resolution controlled by environment variables.
-CVMFS as a whole does not map well to the semantics of a traditional filesystem.
-Projecting CVMFS into such a filesystem can capture at most a single, static version of a repo and must necessarily eliminate some features,
-such as environment-based symlink resolution at runtime.
-
-\if 0
-% https://indico.cern.ch/event/587955/contributions/2937411/attachments/1683156/2705423/WahidBhimji-CHEP18-NERSC.pdf
-Another similar approach in use at NERSC is DVMFS.
-Since creating static container images containing CVMFS is costly and difficult to customize,
-DVMFS instead uses Cray DVS IO forwarders to mount a copy of CVMFS served over NFS.
-This allows for less administrator involvement,
-since the mirrored copy of CVMFS can be automatically updated on a regular basis.
-After dealing with several implementation problems,
-DVMFS is currently deployed at NERSC as an alternative to static images.
-While this approach affords more flexibility in choice of repos and software versions,
-it increases the job startup time and places additional requirements on the infrastructure.
-To be usable in production, DVMFS requires a sufficient number of DVS servers to be provisioned.
-DVMFS is nonetheless a viable solution as long as the costs of developing/maintaining the service and allocating infrastructure do not grow too high.
 \fi
+
+\begin{figure}
+\includegraphics[width=\linewidth]{drawings/methods.pdf}
+\label{choices}
+\caption{Global Filesystem Access Methods}
+\end{figure}
+
+\section{Accessing Global Filesystems}
+
+Modern scientific applications depend upon complex software
+repositories that consist of terabytes of data and millions of files.
+Unlike "traditional" high performance applications that are
+single, statically-linked executables, modern applications are
+agglomerations of software that consist of high level scripts,
+multiple executables written in different languages, support
+libraries that may be dynamically linked, and configuration and
+data files loaded at runtime.
+
+In a well-organized project, the necessary software dependencies
+are collected together into a global repository that is accessible
+from all of the sites that intend to use it.  A common arrangement
+is to publish a filesystem that can be mounted from each execution node that intends to use it.
+For example, the CVMFS~\cite{globalfs-cise-2015} filesystem is used to publish the software used by all of the major LHC
+experiments on hundreds of computing sites around the world.
+This requires that the computing site provide the CVMFS software,
+and permit worker nodes to access the global network.
+
+Figure~\ref{choices} outlines several broad methods for
+accessing such global filesystems.  First, the most desirable
+method is to {\bf mount} the filesystem directly from worker
+nodes, so that the necessary files can be accessed on demand.
+While this is possible at sites that are especially configured
+to support the LHC, it is often not possible at other HPC sites
+that restrict general network access from worker nodes.
+There are many social and technical reasons for these restrictions, but they generally boil down to either security (avoiding intrusions into worker nodes) or efficiency (discouraging high value workers from waiting on network access.)  As a result, many HPC sites are unable
+to use global software repositories.
+
+If that cannot be accomplished, then the next
+alternative is to {\bf copy} the entire repository to
+a shared parallel or distributed filesystem at the site.
+This has the advantage that the filesystem is already
+mounted and available at worker nodes.  The entire
+repository must be duplicated before it can be used,
+regardless of whether an individual job needs it all.
+A more subtle problem is that most parallel filesystems are
+designed for delivering very large files, and are not at all 
+efficient at delivering deep directory trees that
+consist of small files and large amounts of metadata.
+When a large number of workers begin to use the software
+simultaneously, the system can collapse from the metadata
+load.  ~\cite{metafs-pdsw-2017}  (XXX CITE MORE HERE)
+
+A {\bf container} is often used to reduce the metadata load
+on a parallel filesystem.  If a complex software tree is
+deployed into a container, the underlying filesystem will
+treat the container as a single large file, and be more
+efficient at delivering it to many worker nodes simultaneously.
+However, if the entire repository (measured in TB) is placed into
+the container, it is likely to exceed a number of practical
+limits on container size.  Individual worker nodes may have limited
+local disk space, or may be entirely diskless.  Even if the
+large container fits, it is likely that a given job does not need
+\emph{all} of the repository simultaneously, and so it is wasteful
+to transfer unneeded data.
+
+To have containers of reasonable size, it is necessary to {\bf project}
+the required subset of the global repository into a container of reasonable size that can be transferred to each worker node in an
+efficient way. Of course, each job that depends upon a repository may need a different subset of the files, so it will be necessary to
+maintain multiple container images, within the constraints of available 
+storage space.
+
+In this paper, we explore the technique of projecting a global
+filesystem into local containers, in order to efficiently use
+an HPC facility for a large number of jobs.  This requires striking
+a balance between the size of containers, the amount of storage
+space used, and the efficiency of data transfer.
+We consider two phases of the problem: first, how to describe
+and construct a container that serves the needs of a given job,
+and second, how to efficiently manage resources when multiple
+jobs and multiple containers are required.
 
@@
 
+\if 0
 \begin{table*}[t]
@@
 
+\fi
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|r r r|}
+\hline
+Repository & Total Size & inode Count & Deduplicated \\
+ & (GB) & (K) & Data (GB) \\ \hline
+alice & 3,981 & 27,322 & 2,099 \\
+atlas & 6,195 & 120,599 & 3,761 \\
+boss & 120 & 2,418 & 72\\
+cms & 582 & 26,225 & *2,243\\
+geant4 & 298 & 7,128 & 68 \\
+grid & 31 & 714 & 19 \\
+lhcb & 2,135 & 33,737 & 368,070 \\
+na61 & 5 & 26 & 2 \\
+sft & 3,997 & 128,609 & 2,017 \\ 
+ \hline
+\end{tabular}
+\caption{CVMFS repository sizes.}
+\label{tab:repo-sizes}
+This table shows the size, number of inodes (rounded to nearest thousand), and the amount
+of duplicate data that is eliminated using CVMFS.
+\end{center}
+\end{table}
+
 \section{Realization of CVMFS Projection}
@@
 
+\begin{figure}
+    \includegraphics[width=\linewidth]{curated/lxplus_dist/usage_distribution_cdf.pdf}
+    \caption{Usage distribution pulled from CERN LXPlus interactive machines.}
+    \label{fig:usage_distribution}
+\end{figure}
 \if 0
@@
 First, we note that most of the operations are merges.
-This makes sense, as this simulation was at fairly high $\alpha$.
-The total bytes written also closely tracks merges,
-indicating that merging is the dominant source of I/O.
-We can still see inserts over the course of the simulation.
-At more extreme $\alpha$ values,
-we would expect to see one of these operations almost exclusively.
-As the data in cache continues to rise,
-we eventually see the cache limit,
-after which the delete count starts increasing.
-Over the course of the simulation,
-inserts and deletes are filling and emptying the cache such that it remains close to its storage limit.
-We also observe the number of cache hits continue to rise despite deletions.
-As we will see, merging allows for a greater proportion of hits even if the amount of data remains constant.
-This is because frequently used data can be merged,
-reducing duplication.
-The cache limit then ensures that infrequently used parts are eventually removed.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{curated/time-series/time-series.pdf}
-\label{fig:series}
-\if 0
-zoomed out?
-cache limit, deletes start
-merge tracks writes, it's the dominant factor
-hits grow gradually as cache merges and fills
-deletes help prune useless stuff, remaining should be frequent hits
-far ends are either insert or merge, middle alphas have both
-\fi
-\end{figure}
-
-Figure~\ref{fig:series} shows a single simulation of the image management system,
-with several notable details.
-First, we see that most of the operations are merges.
 This makes sense, as this simulation was at fairly high $\alpha$.
@@
 \begin{figure*}
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_cache_percent_plt.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_cache.pdf}
 \hfill
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_cache_percent_plt.pdf}
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_request_percent_plt.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_cache.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_request.pdf}
 \hfill
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_request_percent_plt.pdf}
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_uses_percent_plt.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_request.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_uses.pdf}
 \hfill
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_uses_percent_plt.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_uses.pdf}
 \label{fig:sensitivity}
@@
 \section{Related Work}
-TODO
+
+\subsection{CVMFS}
+The CernVM File System (CVMFS) is a globally distributed filesystem designed for providing efficient,
+read-only access to scientific software.
+CVMFS is primarily used for distribution of High Energy Physics (HEP) software used as part of the Large Hadron Collider (LHC) at CERN.
+Researchers at CERN use CVMFS as the primary means of distributing the analysis and simulation software they develop to the Worldwide LHC Computing Grid (WLCG).
+
+Each WLCG site commits to providing some fixed amount of computational resources for analyzing particle collision events and running simulations.
+Worker nodes at each site must be configured to mount CVMFS as a prerequisite to performing any computational tasks.
+Since CVMFS is implemented as a FUSE module,
+WLCG sites are expected to support FUSE on their worker nodes.
+CVMFS also relies on global configuration files,
+requiring administrative involvement in preparing worker nodes for WLCG jobs.
+In addition, sites generally provision some infrastructure to support CVMFS,
+such as caching proxies to reduce load created by local workers on publicly accessible filesystem servers.
+
+As a result, it is difficult to contribute computational resources toward the LHC experiments without making some site-wide provisioning and configuration changes.
+This is not an issue for the WLCG sites,
+as they provide dedicated computing resources and can customize their setups to meet the requirements of CVMFS.
+For a professor attempting to harness a campus cluster,
+on the other hand, these requirements present a serious barrier.
+Solutions exist such as
+Parrot, a debugging tool and personal filesystem developed at the University of Notre Dame,
+which sidesteps FUSE but still requires unrestricted network access and
+UnCVMFS which locally store all of CVMFS to avoid network access, 
+but relies on large storage and permission mounting.
+For computing at scale, it is important to provide a caching proxy and concise storage,
+which currently require assistance from administrators.
+
+\subsection{HPC}
+
+With the rapidly increasing demands of the LHC experiments at CERN,
+HPC resources are an appealing source of computing power to supplement the WLCG.
+Leadership class machines such as Summit, at Oak Ridge National Laboratory, will by themselves be able to perform a substantial portion of the work of the entire WLCG today.
+Use of national-scale HPC resources across the United States and Europe such as Cori at NERSC, Sierra at Lawrence Livermore National Laboratory, Piz Diant at Swiss National Supercomputing Centre could be key in meeting the computational demands following the LHC's high-luminosity upgrade.
+
+Unfortunately, HPC sites are generally more restrictive than the previously mentioned campus resources.
+In addition to limited privilege on workers and lack of FUSE support,
+HPC sites often impose restrictions on network activity,
+including outbound firewalls preventing access to CVMFS.
+
+The US collaboration of the ATLAS project is currently taking advantage of computing resources at various supercomputers in the United States including Cori at NERSC.
+For the deployment of their software the \texttt{uncvmfs} utility is used to unpack entire CVMFS repositories into standalone images (usually in ext4 or squashfs format) which in turn can be used to build Shifter or Singularity images.
+While these images can very well be scaled out onto a large number of nodes inside the NERSC infrastructure,
+the image build process takes around 24 hours which makes it difficult to deploy up-to-date versions of the software on a regular basis.
+As these images contain complete copies of everything in CVMFS,
+each takes on the order of terabytes,
+necessitating careful storage management.
+In addition, the process requires administrators for image creation, deployment, and cleanup.
+Currently, this approach is used in production on the ATLAS and CMS experiments.
+As additional experiments want to take advantage of the resources at NERSC,
+the administrative burden of managing multiple CVMFS images on multiple software versions increases accordingly.
+For the experiments to use the same images (saving time and cost at NERSC),
+multiple groups would have to come to agreement about a single common set of software versions to use and adjust their pipelines elsewhere.
+Both solutions introduce new inconveniences and administrative burden.
+
+More broadly, the use of static, manually managed container images breaks user assumptions about the operation of CVMFS.
+When running jobs using CVMFS,
+researchers can specify the version of each repo to use,
+allowing precise control over software and dependencies.
+Each experiment has its own policies and procedures for choosing a software version;
+some may use bleeding-edge analysis software or another experiment may prefer older, stable versions.
+CVMFS is able to provide multiple customized views of the global filesystem without incurring much additional cost.
+CVMFS also supports some more exotic features,
+such as symlink resolution controlled by environment variables.
+CVMFS as a whole does not map well to the semantics of a traditional filesystem.
+Projecting CVMFS into such a filesystem can capture at most a single, static version of a repo and must necessarily eliminate some features,
+such as environment-based symlink resolution at runtime.
+
+\if 0
+% https://indico.cern.ch/event/587955/contributions/2937411/attachments/1683156/2705423/WahidBhimji-CHEP18-NERSC.pdf
+Another similar approach in use at NERSC is DVMFS.
+Since creating static container images containing CVMFS is costly and difficult to customize,
+DVMFS instead uses Cray DVS IO forwarders to mount a copy of CVMFS served over NFS.
+This allows for less administrator involvement,
+since the mirrored copy of CVMFS can be automatically updated on a regular basis.
+After dealing with several implementation problems,
+DVMFS is currently deployed at NERSC as an alternative to static images.
+While this approach affords more flexibility in choice of repos and software versions,
+it increases the job startup time and places additional requirements on the infrastructure.
+To be usable in production, DVMFS requires a sufficient number of DVS servers to be provisioned.
+DVMFS is nonetheless a viable solution as long as the costs of developing/maintaining the service and allocating infrastructure do not grow too high.
+\fi
 

### 2019-01-24 paper.tex
Submitted Paper Version

@@
 
-\author{Tim Shaffer}
-\email{tshaffe1@nd.edu}
-\orcid{}
+\author{Tim Shaffer$^\star$, Nicholas Hazekamp$^\star$, Jakob Blomer$^\dagger$, and Douglas Thain$^\star$}
+\email{tshaffe1@nd.edu, nhazekam@nd.edu, jblomer@cern.ch, dthain@nd.edu}
+%\orcid{}
 \affiliation{%
-  \institution{University of Notre Dame}
-  \city{Notre Dame}
-  \state{Indiana}
-  \postcode{46556}
+  \institution{$^\star$ University of Notre Dame   $^\dagger$ European Laboratory for Particle Physics (CERN)}
+%  \city{Notre Dame}
+%  \state{Indiana}
+%  \postcode{46556}
 }
 
+%\author{Jakob Blomer}
+%\email{jblomer@cern.ch}
+%\orcid{}
+%\affiliation{%
+%  \institution{CERN}
+%  \city{Geneva}
+%  \country{Switzerland}
+%}
+
+\if 0  
 \author{Nicholas Hazekamp}
@@
 
-\author{Jakob Blomer}
-\email{jblomer@cern.ch}
-\orcid{}
-\affiliation{%
-  \institution{CERN}
-  \city{Geneva}
-  \country{Switzerland}
-}
-
 \author{Douglas Thain}
@@
 }  
+\fi
 
@@
 applications to make use of local filesystems and container
-technologies.  In this work, we present a technique for efficiently.
+technologies.  In this work, we present a technique for efficiently
 \emph{projecting} a global filesystem into local resources.
@@
 and LSST make use of hundreds of thousands of cores spread across
-many computing facilities around the world in order to achieve the massive scale needed simulate and analyze data from high bandwidth
+many computing facilities around the world in order to achieve the 
+massive scale needed to simulate and analyze data from high bandwidth
 scientific instruments.  In order to ensure consistency of software
@@
 can be quite large, consisting of thousands of software packages
-consuming millions of files and terabytes of data.  Software
+containing millions of files and terabytes of data.  Software
 such as CVMFS~\cite{globalfs-cise-2015} is used to distribute
@@
 actually used, or by static analysis of its dependency tree.
-We describe Shrinkwrap, a tool that performs this analysis on
-a Unix application, and produces a minimal container
-that can be used to execute the application, or one closely related.
-We analyze the efficiency of this tool on a selection of high
-energy physics applications, and demonstrate storage efficiency of this technique, and measure the performance and storage efficiency.
+Using the subset or specification from either method, a projection
+then captures the job's software stack.
+From this captured software, a minimal container can be
+constructed to allow for more efficient storage, transfer, 
+and use. We describe Shrinkwrap, 
+a tool that produces a minimal software stack 
+that can be used to execute the application, or ones closely related.
+We analyze the size and make up of global filesystem repositories,
+specifically in CVMFS, to show why existing solutions fall short.
 
@@
 
-\if 0
-
-Scientific research has a large and rapidly growing need for computation.
-Researchers have a number of options at their disposal to meet those needs,
-including campus-local or organization-local resources,
-cloud providers,
-grid computing,
-and large HPC sites.
-Moving applications between these classes of resources can present problems,
-as the assumptions made during development may no longer hold.
-For example, organization-local resources and some grid sites commonly provide access to shared filesystems at each compute node.
-In addition, there is some flexibility in system configuration to support an organization's primary research activity.
-A new computing environment may be much less open to changes for a specific application,
-especially large, multi-tenant resources such as HPC sites.
-In these cases, it is left to scientists to adapt their application software to the available resources.
-
-Many of our research collaborators have experienced significant frustration in preparing an existing scientific application to run on a new type of resource.
-As a concrete example for this work,
-we consider the challenges in running simulation and analysis workloads at CERN on HPC sites.
-These applications assume several pieces of infrastructure to be in place and require some configuration and customization to the compute environment.
-This is not an issue when running on Worldwide LHC Computing Grid~(WLCG),
-which is a dedicated resource built to meet CERN's substantial computing needs.
-The WLCG currently satisfies most of the computational needs of the LHC experiments at CERN.
-With the upcoming high luminosity upgrades to the LHC,
-the computational resources needed will will increase by at least an order of magnitude by 2026,
-% On CERN's page it estimate 50-100X need, argonne had a report that was 10-20X
-prompting researchers to turn to available high performance computing (HPC) resources to meet demand.
-%In contrast with the WLCG's single slot submissions,
-%HPC applications typically consist of parallel jobs using several cores up to a full machine for each submission.
-
-Most HPC systems, however, have site-specific configuration and security requirements.
-These can take the form of limited network availability at computation nodes,
-limited container support,
-and restricted access to privileged system features (such as FUSE support).
-These new restrictions present challenges when porting existing scientific applications,
-which expect required software to be readily available,
-make a number of assumptions about how the system should be configured,
-and may not be designed with tasks such as converting between image and container formats in mind.
-
-In the case of the simulation and analysis workloads at CERN,
-unrestricted network access and FUSE support are critical for providing global filesystem access.
-Nearly all research software used at CERN assumes the availability of the CernVM File System (CVMFS),
-a global filesystem built for distributing software and configuration data.
-CMVFS provides a read-only, distributed file system using HTTP for network transport and extensive local
-caching to rapidly distribute updates in software across the WLCG.
-Dependence on CVMFS to distribute experiment software currently makes it impossible to take advantage of many HPC sites.
-CVMFS requires unrestricted network access and is implemented as a FUSE module,
-both of which sites often disallow for security and management reasons.
-
-Some sites, such as NERSC,
-have attempted workarounds for providing CVMFS,
-such as creating large images several terabytes in size that contain all of the software stored in CVMFS or
-maintaining local mirrors of CVMFS available via site-local shared filesystems.
-However, both of these existing solutions require 
-hours to days of attention from system administrators to deploy or update,
-which leads to lag between the available software and the current version.
-Additionally for systems that utilize large container images, large amounts of storage are needed to maintain different versions.
-
-Since each LHC experiment job uses 
-only a relatively small portion of CVMFS,
-a more fine-grained approach to preparing container images would reduce the time spent creating images and allow for more precise control of the software versions in use.
-This could also reduce the storage required at each worker node and the network bandwidth spent transferring images.
-It is often possible to determine dependencies for particular applications statically by examining submission commands, setup scripts, etc.
-This approach, while reliable, tends to be labor-intensive and tied to specific experiments and software stacks.
-We therefore modified the reference implementation of CVMFS to allow us to capture the filesystem interactions of live HEP applications.
-Using live tracing and static analysis,
-we are able to determine exactly which parts of the \texttt{/cvmfs} file system tree CVMFS must be included in a container image for a particular job.
-We developed a tool, Shrinkwrap, that can efficiently \emph{project} CVMFS (or a subset thereof) at a particular version into a traditional filesystem.
-Such a \emph{projection} of CVMFS is suitable for inclusion in a static container image that does not depend on FUSE or network access.
-
-While working with projected container images and Shrinkwrap,
-we found that the cost of creating, distributing, and storing customized images for a large number of jobs and users creates unexpected challenges as well.
-Shrinkwrap is designed to efficiently manage projections by deduplicating data in the filesystem.
-Disk images, however, must contain complete copies of all included data.
-The na\"{i}ve approach of creating a projection of CVMFS for each application results in frequent duplication of data and significant overhead in terms of computation and bandwidth.
-Depending on the granularity of the projection,
-image creation time can easily exceed individual job's actual compute time.
-The alternative extreme, i.e.\ building a single large image that satisfies the requirements of all applications,
-forces all jobs to use the same repository revision,
-which can result in incompatibilities across applications.
-Updates to the repository or changes in applications and requirements can again result in large amounts of overhead due to frequently creating, transferring, and storing these very large images.
-
-Observing that jobs frequently have significant overlap in dependencies,
-we designed a system to more efficiently manage container images on job submission.
-First, we defined a metric for determining the similarity between projections.
-This gives the possibility of identifying existing container images that could be re-used to fulfill a new request.
-We also determined that in some cases,
-it is less expensive to augment an existing, similar projection to allow a single container image to support multiple distinct applications.
-We bring these tools together as a proof-of-concept submission system for making online decisions about whether to create, reuse, or augment container images to efficiently support high-throughput HEP jobs on HPC resources.
-
-Compared to existing solutions,
-Shrinkwrap created more compact projections than traditional methods such as rsync copies,
-and allowed for more efficient updates to existing projections.
-Based on the existing projects to produce static CVMFS container images and traces of benchmark applications,
-we examined the sizes and structures of several projected container images.
-Based on these properties,
-we simulated the behavior of our image management system over a large number of job submissions.
-We found that our approach can effectively manage the container images for a continuous stream of submissions with varying software dependencies and versions with less storage overhead than creating a new image per application.
-In addition, our results show that applications can still take advantage of precise dependency and version specifications to reduce image creation and transfer overhead compared to working with large/complete mirrored copies of CVMFS.
-
-\fi
-
 \begin{figure}
 \includegraphics[width=\linewidth]{drawings/methods.pdf}
-\label{choices}
 \caption{Global Filesystem Access Methods}
+\label{fig:choices}
 \end{figure}
@@
 repositories that consist of terabytes of data and millions of files.
-Unlike "traditional" high performance applications that are
+Unlike ``traditional'' high performance applications that are
 single, statically-linked executables, modern applications are
@@
 libraries that may be dynamically linked, and configuration and
-data files loaded at runtime.
+data files loaded at runtime.~\cite{casestudy-poster-chep-2015}
 
@@
 
-Figure~\ref{choices} outlines several broad methods for
-accessing such global filesystems.  First, the most desirable
+\Cref{fig:choices} outlines several broad methods for
+accessing such global filesystems.  The simplest
 method is to {\bf mount} the filesystem directly from worker
@@
 that restrict general network access from worker nodes.
-There are many social and technical reasons for these restrictions, but they generally boil down to either security (avoiding intrusions into worker nodes) or efficiency (discouraging high value workers from waiting on network access.)  As a result, many HPC sites are unable
-to use global software repositories.
+There are many social and technical reasons for these restrictions, but they generally boil down to either security (avoiding intrusions into worker nodes), efficiency (discouraging high value workers from waiting on network access), or configuration (kernel modules are not available to perform the mount.)  As a result, many HPC sites are unable
+to use global filesystems of this form.
 
@@
 efficient at delivering deep directory trees that
-consist of small files and large amounts of metadata.
-When a large number of workers begin to use the software
+consist of small files and large amounts of metadata,
+which describes these repositories as seen in \Cref{fig:cdf}.
+When a large number of workers begin to use the software or perform similar metadata operations
 simultaneously, the system can collapse from the metadata
-load.  ~\cite{metafs-pdsw-2017}  (XXX CITE MORE HERE)
+load.~\cite{metafs-pdsw-2017,spindle,torrent,exa}
 
@@
 to transfer unneeded data.
+This concept is a driving influence on projects like Slacker\cite{194430}.
 
@@
 
-\section{Defining Application Dependencies}
-
-While the entirety of CVMFS is fairly large,
-a single application would never use every piece of data.
-Rather than attempting to provide a complete copy of CVMFS locally,
-which entails significant storage costs,
-we could instead provide a limited subset of the complete filesystem that is sufficient for a particular job.
-This approach is more in line with traditional software packaging practices,
-where software is provided as needed in the form of RPM files or another similar format.
-Working with software packages is simpler in that dependencies are explicitly specified in the package file or database.
-With CVMFS, no such definitive specification exists.
-This issue is common to shared/global filesystems,
-since users expect the complete contents to be available on demand and so devote little time to explicitly defining dependencies and other requirements.
-
-In the case of CVMFS,
-building an accurate 
-specification of the files, libraries, and packages
-needed for each experiment is the first step toward creating usable container images.
-Such file specifications can be used when 
-projecting CVMFS into realized filesystems.
-The correctness and storage overhead of the projected filesystem hinges on the clarity and accuracy of the file specification.
-File specifications can be derived from several
-different sources, such as 
-usage traces, static analysis, or package listings.
-The end goal of any of these methods is to 
-create a comprehensive list of needed files and structure.
-The nature of global filesystems flexible behavior
-further extends file specifications by requiring
-not just the files and structure, but also what
-version/revision was used to verify the contents as well.
-
-\subsection{Creation of Specification}
-File specifications can be created several ways,
-with the obvious option being handwritten.
-However, this quickly becomes unfeasible with
-even a moderate number of requirements.
-This leaves the need for an automate method to create
-specifications.
-
-\subsubsection{Tracing}
-The first method is to create a sample trace for your
-jobs. 
-Creating a trace from a sample job allows 
-for a listing of actual files used, which
-can be more accurate than handwritten listings.
-As these jobs are typically on the order of thousands
-to millions of jobs, sampling a small subset
-can provide good software coverage with little
-additional overhead. % the computation is still used
-The burden of creating these traces is
-dependant on the method used,
-which could be an LD\_PRELOAD library,
-using Parrot to track used files, % I think parrot was mentioned previously
-or building the tracing in to the application.
-
-Tracing is attractive as it provides a consistent 
-behavior between experiments and use cases.
-In addition, it is easier to automate and sample arbitrary applications without assistance from each user.
-This allows the solution to port between applications
-quickly without having to delve in to the applications
-specific configuration.
-As previously mentioned, there are several methods for
-obtaining a trace from a file.
-
-Despite assumption that both CVMFS and FUSE will be limited
-on the HPC site, this methods was used as it is reasonable to
-allow for sampling of these jobs on WLCG. 
-WLCG provides the flexibility of using CVMFS and FUSE, without
-having to implement or provision a separate tracing infrastructure
-at each site.
-Additionally, as many of these jobs would have originally been 
-destine to run on WLCG, the sampling sizes are unlikely to 
-create significant overhead.
-Using CVMFS for the tracing allows CVMFS's just-in-time deployment
-of files to patch the environment and give an accurate view without
-the risk of missing files.
-
-\begin{figure}
+\begin{figure}[h]
 \includegraphics[width=\linewidth]{curated/cdf/cdf.pdf}
-\caption{The distribution of file size for each repository.
-From this we can see that the large majority of files
-accumulate in the kilobyte to megabyte range.}
+\caption{Distribution of file sizes for each repository.}
 \label{fig:cdf}
-\if 0
-lots of small files
-table shows duplication
-long tail of large files
-different scales, similar usage across repos
-
-\fi
 \end{figure}
 
-\subsubsection{Packages}
-The second automated method is to use static
-analysis for discover requirements.
-This would allow the application to be analyzed
-prior to running to determine what packages are
-linked.
-Static analysis can be done at several levels,
-each providing different levels of control 
-and flexibility.
-This could be more language specific in the 
-form of following import lines in python or
-include lines in C.
-The alternative is to use repository specific 
-methods for setting up the environment,
-such as bash scripts for pointing to correct 
+%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Creating Projections}
+
+Having defined a projection of a global filesystem as a subset of the filesystem stored locally,
+the first step toward creating projections of a global repository is deciding which components to include.
+Scientific applications use a wide variety of files, libraries, and packages,
+and tend not to respond well to missing components.
+When directly mounting a global filesystem or making a local copy,
+the entire contents of the filesystem are assumed to be available,
+even if any single job only uses a small portion of the total data.
+As such, there may not be a convenient and complete description of which pieces of data are required.
+Attempting to use a projection based on an incomplete specification leads to potentially intermittent job failures and wasted computing resources.
+On the other hand,
+including too broad a selection of data in a projection is undesirable as it increases overhead due to extraneous data in the container images.
+
+The simplest way to create a specifications for a projection is manually listing each piece to include.
+However, this quickly becomes unfeasible with even a moderate number of software component requirements.
+In addition, manual specifications do not necessarily reflect the actual requirements of an application.
+Changes to software or simple mistakes can result in incomplete specifications,
+and include unnecessary files and data that result in bloated container images.
+Manual specifications also practically require consistent, uniform software organization.
+This is not a safe assumption within a single repository,
+to say nothing of supporting multiple software sources.
+This leaves the need for an automated method to create
+specifications, such as 
+ static analysis or runtime tracing.
+To create a comprehensive list of required files and data,
+it may be necessary to use a combination of these approaches.
+
+\subsection{Static Dependency Analysis}
+
+Static dependency analysis takes advantage of metadata, language, and application-specific knowledge to
+discover requirements.
+This technique allows applications to be analyzed
+prior to running to determine what dependencies from the global filesystem are required.
+Static analysis can be informed by several sources of information,
+each with varying levels of detail and flexibility.
+Language-specific analysis of software components might inspect the \texttt{import} statements in Python or
+\texttt{include} directives in C.
+These approaches only apply to specific languages or frameworks,
+and require some knowledge about how the application is structured.
+The software stack used for experiments can also provide information about dependencies in the way it sets up the working environment.
+This form of static analysis might involve examining shell scripts for selecting  correct 
 software versions, or 
-analyzing SCRAM\cite{DBLP:journals/corr/cs-OH-0306014} lines
+analyzing any number of systems built for 
+declaring dependencies, such as 
+modules\cite{McLay:2011:BPD:2063348.2063360},
+Spack\cite{Gamblin:2015:SPM:2807591.2807623},
+VC3-Builder\cite{tovar-ic2e-2018},
+or Nix\cite{Dolstra2008NixOSAP},
 to determine the configuration.
-In the SFT CVMFS repository, where many core research packages are located,
-dependencies are mapped for many packages using buildinfo files.
-Chaining these buildinfo files together allow for a rough dependency
-tree to be create, which was used to sample for \Cref{fig:picks}.
-With either of these methods the issues
-that arise come from bleeding edge or
-experiment specific code that is not directly
-handled by the configuration.
-There is a significant mix between languages,
-configurations methods, and software organization
-practices even within a single repository,
-which compounds the above issues in this method.
-
-An combination of tracing and static analysis is the
-use of package listings. 
-Package listings provide a high level indication
-of the software needed, and allow for 
-flexibility on the specifying user about the
-exact files needed.
-Using high level packages provides a method
-for handwritten/tuned specifications that 
-include all aspects of the package.
-The downside of using packages is that in
-their flexible, forgiving specification
-they can be bloated and include unnecessary
-files and data.
-Packages also require that there is consistent
-software organization, which often differs
-between repositories.
+Additionally it is not uncommon for experiments 
+to design and use their own methods such as
+SCRAM\cite{DBLP:journals/corr/cs-OH-0306014},
+which is used in LHC experiments.
+This kind of approach is highly tailored to specific applications or software frameworks and is difficult to generalize to other uses.
+
+In the SFT CVMFS repository, where many of the core research packages at CERN are hosted,
+dependency information is included in build metadata associated with each package.
+This metadata gives a convenient way to construct a dependency
+tree of the entire repository.
+We constructed such a dependency tree for the SFT repository to use in the simulations discussed later.
+\Cref{fig:picks} gives an indication of the structure and sizes of software components in the repository.
+To construct that figure,
+we started by repeatedly making random package selections of varying sizes and measuring the total sizes of the selections.
+The two lines near the bottom of the graph show these counts and sizes.
+Then for each selection of packages,
+we used the extracted package metadata to recursively include each package's dependencies.
+This recursive resolution simulates the procedure that a program might perform when starting up and loading libraries.
+After finding the complete set of dependencies required by each selection,
+we plotted the resulting package counts and sizes.
+We repeated this procedure 100 times for each choice of specification size,
+up to specifications of 1,000 packages.
+
+The plot in \Cref{fig:picks} shows the medians of the resulting measurements.
+We observe of course that selection size increases uniformly along the $x$ axis.
+The on-disk size of the selections appears to increase proportionally to selection size.
+Recursively including the dependencies of these same selections,
+however, results in a significant increase in number of packages and size on disk.
+For small selections (less than 100~packages),
+the complete images with dependencies might contain 5x as many packages as requested,
+with storage increasing proportionally.
+With larger selections, however,
+this increase becomes less prominent.
+This non-linear change is the result of the tree structure of the software dependencies.
+There are a number of core components that are transitive dependencies of a large number of packages.
+These common dependencies are only included once,
+so subsequent selections that depend on them do not add them again.
+As the number of selections increases,
+this curve will approach the total repository size.
+Also note that the non-linear increase in size with increasing selection size means that the size of combining two selections of 100 packages (and their dependencies) is not as simple as adding them together.
+Due to duplication, we would expect the resulting combined size to be lower,
+as if we were choosing 200 packages from the start.
 
@@
 non-linear, so bigger ones have more duplicates
-appraoches repo size in limit
+approaches repositories size in limit
 merging is subject to drop-off(more duplication)
@@
 
-\subsection{Unique Consistency Requirements}
+\if 0
+\begin{figure}
+    \includegraphics[width=\linewidth]{curated/lxplus_dist/usage_distribution_cdf.pdf}
+    \caption{Usage distribution pulled from CERN LXPlus interactive machines.}
+    \label{fig:usage_distribution}
+\end{figure}
+
+spread across everyone, lot of commonality/overlap
+few with very high usage, very large number of rarely used
+lot not observed
+\fi
+
+\subsection{Runtime Tracing}
+Rather than examining the static data and metadata associated with a software collection,
+we can use runtime tracing to identify the dependencies used during a particular job.
+This trace of a sample job gives an exact listing of the files used,
+which is more accurate and reliable than handwritten listings.
+A potential drawback is that tracing does not determine every possible dependency,
+only those used during a particular sample.
+This can be problematic for non-deterministic or data-dependent workloads.
+As such, it is preferable to trace a number of jobs.
+Since scientific applications routinely consist of thousands
+to millions of jobs, sampling a small subset
+can provide good software coverage with little
+additional overhead. % the computation is still used
+The performance cost of creating these traces is
+dependant on the method used.
+A tracer library based on \texttt{LD\_PRELOAD} can be extremely lightweight but require modification to the execution environment.
+More invasive methods include using Parrot to track used files, % I think parrot was mentioned previously
+or building the tracing as part of the application.
+
+An important advantage of runtime tracing techniques is that they can function reliably across multiple languages and applications.
+This makes it is easier to automate sampling of unknown applications without assistance from each user,
+since there is no need to delve in to the particular application
+design and configuration.
+
+As part of this work, runtime tracing was introduced
+into the CVMFS FUSE client and can be activated for sampling.
+When tracing an application,
+each filesystem access is recorded and added to the trace log.
+Tracing through CVMFS is convenient as it does not require changes to applications,
+and the FUSE client is widely deployed across compute sites.
+Users have the option of collecting this trace information themselves by re-configuring the client.
+For applications running on the Worldwide LHC Computing Grid~(WLCG),
+it would also be possible to instrument a fraction of runs to build a profile of a given task.
+This introduces the possibility of automatically collecting traces to use when porting applications to more restrictive environments.
+
+\subsection{Building Projections}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|r r r|}
+\hline
+Repository & Total Size & inode Count & Deduplicated \\
+ & (GB) & (K) & Data (GB) \\ \hline
+alice & 3,981 & 27,322 & 2,099 \\
+atlas & 6,195 & 120,599 & 3,761 \\
+%boss & 120 & 2,418 & 72\\
+cms & 582 & 26,225 & *2,243\\
+%geant4 & 298 & 7,128 & 68 \\
+grid & 31 & 714 & 19 \\
+lhcb & 2,135 & 33,737 & 368,070 \\
+%na61 & 5 & 26 & 2 \\
+sft & 3,997 & 128,609 & 2,017 \\ 
+ \hline
+\end{tabular}
+\caption{CVMFS repository sizes.}
+\label{tab:repo-sizes}
+{\small \it This table shows the size, number of inodes (rounded to nearest thousand), and the amount
+of duplicate data that is eliminated using CVMFS.}
+\end{center}
+\end{table}
+
+Several existing utilities such as UnCVMFS and 
+rsync have been used to create projections of CVMFS.
+UnCVMFS is a utility that projects an entire
+CVMFS repository, at a set revision, into the local
+filesystem. 
+UnCVMFS utilizes CVMFS's efficient, content-based storage
+to deduplicate the underlying data, but requires
+entire repositories to be projected.
+When used with several repositories for large projects,
+unCVMFS can easily consume terabytes of disk space for a single projection.
+In addition, it is expensive to update to different revisions of a repository using unCVMFS,
+which can create headaches when working with different experiments and analyses that may have conflicting revision requirements.
+
+Rsync allows for file lists to specify and download
+subsets of CMVFS repositories, but because it
+was designed for general filesystem synchronization
+rsync is not aware of the internal organization of CVMFS.
+This leads to slow operations and unnecessary duplication of data,
+especially when creating multiple projections.
+
+We built the CVMFS Shrinkwrap utility to create projections of CVMFS with better efficiency and finer control than is possible either unCVMFS or rsync individually.
+Shrinkwrap uses the CVMFS's client interface to
+create and update CVMFS projections.
+Shrinkwrap uses a CVMFS configuration to specify the
+revision of the repository, a file-specification list
+to determine the repository subset to project, and
+deduplicates the data in a user specified location.
+Shrinkwrap can be configured to share the data
+repository between several repositories and even
+projections of the same repositories.
+This allows for multiple realized projections
+to coexist, with the desired projection
+mounted for use.
+In \Cref{figure:shrinkwrap-arch} you can see an outline
+of the design of the Shrinkwrap utility.
+
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{drawings/shrinkwrap-structure.png}
+\caption{Operation of Shrinkwrap}
+\label{figure:shrinkwrap-arch}
+\end{figure}
+
+\subsection{Building Container Images}
+
+After creating a file specification for an application and projecting the required data from the global filesystem,
+the next step is to create a container image to be used locally.
+The simplest approach for automatically projecting container images is to create a suitable image on each batch job submission.
+Assuming a file specification has already been created,
+it is passed to Shrinkwrap to create the requested projection,
+which is then processed into a container image,
+for example using \texttt{mksquashfs}.
+Unfortunately, this solution is very wasteful of computing time and storage space.
+Even a minimal image consisting of a precise set of required libraries and frameworks ranges in the tens of gigabytes,
+depending on the application.
+A full copy of a repository would consume multiple terabytes of disk space and take the majority of a day to create.
+In addition, image creation must be completed before any actual processing occurs,
+i.e.\ there is no way to start with a partial image that supports the first steps of a pipeline and finish building the image as the pipeline runs.
+There is also the cost of transferring and storing freshly created images on any nodes that will use them,
+further delaying job dispatch and increasing load on the network and storage infrastructure.
+
+The disk I/O alone required to create even a targeted image is likely to take more time than a short simulation or analysis job.
+Thus a simple performance improvement is to re-use previously created container images for multiple jobs.
+A single experiment submitting computational tasks is likely to request the same projection,
+at least for jobs submitted around the same time.
+It is straightforward to check for exact equality between projections as long as the system tracks the set of files that went into their creation.
+If there are few CVMFS users at a site,
+this method could be sufficient.
+Users simply maintain image collections for use with their jobs.
+
+Unfortunately, a number of real-world usage patterns are poorly handled by this approach.
+A small change to a projection,
+such as adding a library or updating dependencies,
+results in the creation of a second, unrelated image.
+Even a small change such as removing an unnecessary documentation file results in a completely new container image.
+This is a manageable nuisance when manually testing or developing a system,
+but would present a problem for a system managing images automatically and receiving updated specifications over time from tracing jobs.
+
+In addition to unnecessary duplication for a single user/application,
+there is also significant duplication across multiple users.
+In the simulation and analysis pipelines at CERN,
+a number of large components (such as ROOT) are used nearly universally.
+These components provide basic functionality and infrastructure that each experiment's software stack builds upon.
+Moreover, these same common infrastructure components must be included regardless of user or application.
+Thus even if every user of the system only creates a single container image,
+there is a duplicate copy of all such common infrastructure components for each user.
+This is problematic if we would like to design a system that can support many users at a site.
+
+These two sources of duplication tend to compound each other,
+leading us to conclude that a more sophisticated approach is necessary to work with realistically-provisioned workers and system infrastructure.
+These issues are fundamentally related to unnecessary duplication of data among container images.
+The same types of issues motivate the use of dynamic libraries for everyday applications.
+With statically linked executables,
+each application includes a complete set of required libraries.
+While very stable and reliable,
+OS distributions prefer to use dynamic linking.
+This allows applications to share a single copy of each library,
+so that foundational components such as libc are not copied into every executable,
+and so that updates and changes to the set of installed packages do not require rebuilding the entire system.
+For this approach to work,
+package maintainers must arrange all software packages in the system to ensure compatibility,
+and the OS distribution must set policies for system organization, update procedures, etc.
+Container images here function similarly to statically linked executables,
+but since multiple complete OS distributions are stored within CVMFS,
+and since we are dealing with multi-terabyte repositories and entire frameworks and software stacks rather than individual OS libraries,
+the scale involved is of course larger.
+In the case of CVMFS, there is no notion of a ``package mainainer'' or central authority over all software.
+Each experiment manages its own software repository as it chooses,
+with some components and repositories shared between multiple applications.
+Since we would like to be able to automatically handle existing applications that are not designed with central management in mind,
+we must instead look to other approaches for dealing with data duplication as a result of copying data into container images.
+
+\section{Managing Projected Images}
+
+The representation of projections as disk images is problematic for deduplicating data.
+Internally, Shrinkwrap uses hard links to a common data store so that a single file is only stored once across all projections in the system.
+On creating disk images, however,
+these links are broken.
+Each disk image must contain a complete copy of all data,
+so that a file must be included (and take up space) in every image that includes it.
+Docker's solution is to build container from layers of archives that can be reused.
+Docker layers are explicitly defined when writing a Dockerfile.
+With CVMFS, however, there is no existing layer-like organization or universal rule that could derive layers for existing applicaions.
+Working with layers defined from the outset is also a poor fit if we would like to support partial, gradually improved information from traces.
+We therefore desire some sort of deduplication that can be applied for existing applications to an assortment of container images after creation.
+
+To develop a deduplication strategy,
+we started with some observations about the projection and image properties.
+For ``large'' images,
+there is limited opportunity for deduplication.
+Each large image is a fairly complete copy of a particular repository.
+While there is bound to be some overlap between versions of the same repository,
+an updated image will contain a large number of updated software components,
+making a ``deduplicated'' between two versions significantly larger than either individually.
+There are also likely to be some conflicts in the filesystem if software components re-use the same paths for different files between versions.
+This case is akin to combining RHEL~6 and RHEL~7 in an attempt to save space.
+While there are similarities between the two,
+the significant changes and conflicts make this a poor strategy.
+Thus for our strategy we would like to avoid ``deduplicating'' images that have large non-overlapping portions.
+
+Likewise, there is little reason to attempt to deduplicate large images built from different repositories.
+Such images would have little to no overlap,
+so the only benefit to combining the two would be reducing the number of images by one.
+This comes at the cost of needlessly storing and transferring the unused half.
+This case is akin to bundling a complete Debian installation alongside RHEL.
+Our deduplication strategy should only take effect for images with significant overlap.
+
+It is also undesirable to combine images of drastically different size.
+As an example, suppose a user creates a small projection as a subset of a repository for running a specific application.
+This small image overlaps completely with a large image of the whole repository,
+and there is likely no conflict.
+Automatically combining the large and small image, however,
+can have unintended consequences.
+Small, tailored images minimize transfer and storage overhead,
+and can be important for operating at scale.
+It is therefore unwise to surprise users by suddenly making their container images several orders of magnitude larger.
+This is akin to distributing a library exclusively by placing it within a disk image of a Debian installation.
+We would therefore like to attempt deduplication only for
+{\bf similarly-sized images, with large overlap and with no conflicting paths}.
+
+\subsection{Measuring Similarity Between Projections}
+
+    Having set out the conditions under which to consider deduplicating projected images,
+we are in a position to define a metric to quantify the similarity between images.
+This would allow us to identify projections that are ``close'' (for some definition of close) as candidates for optimization.
+We chose the Jaccard distance under appropriate choice of set elements as it conveniently satisfies the desired qualities set out above,
+and is very well used and studied.
+In our case, we must describe a container image as consisting of a set of (path, content hash) tuples.
+CVMFS already stores files by the hashes of their contents,
+so this information is readily available.
+Considering both the file names and contents is important as CVMFS allows for updates to file contents over time.
+We chose fine-grained tracking of file contents rather than simply recording the version of the CVMFS repository to afford additional opportunities for dedupliaction.
+Since each update to a repository does not change the contents of every (or even the majority of) files,
+data from multiple repository versions may well be able to coexist in a single container image.
+
+For two sets $A$ and $B$,
+the Jaccard distance $d_j$ is defined as follows.
+\[
+d_j(A, B) = 1 - \frac{|A \cap B|}{|A \cup B|} = \frac{|A \cup B| - |A \cap B|}{|A \cup B|}
+\]
+
+For this application in particular,
+the Jaccard distance captures several desirable properties of projections.
+In the case of two projections that differ only by one file,
+the Jaccard distance will be small.
+Likewise for a pair of projections with nothing in common,
+the distance will be high.
+And since the Jaccard distance is a proper metric,
+related projections will tend to occur in clusters;
+if two projections are both found to be extremely close to a third projection,
+we can conclude that the two are themselves close by the triangle inequality.
+Other choices of metric may well be viable,
+but we found the Jaccard distance to be both well-studied and amenable to efficient approximate implementation.
+As described in the following,
+we used MinHash~\cite{minhash} as a constant-time approximation of the Jaccard metric as a first pass at selecting similar images.
+Such an approximation is important in practice due to the sizes of the data involved;
+metadata listings alone (not including any file data) consumed multiple gigibytes of storage in some of the container images we evaluated.
+
+\subsection{Container Management as an Online Problem}
+
+\begin{figure}
+\centering
+\fbox{\includegraphics[width=\linewidth]{drawings/system_architecture.pdf}}
+\caption{Architecture of a Site-wide Image Management System}
+\label{figure:sys-arch}
+\end{figure}
+
+ Combining the tools and techniques introduced so far,
+we are now ready to address the central problem in managing projected images as a site-wide service.
+Rather than relying on administrators to manually manage images,
+we must define a method for making a decision online as to how to efficiently satisfy the dependency requirements for submitted jobs.
+Our method must be suitable for inclusion either in individual researchers' workflows or as part of the setup for batch job scheduling, and must be efficient in terms of both computation time and storage space.
+Our approach will be to track similarity between projections and allow existing images to be augmented rather than creating new ones for each job request.
+\Cref{figure:sys-arch} gives an overview of the structure of this system.
+
+Our method will be based around a cache of container images used to satisfy job specifications.
+Upon receiving a request for a projection,
+we first hash the new projection according to the MinHash algorithm.
+We always store the hash values computed for every projection we build to allow for fast clustering.
+Next, we use MinHash to approximate the Jaccard distance between the new projection and all projections previously built.
+This computed distance can be cached to ensure that we estimate the distance between each pair of images only once.
+Since comparisons using MinHash take constant time for a given probability bound independent of the size of the sets involved,
+we can quickly identify cached projections that are similar to the new request.
+
+To decide if two projections are ``close enough'',
+we define the parameter $\alpha$ as the maximal Jaccard distance between closely related projections.
+Since Jaccard distance is by definition between zero and one,
+$\alpha$ must be in the same range.
+This $\alpha$ parameter is something like the ``globbiness'' of the system.
+Small choice of $\alpha$ requires that projections be extremely close before considering them for merging.
+In the extreme case with $\alpha = 0$,
+only \emph{identical} images will be considered close so no images will be merged.
+This results in a larger number of independent images.
+Choosing $\alpha$ to be larger makes it more likely for images to be considered similar and merged.
+This results in more augmented images that serve multiple tasks.
+In the extreme case of $\alpha = 1$,
+\emph{every} pair of images is considered close and merged if possible.
+This results in large container images that have accumulated many projections.
+We explore tradeoffs in terms of efficiency and performance when choosing $\alpha$ in \Cref{sec:eval}.
+
+Having selected a set of candidate projections that are close enough to the new request,
+we next check each in turn to find either a superset of the requested projection or one without version incompatibilities.
+We first sort the list of candidate images by Jaccard distance.
+This ensures that our system prioritizes candidates that are similar in size to the requested image.
+In the case of finding a superset,
+we can simply use the corresponding cached container image.
+Presumably this superset was created by the merging process below,
+so we are done.
+Otherwise, we continue to look for a compatible projection
+Two projections are incompatible if they both include the same path but point to different data.
+This situation can arise if for example a library in CVMFS is updated in place.
+Old and new projections would include an identical filesystem path,
+but the new version would have different contents.
+This situation can arise in more subtle ways,
+for example with symlinks controlled by environment variables.
+Since projections of CVMFS must statically resolve such links,
+incompatibilities can arise despite paths appearing compatible under CVMFS.
+Any candidate projections that are incompatible with the new request are removed from further consideration.
+
+If no compatible projections were found among the candidates,
+we create a container image for the projection as requested and add it to the cache.
+Otherwise, on finding a compatible cached projection the next step is to augment it with the contents of the new request.
+We choose the first compatible image from the list of candidates for merging,
+i.e.\ the compatible image with the smallest Jaccard distance,
+which again prioritizes images that are similar in size and contents.
+We can use Shrinkwrap to efficiently augment projections to include new or updated contents.
+Next, we build a container image based on the newly updated projection.
+The computational cost of this step could become high since we are creating larger combined projections.
+Appropriate choice of $\alpha$, however,
+bounds the distance between the combined projections such that the cost of creating the augmented container is arbitrarily close to cost of simply fulfilling the original request.
+Since the augmented projection is the union of two different projections,
+it can be used for jobs requiring either.
+Thus on merging, the augmented projection can replace the existing image,
+so that the net increase in storage space comes only from the newly added files.
+Since a single augmented image now serves two different batch job submissions,
+we reduce duplication in the cache by avoiding making two copies of the common files and data.
+
+One potential issue with this merging strategy is that there is no way to prevent poor merging decisions.
+In the case of a one-off submission,
+the system might merge an existing, frequently used image.
+This is not problematic for the one-off job itself,
+but increases overhead for subsequent jobs.
+The original, frequently used image would increase in size with no additional benefit to other tasks.
+Images might become ``bloated'' in this way,
+accumulating infrequently used dependencies and increasing overhead indefinitely for future tasks.
+The Jaccard distance metric gives a natural way to capture and quantify this effect.
+As an image becomes bloated due to repeated merges,
+its distance from every individual request increases.
+After sufficient growth,
+the image will become too far from any request to even be considered.
+Without regular use,
+the bloated image will eventually be removed in LRU order.
+Choice of $\alpha$ allows an upper limit on the amount of useless bloat in images.
+
+\subsection{Simulating Projections}
+
+\if 0
+Based on correspondence with researchers,
+we made some general observations about the behaviors and requirements of applications we need to support.
+First, there are some components such as ROOT that serve as common, foundational components to almost every other simulation and analysis stack.
+These core components are often stored separately from the rest of the software collections available in CVMFS.
+Second, there are shared components such as setup scripts, configuration files, and calibration or condition data (required to interpret readings from detectors) that are also nearly universally required but are more particular to a given experiment or group.
+Finally, there is the collection of software maintained by each experiment.
+Software collections may contain multiple copies of the same software for different architectures and versions.
+A single analysis job does not use the majority of the available software.
+There is some overlap among these categories of software,
+for example some software components are widely enough used to be considered core components,
+and usage and availability of components varies over time.
+\fi
+
+    Since there are currently few regular CVMFS on HPC users,
+we decided to simulate the activity of users requesting container images.
+We decided to simulate at the granularity of packages rather than individual files,
+because this results in more robust container images.
+As mentioned previously,
+fine-grained tracing runs the risk of leaving out important dependencies.
+In addition, working only based on the properties of individual files ignores the structure inherent in software collections.
+Looking only at the properties of individual files,
+Figure~\ref{fig:cdf} shows cumulative distributions for file sizes in several of the largest and most active CVMFS repository.
+The repository vary in scale,
+but all show similar usage patterns.
+We see that across all repositories,
+the majority of files are small (between 1~KB and 1~MB).
+The repositories do contain, however,
+a small number of large files.
+This fits our expectation that the CVMFS repositories primarily contain software.
+Most individual files are small,
+but the repository size is defined by the large number of packages in the software collection.
+
+We tested two schemes for generating image requests, taking into account the properties of the software collections,
+as well as recorded user activity patterns.
+We also generated images via a uniform random scheme for comparison.
+We consulted with the developers of CVMFS,
+as well as HEP researchers at the University of Notre Dame collaborating with CERN to determine how current users interact with CVMFS.
+Since CVMFS is not currently widely deployed on HPC resources,
+the prospective user base for our container management system is limited.
+We were, however, informed of some researchers and groups experimenting with static mirrors of CVMFS for container images,
+including projects at NERSC and Stanford University,
+with one of the setups available publicly at \url{https://github.com/wyang007/atlas-fat-container}.
+
+ We use these observations about the structure and behavior of applications on CVMFS to build a simplified simulation taking into account dependency requirements for evaluating our container image management scheme.
+For the purpose of simulation,
+we assume that the requirements of a job are given as discrete components or collections of files.
+While Shrinkwrap can operate at the granularity of individual files,
+both static analysis and runtime tracing produce file collections to include.
+We are intentionally somewhat vague here to allow for both explicitly defined software components and tracing outputs.
+We examined the software collections available in the SFT repository and constructed a DAG of dependencies among the packages.
+The currently available software collection in the SFT repository consists of 9660~packages,
+where a program or library typically provides packages for multiple versions and configurations.
+When building a projection,
+we recursively include dependencies of requested software.
+This more closely approximates the structures of actual container images,
+while still allowing for randomly selected package requests.
+For each simulated image request,
+we chose a random selection of packages (either uniform random selection or subject to the process below)
+and then added the closure of the package dependencies.
+This first image simulation scheme captures the structure inherent in the software collection,
+in that packages in addition to those requested are automatically included so as to ensure a functional image.
+The initial selection of packages,
+however, is simply uniformly random.
+This is not necessarily realistic
+so we also took usage patterns into account.
+
+We assumed that certain components are included in most container images,
+while other components are relatively rarely used.
+While multiple versions and variations might exist,
+these components have a very high likelihood of appearing in every container image.
+These components correspond to the core components, setup scripts, calibration data, etc.\ described previously.
+For the purposes of simulation,
+we would like to see these components in a large proportion of jobs across all users and experiments.
+
+We obtained CVMFS usage logs from frontend nodes at CERN,
+which give rough usage frequencies of the packages in the SFT repository.
+To generate simulated container requests,
+we made random package selections subject to the observed distribution of package use.
+Many packages in the repository were not observed being used at all,
+so we added a small, non-zero probability of selecting packages lacking usage data.
+We applied this technique in combination with the package dependency graph,
+so that images contained a weighted random core selection along with any dependent packages.
+This second image simulation scheme takes into account the dependency structure of the repository as well as usage patterns observed in the wild.
+
+As a base for comparison,
+we also generated simulated images consisting of packages chosen in a uniform random way.
+To ensure that total size (or at least total number of packages)
+is comparable to images generated by the previous two methods,
+for this approach we started with an image request generated via the first scheme discussed
+(uniform random core selection with dependencies added).
+We considered only the number of software packages in the resulting image,
+and then chose the same number of packages uniformly randomly from the entire repository,
+ignoring usage information and package dependencies.
+while images generated in this way are not particularly realistic,
+they will be useful for evaluating image management schemes.
+By comparing results with random images to those with the previous two image generation schemes,
+we can compare the general case of containers as collections of arbitrary data to the specific focus of this work,
+i.e.\ containers with selections of experimental software.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/time-series/time-series.pdf}
+\caption{Behavior of a Single Simulation}
+\label{fig:series}
+\if 0
+zoomed out?
+cache limit, deletes start
+merge tracks writes, it's the dominant factor
+hits grow gradually as cache merges and fills
+deletes help prune useless stuff, remaining should be frequent hits
+far ends are either insert or merge, middle alphas have both
+\fi
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/comparative/operation_count.pdf}
+\caption{Operation Counts for a Single Cache Configuration}
+\label{fig:opcounts}
+\if 0
+inserts and merges switch as alpha goes up
+more hits with merging
+allows better responsiveness? image is immediately ready
+inserts and deletes go to zero, so storage use will taper off
+drop off: we keep everything, so 500 requests get merged; hit count goes up correspondingly since everything is in cache
+not gonna be close to cache limit
+inserts and deletes track: corresponds to LRU, inserts while filling then balanced
+\fi
+\end{figure}
+
+    Figure~\ref{fig:series} shows a single simulation of our image management system
+with $\alpha=0.0.75$ and cache size of 1.4~TB processing 500 unique job specifications, each one repeated five times.
+First, we note that most of the operations are merges.
+This is to be expected,
+as this simulation ran with a fairly high $\alpha$ value.
+The total bytes written also closely tracks merges,
+indicating that merging is the dominant source of I/O.
+We can still see inserts over the course of the simulation.
+At more extreme $\alpha$ values,
+we would expect to see one of these operations almost exclusively.
+As the data in cache continues to rise,
+we eventually see the cache limit,
+after which the delete count starts increasing.
+Over the course of the simulation,
+inserts and deletes are filling and emptying the cache such that it remains close to its storage limit.
+We also observe the number of cache hits continue to rise despite deletions.
+As we will see, merging allows for a greater proportion of hits even if the amount of data remains constant.
+This is because frequently used data can be merged,
+reducing duplication.
+The cache limit then ensures that infrequently used parts are eventually removed.
+
+\section{Evaluation}
+\label{sec:eval}
+
+To evaluate the viability of our approach in automatically deduplicating a container image store,
+we simulated the behavior of the system over time for a range of $\alpha$ values.
+Choice of $\alpha$ is important here to ensure that common components can be detected and merged in container images,
+and that old images or poor decisions are eventually removed from the system.
+Choosing $\alpha$ too far in either direction will result in excessive storage overhead due to duplicated components,
+or excessive compute and network overhead repeatedly merging largely unrelated images.
+Our goal in this evaluation is therefore to choose $\alpha$ so as to minimize the storage and compute costs associated with maintaining a collection of images.
+
+    %sweeping over alpha
+
+    Since our simulation uses random simulated requests,
+there is noticable variability between individual simulations.
+Thus for a given choice of cache size, job count, etc.
+we repeated the simulation 20 times and reported the median behavior over the runs.
+At each choice of $\alpha$ (in steps of 0.05)
+we performed a set of 20 simulated runs,
+allowing us to plot various measurements of the system versus $\alpha$.
+
+Sweeping over the range of $\alpha$ values in this way,
+we can immediately see differences in the frequency of simulated operations.
+Figure~\ref{fig:opcounts} shows the upper range of $\alpha$ values where behavior differences appear.
+From the lower values on the left,
+the insert and delete counts are the primary (or only) operations,
+with number of hits relatively constant.
+This corresponds to a simple LRU-based cache.
+The insert count is higher due to cache filling,
+but the two tend to increase in lockstep.
+As $\alpha$ increases,
+image merges become more probable.
+The merge count in the figure steadily increases throughout most of the upper range,
+while inserts and deletes decrease.
+This suggests that at high $\alpha$ values,
+the cache space would be more efficiently used,
+with some of the duplication merged out.
+In the extreme case with $\alpha=1$,
+every request is merged into a single image in cache,
+hence the sudden increase in hit rate and decrease in total merges at the far right of Figure~\ref{fig:opcounts}.
+Note that as merging increases,
+the proportion of hits generally also increases.
+This indicates that more of the requests can be satisfied by previously merged images in the cache.
+From the perspective of the system,
+this is beneficial as there is no extra time or compute cost for handling a larger proportion of images.
+
+\subsection{Measuring Efficiency of the System}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/comparative/distribution_efficiency.pdf}
+\caption{Cache and Container Efficiency for a Representative Configuration}
+\label{fig:dist-eff}
+\if 0
+explain what happened
+metrics: why we care about efficiency?
+cache efficiency: gauge duplication (waste), keeping working set costs lots of storage
+container efficiency: unexpected overhead for each request; compute, network, etc.
+place limits based on site, conditions
+limits exist, we won't tell you
+maybe just buy more cache?
+large scale (10-100 TB), no can do...so need to merge sometimes
+why 100 and 20 on the left
+why they switch places: extreme cases
+strike a balance
+one curve might be more important for you
+effects are not uniform/balanced gentle drop, massive spike in overhead on right
+\fi
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/comparative/cache_efficiency.pdf}
+\caption{Duplication of Data in Cache}
+\label{fig:cache-eff}
+\if 0
+justification for cache efficiency
+unique vs total
+full x range?
+merging can give more unique data in less cache space
+\fi
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/comparative/container_efficiency.pdf}
+\caption{Cumulative I/O Overhead}
+\label{fig:container-eff}
+\if 0
+related to container efficiency: not exact but shows it
+too much jitter from individual requests, so cumulative
+crossover: hit rate allows reuse, so slightly less bytes written than requested
+requested is based on random stream, constant
+actual written goes up fast with alpha, so more overhead
+proxy for compute/network/etc.
+really steep at high alpha
+\fi
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{curated/comparative/distribution_efficiency_comp.pdf}
+\caption{Efficiency Across Package Selection Methods}
+\label{fig:comp-eff}
+\if 0
+different data patterns cause different effects
+random has high utilization
+pushed right, limited opportunity for merge, since no overlap
+merging doesn't help
+random is not realistic though
+merging is not general solution, but for trees of software, look at this
+
+tree has more duplication, starts merging much sooner
+closure is very non-controvertial, but results in serious changes in efficiency
+common to any software, so opportunity for optimization
+common core that can merge to avoid dups
+including usage stats (dist) amplifies these somewhat
+real usage should be stronger than dist
+
+use dist from now on because reasonable approximation to actual usage
+similar to tree anyway
+mild but noticable difference from tree, more natural approximation to real usage
+\fi
+\end{figure}
+
+When sweeping over the range of $\alpha$ values,
+there are a number of metrics available to summarize each simulation run.
+Many, however, are highly coupled with the particular system configuration and difficult to compare as we vary the paramaters of the simulations.
+
+We therefore chose to define two metrics, cache efficiency and container efficiency,
+to indicate the relative utilization of the system.
+Figure~\ref{fig:dist-eff} plots these two metrics for a representative simulation.
+We defined cache efficiency as the ratio of unique data to total data in the cache.
+In our case, this is equivalent to the ratio of the size of the unique packages to the total cache size.
+If many images contain copies of the same packages,
+the cache efficiency decreases.
+This metric captures the extent of duplication within the cache across all images.
+With no merging there is a high degree of duplication,
+so the cache efficiency is low.
+On the other end of the spectrum,
+maintaining a single, large image containing all data results in cache efficiency of 100\%,
+as the entire cache resides in that single image with no duplication
+
+We defined container efficiency as the ratio of the size of the requested container
+(in our case a set of requested packages plus all dependencies)
+to the size of the container the system actually used for the job.
+In the absence of merging,
+these two are equal so the container efficiency is 100\%;
+jobs are run with exactly what was requested.
+By merging to allow for image reuse,
+we include additional, unrequested data in container images.
+The container efficiency measures this difference between requests and containers.
+In the extreme case of $\alpha=1$ with a single large image, for example,
+the container efficiency is poor because the entire repository is used for every request,
+regardless of size.
+
+These two extreme cases, no merging among many images and a single merged image,
+can both be useful in some situations.
+If the requirements of all jobs are known in advance and worker storage and bandwidth are sufficient,
+maintaining a single, large image for all jobs on all workers is an useful strategy.
+After the initial (costly) image creation and transfer,
+there are no continuing costs or delays.
+When resources are limited or requirements change regularly,
+however, this approach becomes prohibitively expensive.
+The high compute and bandwidth overhead apply for every image update,
+which in the worst case could be every job.
+
+The other extreme situation,
+simply caching requests with no merging,
+can also be viable.
+This approach is simpler and does not come with extreme per-job overhead.
+In addition, frequent changes to the jobs and requirements do not degrade the efficiency of each job.
+At large scale, however,
+the overall system efficiency suffers.
+Due to duplication among images,
+larger caches contain less and less unique data.
+To support a given repository,
+it can become necessary to provision a cache much larger than the size of the repository.
+With common CVMFS repositories consuming several terabytes of storage each,
+the amount of cache space grows quickly.
+
+These two limiting factors,
+the compute and transfer cost in the highly merged case and the cache efficiency in the unmerged case,
+serve as limits on the viable range of $\alpha$ values for the system.
+Figure~\ref{fig:dist-eff} shows two vertical lines indicating these limits on a plot of efficiencies.
+The left line shows a lower limit of around 30\% on the cache efficiency.
+Choosing $\alpha$ too low results in cache efficiency reaching below 20\%.
+On the right, the line gives an upper limit on the likelihood of merging.
+As shown in Figure~\ref{fig:opcounts},
+the amount of I/O and compute to update images becomes much larger if $\alpha$ is set too high.
+These two limits define a range of viable values for $\alpha$.
+There is no general rule for the placement of these limits,
+which depends strongly on the performance characteristics of the execution environment,
+as well as the priorities of the administrators in optimizing the system.
+Rather than defining where these limits fall,
+we give per-site considerations in choosing limits in Section~\ref{ANALYSIS_SECTION}.
+
+Figure~\ref{fig:cache-eff} shows the actual data sizes used to calculate the cache efficiency.
+Without merging, the unique data makes up a small proportion of the cache.
+With increased merging, the amount of unique data increases.
+For sufficiently high $\alpha$,
+merges occur more frequently than cache inserts,
+resulting a drop in total storage size.
+On the far right of Figure~\ref{fig:cache-eff},
+the entire cache contents are merged into a single large image,
+so that the unique and total data sizes become equal.
+The cache efficiency metric is affected both by the increase in unique data and by the decrease in total cache size.
+
+Figure~\ref{fig:container-eff} shows the cumulative costs of creating and updating images relative to the cost of simply providing the requested images.
+These costs illustrate the effect of container efficiency over the course of a simulated run.
+Without merging, the actual I/O and compute costs in the system closely follow the requests.
+In Figure~\ref{fig:container-eff} the actual I/O is slightly smaller than the requested amount due to caching:
+the system can sometimes reuse images without performing any I/O.
+As $\alpha$ increases, the cost of updating and merging images comes to dominate the total cost.
+Administrators migh decide, for example,
+to spend at most twice the compute and I/O time relative to the the actual requested operations.
+This would correspond to the upper compute limit shown on \Cref{fig:dist-eff}.
+Depending on the priorities at a particular execution site,
+either the storage utilization or compute and I/O time could be the critical factor in optimizing performance.
+
+\begin{figure*}[ht]
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_cache.pdf}
+\hfill
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_cache.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_request.pdf}
+\hfill
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_request.pdf}
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_uses.pdf}
+\hfill
+\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_uses.pdf}
+\caption{Effects of Simulation Parameters on System Efficiency}
+\label{fig:sensitivity}
+\if 0
+[t] "just buy more cache"
+efficiency is worse, since cache holds on to lots of questionable data
+loads of opportunities to try merges, so we see a lot of unnecessary merges
+less force removing bad choices, so more cruft can accumulate in many images
+better cache utilization with tiny cache, but of course more thrashing
+at scale, bad idea
+large cache is too pricey at scale
+[m] varying reuse doesn't affect too much
+below 100 doesn't seem full/steady state yet
+cache isn't settled? so haven't dropped bad merges; more unique data in cache boosts efficiency
+[b] high number of requests, repeats don't matter
+unique data dominates
+hard to tell difference between [m] and [b], so not too sensitive to these
+once you hit steady state, jobs count/repeats has little effect on efficiency
+general trends of merging, not interesting by themselves
+\fi
+\end{figure*}
+
+\subsection{Comparison of Simulated Image Types}
+
+We also compared the efficiencies of several types simulated images under a range of $\alpha$ values.
+For the previous discussion, synthetic images were generated taking into account both the tree of package dependencies and usage statistics.
+Figure~\ref{fig:comp-eff} shows a representative simulation with two other synthetic image types included.
+First, we measured images considering only the dependency relationships but not usage.
+These images show similar behavior to those previously discussed,
+albeit with slightly lower efficiencies.
+More realistic images primarily make the same trends more pronounced.
+This suggests that even with random, non-realistic requests,
+the tree structure of package dependencies is sufficient to produce duplication in the cache.
+The other type of synthetic image shown in Figure~\ref{fig:comp-eff} is purely randomly generated.
+The set of packages in each image was chosen uniformly randomly from the full SFT repository.
+These random images do not take into account usage statistics or package dependencies,
+making the software in them entirely nonfunctional.
+For the random images,
+we observe markedly different cache behavior.
+The cache efficiency is much higher,
+indicating a greater proportion of unique data in the cache.
+Since the image contents were selected purely at random,
+this is expected.
+The container efficiency is also high over a much larger range of $\alpha$ values.
+This is because there is effectively no merging until extremely high $\alpha$.
+Since image contents are purely random,
+there is no correlation between different images.
+Thus, it is much more difficult to find images similar enough to merge until the $\alpha$ value is very lax.
+This would indicate that our merging strategy is not applicable to arbitrary collections of data.
+Random images show little to no effect for most $\alpha$ values.
+Our merging strategy, which takes advantage of duplicated content included as a result of dependencies in software,
+would be ill advised for situations that are not known to follow similar patterns of duplication.
+
+\subsection{Sensitivity Analysis}
+\label{sec:sensitivity}
+
+    In Figure~\Cref{fig:sensitivity}, we plot efficiency curves for a range of simulation conditions.
+The left column shows container efficiency,
+while the right column shows cache efficiency.
+In the first row, the number of jobs and the amount of repetition are constant while the cache size is varied.
+In the second row, the number of unique requests is varied with the other parameters constant,
+and the third row varies the number of times each job is submitted with the rest constant.
+
+The size of the cache has an inverse relationship with both the container and cache efficiency.
+A larger cache can of course hold a larger number of images,
+but since each image contains significant duplicated portions,
+increasing cache size tends to decrease cache efficiency.
+Conversely, small caches more quickly evict images so that ineffective merges and similar images tend not to remain in cache too long.
+A larger cache also allows for more opportunities to merge images,
+leading to decreased container efficiency.
+When deciding how to handle a request,
+a large cache full of images is much more likely to contain an image suitable for merging.
+With a small cache, opportunities to merge are much more dependent on the order in which requests happen to come.
+
+The effect of varying the number of unique jobs is less pronounced than the effect of cache size.
+Streams of 500 and 1000 unique jobs show nearly indistinguishable behavior,
+indicating that by 500 jobs the system has reached a steady state.
+Continuing with an arbitrarily long stream should not result in significant performance changes.
+It appears that 100 unique jobs were not sufficient to fill the cache and reach a steady state.
+In this case the container efficiency is slightly decreased over $\alpha$,
+suggesting that some ineffective merges had not made their way out of the cache yet.
+The cache efficiency in this case is slightly increased.
+This would suggest that before reaching a steady state,
+the cache contents are more assorted and some unnecessary data remains cached.
+
+The effect of varying the number of times each unique job is submitted appear minimal at best.
+As discussed with the previous row,
+a stream of 500 unique images should be sufficient to reach a steady state.
+For a sufficiently large cache, therefore,
+it appears that the efficiency is determined by the amount of unique data and fairly insensitive to how much those pieces of data are used.
+This simulation is somewhat unrealistic in that each job is repeated a uniform number of times.
+In real systems, we expect more variation in the frequency of repeated jobs.
+We predict that a steady, non-uniform distribution of frequencies would result in the same behavior as shown in the bottom row of Figure~\ref{fig:sensitivity},
+while changes in the frequencies of jobs (e.g. a new user starting an experiment)
+would disturb the steady state and produce effects similar to those in the middle row.
+
+\subsection{Choosing $\alpha$ Threshold}
+\label{ANALYSIS_SECTION}
+
+Based on the previous analyses of simulated cache behavior,
+some trends emerge in the effect of choice of $\alpha$ with varying cache parameters.
+While there is no general rule for applying our approach to real sites,
+these trends give some guidance.
+The absolute simplest method to improve caching behavior is not to merge images and to increase the size of the cache.
+For a small number of users or infrequent changes to containers,
+this might be all that is needed.
+At the scale under consideration here, however,
+this somewhat tongue-in-cheek advice easily becomes unfeasible.
+As shown in \Cref{fig:sensitivity},
+the cache efficiency without merging often falls below 20\%.
+When working with one or more repositories containing several terabytes of data each,
+provisioning a cache large enough to hold highly duplicated images becomes expensive.
+This storage cost gives a lower limit in choosing $\alpha$.
+Rather than committing tens of terabytes to a cache,
+sites may prefer to devote a larger proportion of compute and network bandwidth to image management.
+Since the compute overhead grows very quickly at high $\alpha$,
+it would be ill advised to push this value too high.
+The safer approach to choosing $\alpha$ is to start at the lower limit and gradually increase $\alpha$,
+observing the changes in compute overhead.
+
+%%%%%%%%%%%%%%%%%%%%
+\if 0
+where software is provided as needed in the form of RPM files or another similar format.
+Working with software packages is simpler in that dependencies are explicitly specified in the package file or database.
+With CVMFS, no such definitive specification exists.
+This issue is common to shared/global filesystems,
+since users expect the complete contents to be available on demand and so devote little time to explicitly defining dependencies and other requirements.
+
 CVMFS differs from traditional POSIX filesystems in that previous versions of all content remain available.
@@
 and providing this information for verification.
+\fi
+
+\section{Related Work}
+\subsection{Container Technology}
+A core assumption presented in this paper is that 
+container technology is relied on for packaging and
+transporting the projections to the compute site.
+Unfortunately, which of a variety of solution is used 
+depends on the actual site.
+Currently, the most popular solutions are
+Docker\cite{Merkel:2014:DLL:2600239.2600241} and
+Singularity\cite{Singularity}.
+Docker is more ubiquitous, but its 
+obfuscated implementation and privileged user requirements 
+are leading more scientist and site administrator
+to Singularity.
+Additionally, site often spin-off highly tailored implementation
+like Shifter\cite{Jacobsen2015ContainT} at NERSC
+and Slacker\cite{194430}, which provide slim containers
+for moving around that bootstrap larger systems.
+
+In recent years, with the announcement of user namespaces
+and unprivileged filesystem mounts~\cite{user-filesys-mount}, notably FUSE, 
+containers, image formats, and different environment specification
+use will increase and proliferate increasing the need for a container
+agnostic approach until more consistency is reached.
+This will likely lead to more simplified container implements such as
+CharlieCloud\cite{Priedhorsky:2017:CUC:3126908.3126925}.
+
+\section{Conclusion}
+
+Global filesystems offer a convenient way to organize and distribute the large, complex, and frequently updated software repositories that support international scientific experiment collaborations such as ATLAS, CMS, LIGO, and LSST.
+The software in these repositories consists of millions of files and terabytes of data.
+The massive scale of these projects would not be possible without software such as CVMFS to ensure consistency and availability of software at sites across the globe.
+At some computing facilities, especially national HPC sites,
+restrictions on the worker activity and network access make global filesystem access difficult or impossible.
+To take advantage of local high performance filesystems and container technologies available at such sites,
+we introduced the concept of projecting subsets of the global filesystem into local resources.
+We instrumented CVMFS to add support for tracing application dependencies,
+and developed a tool for efficiently projecting CVMFS into local container images suitable for use at HPC facilities.
+When working with many users and at large scale, however,
+the straightforward approaches to managing these projected images fall short.
+We proposed a simple strategy for managing a cache of projected images,
+taking advantage of usage patterns and duplication of software collections to reduce data duplication among containers.
+Our approach allows users or system administrators to balance the storage cost associated with small, tailored container images against the computational and bandwidth costs of maintaining large, multi-job images.
+Using our cache management approach,
+applications can continue to take advantage of precise dependency requirements,
+while users and administrators can fit performance and storage costs to meet the requirements of their particular execution sites.
 
 \if 0
-\begin{table*}[t]
-\begin{center}
-\begin{tabular}{|c|r r r|r r r|}
-\hline
-CVMFS & & ls -lr & & & Shrinkwrap & \\
-Repository & Total Size (MB) & inode Count & Time to walk & 
-  Data Files & Deduplicated Data Files & Deduplicated Data (MB) \\ \hline
-%alice-ocdb & 482,642,307,428 & 1,621,661 & 2:22 &
-% 1,621,282 & 1,595 & 24,433,131,203\\
-%alice & 3,980,875,747,986 & 27,322,350 & 51:53 & 
-%  797,404 & 23,996,541 & 2,098,509,015,172\\
-alice & 3,980,875 & 27,322,000 & 51:53 & 
-  797,000 & 23,997,000 & 2,098,509\\
-%atlas-condb & 719,548,941,240 & 11,781 & 0:04 &
-%  10,940 & 229 & 471,492,086 \\
-%atlas & 6,195,013,689,056 & 120,598,601 & 3:48:07 &
-% 8,823,179 & 85,554,444 & 3,760,712,059,548 \\
-atlas & 6,195,013 & 120,599,000 & 3:48:07 &
- 8,823,000 & 85,554,000 & 3,760,712 \\
-%boss & 120,048,527,315 & 2,418,025 & 4:18 &
-% 256,061 & 1,868,192 & 72,301,955,666\\
-boss & 120,048 & 2,418,000 & 4:18 &
- 256,000 & 1,868,000 & 72,302\\
-%cms & 581,789,406,847 & 26,224,567 & 1:07:30 &
-% 18,667,178 & 101,448,867 & *2,242,658,825,701\\
-cms & 581,789 & 26,225,000 & 1:07:30 &
- 18,667,000 & 101,449,000 & *2,242,658\\
-%geant4 & 297,571,283,648 & 7,128,024 & 10:36 &
-% 323,280 & 6,219,891 & 68,381,969,660 \\
-geant4 & 297,571 & 7,128,000 & 10:36 &
- 323,000 & 6,220,000 & 68,382 \\
-%grid & 30,642,826,259 & 714,235 & 1:17 &
-% 141,622 & 583,923 & 18,804,003,040 \\
-grid & 30,643 & 714,000 & 1:17 &
- 142,000 & 584,000 & 18,804 \\
-%lhcb & 2,134,893,079,440 & 33,736,802 & 1:01:26 &
-% 10,032,339 & 20,406,235 & 368,070,277,634 \\
-lhcb & 2,134,893 & 33,737,000 & 1:01:26 &
- 10,032,000 & 20,406,000 & 368,070 \\
-%na61 & 5,222,967,734 & 25,602 & 0:02 &
-% 8,609 & 15,001 & 2,425,507,431 \\
-na61 & 5,222 & 26,000 & 0:02 &
- 9,000 & 15,000 & 2,426 \\
-%sft & 3,996,527,275,237 & 128,608,792 & 4:17:58 &
-% 6,821,699 & 63,210,139 & 2,016,609,810,573 \\ 
-sft & 3,996,527 & 128,609,000 & 4:17:58 &
- 6,822,000 & 63,210,000 & 2,016,610 \\ 
- \hline
-\end{tabular}
-\caption{CVMFS repository sizes.}
-\label{tab:repo-sizes}
-This table shows the size, number of inodes (rounded to nearest thousand), and the amount
-of duplicate data that is eliminated using CVMFS.
-\end{center}
-\end{table*}
-
-\fi
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|c|r r r|}
-\hline
-Repository & Total Size & inode Count & Deduplicated \\
- & (GB) & (K) & Data (GB) \\ \hline
-alice & 3,981 & 27,322 & 2,099 \\
-atlas & 6,195 & 120,599 & 3,761 \\
-boss & 120 & 2,418 & 72\\
-cms & 582 & 26,225 & *2,243\\
-geant4 & 298 & 7,128 & 68 \\
-grid & 31 & 714 & 19 \\
-lhcb & 2,135 & 33,737 & 368,070 \\
-na61 & 5 & 26 & 2 \\
-sft & 3,997 & 128,609 & 2,017 \\ 
- \hline
-\end{tabular}
-\caption{CVMFS repository sizes.}
-\label{tab:repo-sizes}
-This table shows the size, number of inodes (rounded to nearest thousand), and the amount
-of duplicate data that is eliminated using CVMFS.
-\end{center}
-\end{table}
-
-\section{Realization of CVMFS Projection}
-
-\subsection{Definition of Projection}
-
-In the case of global filesystems,
-every time the filesystem is used locally
-the flexible global state has to be 
-locked for that concrete use case.
-In the case of CVMFS,
-this includes specifying the date-time
-or revision number of the repository.
-This state is used to determine 
-what the global filesystem's directory
-structure and file contents are, 
-allowing for future use of the same state.
-CVMFS is able to flexibly move
-between these revisions using the FUSE module
-to determine the structure and files and
-relies on HTTP to pull any missing files.
-However, with neither the HTTP access or
-FUSE permissions, a method is needed to 
-create and track realizations of this
-state for future use.
-
-We call these realizations of the global filesystem
-projections.
-A projection is a concrete filesystem image that can be
-mounted, loaded, or copied into an existing system.
-Projections can then be used to run experiments
-and analysis using the global filesystem without
-the previously mentioned permissions.
-Projections have been used with CVMFS before,
-in the form of unCVMFS and rsync.
-UnCVMFS is a utility that projects an entire
-CVMFS repository, at a set revision, into the local
-filesystem. 
-UnCVMFS utilizes CVMFS's data/file separation
-to deduplicate the underlying data, but requires
-entire repositories to be projected.
-When used with several repositories for large projects,
-unCVMFS can easily consume 500GB for single projection.
-As is often the case, different experiments/analyses 
-require different revisions that may conflict.
-
-Rsync allows for file lists to specify and download
-subsets of CMVFS repositories, but because it
-was designed for general filesystem synchronization
-rsync flattens the files, introducing duplicate data.
-This direct representation of the file tree also
-excludes the possibility of shared data repositories
-between different projections.
-
-% Probably need to mention that many applications assume
-% CVMFS software is installed/mounted at /cvmfs
-
-\begin{figure}[h]
-\includegraphics[width=\columnwidth]{drawings/shrinkwrap-structure.png}
-\caption{Shrinkwrap Approach}
-\label{figure:shrinkwrap-arch}
-\end{figure}
-
-We built the CVMFS Shrinkwrap utility,
-a solution that leverages the best of both
-unCVMFS and rsync.
-Shrinkwrap uses the CVMFS's client interface to
-create and update CVMFS projections.
-Shrinkwrap uses a CVMFS configuration to specify the
-revision of the repository, a file-specification list
-to determine the repository subset to project, and
-deduplicates the data in a user specified location.
-Shrinkwrap can be configured to share the data
-repository between several repositories and even
-projections of the same repositories.
-This allows for multiple realized projections
-to coexist, with the desired projection
-mounted for use.
-In \Cref{figure:shrinkwrap-arch} you can see an outline
-of the design of the Shrinkwrap utility.
-
-\subsection{Packaging Projections}
-
-All of the above methods described utilize CVMFS
-and rely on the internet connection to create
-the projections.
-A core assumption is that external network
-access is limited at the compute nodes, so a method
-is needed to transform and use these projections
-for computation.
-For this several methods for providing
-these projections using 
-Container images (images specific to Docker, Singularity, ect.)
-or SquashFS images can be used.
-The goal is to find a flexible, compatible intermediate
-format that allows projections to be moved between sites
-and technologies.
-For this purpose, Docker images, Singularity images, and 
-SqaushFS are interchangeable as outputs of the Shrinkwrap process
-and changed as needed for each use case as seen below.
-
-%are compared on their time to create,
-%storage requirements, and most importantly compatibility
-%with several used technologies, namely 
-%Docker, Singularity, and Shifter.
-
-The first aspect is the creation of these images.
-From a defined projection of a repository
-what are the steps needed to create each of these image types.
-
-\subsubsection{Atlas at NERSC}
-
-As previously described, NERSC is working in collaboration the LHC ATLAS project
-to provide HPC resources for simulation and experimentation.
-NERSC is currently deploying CVMFS images as monolithic container 
-built using UnCVMFS, packaged in Squashfs images, and then containerized 
-into Shifter or Singularity format, whose final format
-relies on the target site.
-This entire process can take up to a day to complete with the majority
-of compute used to pull down and store the full CVMFS copy and storage
-needs in excess of terabytes for images.
-By combining either traced samples or statically defined requirements for 
-jobs, Shrinkwrap can reduce these images orders of magnitudes.
-Shrinkwrap which relies on the libcvmfs library reduces 
-the total overhead and eliminates unnecessary files transfers
-over UnCVMFS by allowing selections and rsync using data deduplication.
-
-\subsubsection{Benchmarking Containers}
-For benchmarking purposes, the CERN IT-Department is interested in a ``Benchmark in a Container'' solution which allows the distribution of standalone container images for system performance evaluations.
-For such benchmarks the IT-Department typically relies on simulation workflows provided by CERN's experiment collaborations.
-To execute these benchmarks in an easily portable and reproducible manner, it is of high interest to package these software workflows into containers.
-Since the workflows' software stacks are usually stored inside CVMFS repositories, this is a perfect use case for the shrinkwrap utility and its Docker injection functionality.
-
-For the image export the workflow’s resource-needs would be examined through the CVMFS tracing feature in a first step.
-This would then result in a log containing detailed insights on the paths which were accessed by the benchmarking workload.
-In a second step we can then produce a specification based on the tracing log which will allow an easy, efficient export of the necessary files from CVMFS.
-Afterwards the Docker injector can be used to inject the exported files into a prepared benchmarking container stored in a registry.
-Later, the same utility can be used to update the current container to an up-to-date software version or — depending on the exact use case — the utility could even be integrated into a continuous deployment mechanism which produces new benchmarking containers on every software update.
-
-\section{Managing Projections as a Site-wide Resource}
-
-With static analysis and dynamic tracing of tasks,
-we have methods to choose subsets of CVMFS to project to run specific tasks.
-Users can collect this information themselves,
-or tracing, too, could be handled automatically.
-We have seen two approaches to this in section ??.
-For applications already running on the WLCG,
-it is also possible to instrument a fraction of runs to build a profile of the type of task.
-Alternatively, we could integrate tracing with image creation by automatically running a small number of sample tasks on the WLCG or on dedicated bridge nodes and collecting traces as necessary.
-These approaches give a fair amount of flexibility in determining accurate dependency information.
-Based on this information,
-we are interested in automatically providing suitable projections of CVMFS to users.
-
-Shrinkwrap provides a mechanism for projecting parts of CVMFS,
-but is not by itself a complete solution as it does not integrate with job submission or manage images after creation.
-In the case of a small number of manually created container images,
-this is not a serious problem.
-Administrators could manage a small collection of images in this case and adjust jobs to work with their setup.
-Thus for integrating with the existing systems at NERSC,
-Shrinkwrap alone might be sufficient.
-To support CVMFS as a site-wide service with a potentially varied user base and workload,
-a more automated solution
-(e.g.\ a batch system plugin that can automatically prepare container images on job submission)
-is required.
-This both minimizes direct administrator burden and more closely approximates user experience on other resources like the WLCG.
-Our first attempt at such a system,
-described in the next sections,
-took the simplest possible approach to image.
-As we were developing this approach, however,
-a number of issues in real-world usage led us to conculde that a somewhat more sophisticated solution was required.
-
-\subsection{Simulating Job Submissions}
-
-Since there are currently few regular CVMFS on HPC users,
-we decided to simulate the activity of users requesting container images.
-We decided to simulate at the granularity of packages rather than individual files,
-because this results in more robust container images.
-As mentioned previously,
-fine-grained tracing runs the risk of leaving out important dependencies.
-In addition, working only based on the properties of individual files ignores the structure inherent in software collections.
-Looking only at the properties of individual files,
-Figure~\ref{fig:cdf} shows cumulative distributions for file sizes in several of the largest and most active CVMFS repos.
-The repos vary in scale,
-but all show similar usage patterns.
-We see that across all repos,
-the majority of files are small (between 1~KB and 1~MB).
-The repos do contain, however,
-a small number of large files.
-This fits our expectation that the CVMFS repos primarily contain software.
-Most individual files are small,
-but the repo size is defined by the large number of packages in the software collection.
-
-We tested two schemes for generating image requests taking into account the properties of the software collections,
-as well as recorded user activity patterns.
-We also generated images via a uniform random scheme for comparison.
-We consulted with the developers of CVMFS,
-as well as HEP researchers at the University of Notre Dame collaborating with CERN to determine how current users interact with CVMFS.
-Since CVMFS is not currently widely deployed on HPC resources,
-the prospective user base for our container management system is limited.
-We were, however, informed of some researchers and groups experimenting with static mirrors of CVMFS for container images,
-including projects at NERSC and Stanford University,
-with one of the setups available publicly at \url{https://github.com/wyang007/atlas-fat-container}.
-
-To evaluate the effect of including dependencies in container images,
-we made a large number of random selections of packages from the SFT repo and then compared with the set of dependencies required by each selection.
-Figure~\ref{fig:picks} shows the results over a large number of selection sizes.
-We observe of course that selection size increases uniformly along the $x$ axis.
-The on-disk size of the selections appears to increase proportionally to selection size.
-Recursively including the dependencies of these same selections,
-however, results in a significant increase in number of packages and size on disk.
-For small selections (less than 100~packages),
-the complete images with dependencies might contain 5x as many packages as requested,
-with storage increasing proportionally.
-With larger selections, however,
-this increase becomes less prominent.
-This non-linear change is the result of the tree structure of the software dependencies.
-There are a number of core components that are transitive dependencies of a large number of packages.
-These common dependencies are only included once,
-so subsequent selections that depend on them do not add them again.
-As the number of selections increases,
-this curve will approach the total repo size.
-Also note that the non-linear increase in size with increasing selection size means that the size of combining two selections of 100 packages (and their dependencies) is not as simple as adding them together.
-Due to duplication, we would expect the resulting combined size to be lower,
-as if choosing 200 packages from the start.
-
-\begin{figure}
-    \includegraphics[width=\linewidth]{curated/lxplus_dist/usage_distribution_cdf.pdf}
-    \caption{Usage distribution pulled from CERN LXPlus interactive machines.}
-    \label{fig:usage_distribution}
-\end{figure}
-\if 0
-spread across everyone, lot of commonality/overlap
-few with very high usage, very large number of rarely used
-lot not observed
-\fi
-
-Based on correspondence with researchers,
-we made some general observations about the behaviors and requirements of applications we need to support.
-First, there are some components such as ROOT that serve as common, foundational components to almost every other simulation and analysis stack.
-These core components are often stored separately from the rest of the software collections available in CVMFS.
-Second, there are shared components such as setup scripts, configuration files, and calibration or condition data (required to interpret readings from detectors) that are also nearly universally required but are more particular to a given experiment or group.
-Finally, there is the collection of software maintained by each experiment.
-Software collections may contain multiple copies of the same software for different architectures and versions.
-A single analysis job does not use the majority of the available software.
-There is some overlap among these categories of software,
-for example some software components are widely enough used to be considered core components,
-and usage and availability of components varies over time.
-
-We use these observations about the structure and behavior of applications on CVMFS to build a simplified simulation taking into account dependency requirements for evaluating our container image management scheme.
-For the purpose of simulation,
-we assume that the requirements of a job are given as discrete components or collections of files.
-While Shrinkwrap can operate at the granularity of individual files,
-both static analysis and runtime tracing produce file collections to include.
-We are intentionally somewhat vague here to allow for both explicitly defined software components and tracing outputs.
-We examined the software collections available in the SFT repo and constructed a DAG of dependencies among the packages.
-The currently available software collection in the SFT repo consists of NNN packages,
-where a program or library typically provides packages for multiple versions and configurations.
-When building a projection,
-we recursively include dependencies of requested software.
-This more closely approximates the structures of actual container images,
-while still allowing for randomly selected package requests.
-For each simulated image request,
-we chose a random selection of packages (either uniform random selection or subject to the process below)
-and then added the closure of the package dependencies.
-This first image simulation scheme captures the structure inherent in the software collection,
-in that packages in addition to those requested are automatically included so as to ensure a functional image.
-The initial selection of packages,
-however, is simply uniformly random.
-This is not necessarily realistic, however,
-so we next took usage patterns into account.
-
-We assumed that certain components are included in most container images,
-while other components are relatively rarely used.
-While multiple versions and variations might exist,
-these components have a very high likelihood of appearing in every container image.
-These components correspond to the core components, setup scripts, calibration data, etc.\ described previously.
-For the purposes of simulation,
-we would like to see these components in a large proportion of jobs across all users and experiments.
-
-We obtained CVMFS usage logs from frontend nodes at CERN,
-which give rough usage frequencies of the packages in the SFT repo.
-To generate simulated container requests,
-we made random package selections subject to the observed distribution of package use.
-Many packages in the repo were not observed being used at all,
-so we added a small, non-zero probability of selecting packages lacking usage data.
-We applied this technique in combination with the package dependency graph,
-so that images contained a weighted random core selection along with any dependent packages.
-This second image simulation scheme takes into account the dependency structure of the repo as well as usage patterns observed in the wild.
-
-As a base for comparison,
-we also generated simulated images consisting of packages chosen in a uniform random way.
-To ensure that total size (or at least total number of packages)
-is comparable to images generated by the previous two methods,
-for this approach we started with an image request generated via the first scheme discussed
-(uniform random core selection with dependencies added).
-We considered only the number of software packages in the resulting image,
-and then chose the same number of packages uniformly randomly from the entire repo,
-ignoring usage information and package dependencies.
-while images generated in this way are not particularly realistic,
-they will be useful for evaluating image management schemes.
-By comparing results with random images to those with the previous two image generation schemes,
-we can compare the general case of containers as collections of arbitrary data to the specific focus of this work,
-i.e.\ containers with selections of experimental software.
-
-\subsection{An initial approach}
-
-The simplest approach for automatically projecting container images is to create a suitable image on each batch job submission.
-Under this design,
-on job submission the bridge machine first checks if the same projection had been requested before.
-If so, it arranges to transfer its cached container image to worker nodes that require it.
-If the bridge node does not have that projection built already,
-it uses Shrinkwrap to create the requested projection,
-then invokes mksquashfs to create the container image.
-Once the bridge node places the newly created image in its cache,
-the process proceeds as in the previous case.
-The architecture of this initial approach is shown in figure ???.
-
-Unfortunately, this solution is very wasteful of computing time and storage space.
-Even a minimal image consisting of a precise set of required libraries and frameworks ranges in the tens of gigabytes,
-depending on the application.
-The full-repo images being tested at NERSC are multiple terabytes in size and take the majority of a day to create.
-In addition, image creation must be completed before any actual processing occurs,
-i.e.\ there is no way to start with a partial image that supports the first steps of a pipeline and finish building the image as the pipeline runs.
-There is also the cost of transferring and storing freshly created images on any nodes that will use them,
-further delaying job dispatch and increasing load on the network and storage infrastructure.
-
-The disk I/O alone required to create even a targeted image is likely to take more time than a single simulation or analysis job.
-Thus a simple performance improvement is to re-use previously created container images for multiple jobs.
-A single experiment submitting computational tasks is likely to request the same projection of CVMFS,
-at least for jobs submitted around the same time.
-It is straightforward to check for exact equality between projections as long as the system retains the set of files that went into their creation.
-Note that it is important to retain both the paths and the hash of the file contents used in images.
-This ensures that multiple versions of the same file over time are recognized as distinct entities.
-If there are few CVMFS users at a site,
-this method could be sufficient.
-
-\subsection{Problems with the initial approach}
-
-In developing our initial approach,
-we found that a number of real-world usage patterns are poorly handled.
-A small change to a projection,
-such as adding a library or updating dependencies,
-results in the creation of a second, unrelated image.
-Even a small change such as removing an unnecessary documentation file results in a completely new container image.
-This is a manageable nuisance when manually testing or developing a system,
-but potentially results in problematic proliferation of images in a system managing images automatically and receiving updated specifications over time from tracing jobs.
-
-In addition to unnecessary duplication for a single user/application,
-there is also significant duplication across multiple users.
-In the simulation and analysis pipelines at CERN,
-a number of large components such as ROOT are used nearly universally.
-These components provide basic functionality and infrastructure that each experiment's software stack builds upon.
-Moreover, these same common infrastructure components must be included regardless of user or application.
-Thus even if every user of the system only creates a single container image,
-there is a duplicate copy of all such common infrastructure components for each user.
-This is problematic if we would like to design a system that can support many users at a site.
-
-These two sources of duplication tend to compound each other,
-leading us to conclude that a more sophisticated approach is necessary to work with realistically-provisioned workers and system infrastructure.
-These issues are fundamentally related to unnecessary duplication of data among container images.
-The same types of issues motivate the use of dynamic libraries for everyday applications.
-With statically linked executables,
-each application includes a complete set of required libraries.
-While very stable and reliable,
-OS distributions prefer to use dynamic linking.
-This allows applications to share a single copy of each library,
-so that foundational components such as libc are not copied into every executable,
-and so that updates and changes to the set of installed packages do not require rebuilding the entire system.
-For this approach to work,
-package maintainers must arrange all software packages in the system to ensure compatibility,
-and the OS distribution must set policies for system organization, update procedures, etc.
-Container images here function similarly to statically linked executables,
-but since multiple complete OS distributions are stored within CVMFS,
-and since we are dealing with multi-terabyte repos and entire frameworks and software stacks rather than individual OS libraries,
-the scale involved is of course larger.
-In the case of CVMFS, there is no notion of a ``package mainainer'' or central authority over all software.
-Each experiment manages its own software repo as it chooses,
-with some components and repos shared between multiple applications.
-Since we would like to be able to automatically handle existing applications that are not designed with central management in mind,
-we must instead look at other approaches for dealing with data duplication.
-
-\subsection{Deduplicating Container Images}
-
-The representation of projections as disk images is problematic for deduplicating data.
-Internally, Shrinkwrap uses hard links to a common data store so that a single file is only stored once across all projections in the system.
-On creating disk images, however,
-these links are broken.
-Each disk image must contain a complete copy of all data,
-so that a file must be included (and take up space) in every image that includes it.
-Docker's solution is to build container from layers of archives that can be reused.
-Docker layers are explicitly defined when writing a Dockerfile.
-With CVMFS, however, there is no existing layer-like organization or universal rule that could derive layers for existing applicaions.
-Working with layers defined from the outset is also a poor fit if we would like to support partial, gradually improved information from traces.
-We therefore desire some sort of deduplication that can be applied for existing applications to an assortment of container images after creation.
-
-To develop a deduplication strategy,
-we started with some observations about the projection and image properties.
-For ``large'' images,
-there is limited opportunity for deduplication.
-Each large image is a fairly complete copy of a particular repo.
-While there is bound to be some overlap between versions of the same repo,
-an updated image will contain a large number of updated software components,
-making a ``deduplicated'' between two versions significantly larger than either individually.
-There are also likely to be some conflicts in the filesystem if software components re-use the same paths for different files between versions.
-This case is akin to combining RHEL~6 and RHEL~7 in an attempt to save space.
-While there are similarities between the two,
-the significant changes and conflicts make this a poor strategy.
-Thus for our strategy we would like to avoid ``deduplicating'' images that have large non-overlapping portions.
-
-Likewise, there is little reason to attempt to deduplicate large images built from different repos.
-Such images would have little to no overlap,
-so the only benefit to combining the two would be reducing the number of images by one.
-This comes at the cost of needlessly storing and transferring the unused half.
-This case is akin to bundling a complete Debian installation alongside RHEL.
-Our deduplication strategy should only take effect for images with significant overlap.
-
-It is also undesirable to combine images of drastically different size.
-As an example, suppose a user creates a small projection as a subset of a repo for running a specific application.
-This small image overlaps completely with a large image of the whole repo,
-and there is likely no conflict.
-Automatically combining the large and small image, however,
-can have unintended consequences.
-Small, tailored images minimize transfer and storage overhead,
-and can be important for operating at scale.
-It is therefore unwise to surprise users by suddenly making their container images several orders of magnitude larger.
-This is akin to distributing a library exclusively by placing it within a disk image of a Debian installation.
-We would therefore like to attempt deduplication only for
-\begin{itemize}
-\item similarly-sized images.
-\item with large overlap.
-\item and with no conflicting paths.
-\end{itemize}
-
-\subsection{Quantifying Similarity between Projections}
-
-Having set out the conditions under which to consider deduplicating projected images,
-we are in a position to define a metric to quantify the similarity between images.
-This would allow us to identify projections that are ``close'' (for some definition of close) as candidates for optimization.
-We chose the Jaccard distance under appropriate choice of set elements as it conveniently satisfies the desired qualities set out above,
-and is very well used and studied.
-In our case, we must describe a container image as consisting of a set of (path, content hash) tuples.
-CVMFS already stores files by the hashes of their contents,
-so this information is readily available.
-Considering both the file names and contents is important as CVMFS allows for updates to file contents over time.
-We chose fine-grained tracking of file contents rather than simply recording the version of the CVMFS repo to afford additional opportunities for dedupliaction.
-Since each update to a repo does not change the contets of every (or even the majority of) files,
-data from multiple repo versions may well be able to coexist in a single container image.
-
-For two sets $A$ and $B$,
-the Jaccard distance $d_j$ is defined as follows.
-\[
-d_j(A, B) = 1 - \frac{|A \cap B|}{|A \cup B|} = \frac{|A \cup B| - |A \cap B|}{|A \cup B|}
-\]
-The Jaccard distance is a metric in the mathematical sense,
-making many intitive assumptions about distance and relationships between elements mathematically sound.
-For the purposes of comparisons,
-
-For this application in particular,
-the Jaccard distance captures several desirable properties of projections.
-In the case of two projections that differ only by one file,
-the Jaccard distance will be small.
-Likewise for a par of projections with nothing in common,
-the distance will be high.
-And since the Jaccard distance is a proper metric,
-related projections will tend to occur in clusters;
-if two projections are both found to be extremely close to a third projection,
-we can conclude that the two are themselves close by the triangle inequality.
-Other choices of metric may well be viable,
-but we found the Jaccard distance to be both well-studied and amenable to efficient approximate impelementation.
-As described in section ??,
-we used MinHash~\cite{minhash} as a constant-time approximation of the Jaccard metric as a first pass at selecting similar images.
-Such an approximation is important in practice due to the sizes of the data involved;
-metadata listings alone (not including any file data) consumed multiple gigibytes of storage in some of the container images we evaluated.
-
-\subsection{Image Management as an Online Problem}
-
-\begin{figure*}[t]
-\centering
-\fbox{\includegraphics[width=0.95\textwidth]{drawings/system_architecture.pdf}}
-\caption{Architecture of a Site-wide Image Management System}
-\label{figure:sys-arch}
-\end{figure*}
-
-Combining the tools and techniques introduced so far,
-we are now ready to address the central problem in managing projected images as a site-wide service.
-Rather than relying on administrators to manually manage images,
-we must define a method for making a decision online as to how to efficiently satisfy the dependency requirements for batch jobs.
-Our method must be suitable for inclusion as part of the setup for batch job scheduling and must be efficient in terms of both computation time and storage space.
-We will extend the initial implementation discussed in section ?? by tracking distances between projections and allowing existing images to be augmented rather than creating new ones for each request.
-\Cref{figure:sys-arch} gives an overview of the structure of this system.
-
-Upon receiving a request for a projection,
-we first hash the new projection according to the MinHash algorithm.
-We always store the hash values computed for every projection we build to allow for fast clustering.
-Next, we use MinHash to approximate the Jaccard distance between the new projection and all projections previously built.
-This computed distance can be cached to ensure that we estimate the distance between each pair of images only once.
-Since comparisons using MinHash take constant time for a given probability bound independent of the size of the sets involved,
-we can quickly identify cached projections that are similar to the new request.
-
-To decide if two projections are ``close'',
-we define the parameter $\alpha$ as the maximal Jaccard distance between closely related projections.
-Since Jaccard distance is by definition between zero and one,
-$\alpha$ must be in the same range.
-This $\alpha$ parameter is something like the ``globbiness'' of the system.
-Small choices of $\alpha$ requires that projections be extremely close before considering them for merging.
-In the extreme case with $\alpha = 0$,
-only \emph{identical} images will be considered close so no images will be merged.
-This results in a larger number of independent images.
-Choosing $\alpha$ to be larger makes it more likely for images to be considered similar and merged.
-This results in more augmented images that serve multiple tasks.
-In the extreme case of $\alpha = 1$,
-\emph{every} pair of images is considered close and merged if possible.
-This results in large container images that have accumulated many projections.
-We explore tradeoffs when choosing $\alpha$ and give recommendations based on observed patterns of activity in section ??.
-
-Having selected a set of candidate projections that are close enough to the new request,
-we next check each in turn to find either a superset of the requested projection or one without version incompatibilities.
-We first sort the list of candidate images by Jaccard distance.
-This ensures that our system prioritizes candidates that are similar in size to the requested image.
-In the case of finding a superset,
-we can simply use the corresponding cached container image.
-Presumably this superset was created by the merging process below,
-so we are done.
-Otherwise, we continue to look for a compatible projection
-As discussed in section ??,
-two projections are incompatible if they both include the same path but point to different data.
-This situation can arise if for example a library in CVMFS is updated in place.
-Old and new projections would include an identical filesystem path,
-but the new version would have different contents.
-This situation can arise in more subtle ways,
-for example with symlinks controlled by environment variables.
-Since projections of CVMFS must statically resolve such links,
-incompatibilities can arise despite paths appearing compatible under CVMFS.
-Any candidate projections that are incompatible with the new request are removed from further consideration.
-
-If no compatible projections were found among the candidates,
-we create a container image for the projection as requested and add it to the cache.
-Otherwise, on finding a compatible cached projection the next step is to augment it with the contents of the new request.
-We choose the first compatible image from the list of candidates for merging,
-again prioritizing images that are similar in size and contents.
-As previously discussed,
-Shrinkwrap can update a projection that was previously created;
-since we keep the projection used to generate each container image in the cache,
-we can simply augment the selected projection with the new files from the request.
-Next, we build a container image based on the newly updated projection.
-The computational cost of this step could become high since we are creating larger combined projections.
-Appropriate choice of $\alpha$, however,
-bounds the distance between the combined projections such that the cost of creating the augmented container is arbitrarily close to cost of simply fulfilling the original request.
-Since the augmented projection is the union of two different projections,
-we can substitute it for both.
-The container image for the augmented projection can replace the previous image,
-so that the net increase in storage space comes only from the newly added files.
-Since a single augmented image now serves two different batch job submissions,
-we avoid making two copies of the common files and data.
-
-One potential issue with this merging strategy is that there is no way to prevent poor merging decisions.
-In the case of a one-off submission,
-the system might merge an existing, frequently used image.
-This is not problematic for the one-off job itself,
-but increases overhead for subsequent jobs.
-The original, frequently used image would increase in size with no additional benefit to other tasks.
-Images might become ``bloated'' in this way,
-accumulating infrequently used dependencies and increasing overhead indefinitely for future tasks.
-The Jaccard distance metric gives a natural way to capture and quantify this effect.
-As an image becomes bloated due to repeated merges,
-its distance from every individual request increases.
-After sufficient growth,
-the image will become too far from any request to even be considered.
-Without regular use,
-the bloated image will eventually be removed in LRU order.
-Choice of $\alpha$ allows an upper limit on the amount of useless bloat in images.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{curated/time-series/time-series.pdf}
-\label{fig:series}
-\if 0
-zoomed out?
-cache limit, deletes start
-merge tracks writes, it's the dominant factor
-hits grow gradually as cache merges and fills
-deletes help prune useless stuff, remaining should be frequent hits
-far ends are either insert or merge, middle alphas have both
-\fi
-\end{figure}
-
-Figure~\ref{fig:series} shows a single simulation of the image management system,
-with $\alpha=0.0.75$, cache size XXX, ...
-First, we note that most of the operations are merges.
-This makes sense, as this simulation was at fairly high $\alpha$.
-The total bytes written also closely tracks merges,
-indicating that merging is the dominant source of I/O.
-We can still see inserts over the course of the simulation.
-At more extreme $\alpha$ values,
-we would expect to see one of these operations almost exclusively.
-As the data in cache continues to rise,
-we eventually see the cache limit,
-after which the delete count starts increasing.
-Over the course of the simulation,
-inserts and deletes are filling and emptying the cache such that it remains close to its storage limit.
-We also observe the number of cache hits continue to rise despite deletions.
-As we will see, merging allows for a greater proportion of hits even if the amount of data remains constant.
-This is because frequently used data can be merged,
-reducing duplication.
-The cache limit then ensures that infrequently used parts are eventually removed.
-
-\section{Evaluation}
-
-To evaluate the viability of our approach in automatically deduplicating a container image store,
-we simulated the behavior of the system over time for a range of $\alpha$ values.
-Choice of $\alpha$ is important here to ensure that common components can be detected and merged in container images,
-and that old images or poor decisions are eventually removed from the system.
-Choosing $\alpha$ too far in either direction will result in excessive storage overhead due to duplicated components,
-or excessive compute and network overhead repeatedly merging largely unrelated images.
-Our goal in this evaluation is therefore to choose $\alpha$ so as to minimize the storage and compute costs associated with maintaining a collection of images.
-
-Since our simulation uses random simulated requests,
-there is noticable variability between individual simulations.
-Thus for a given choice of cache size, job count, etc.
-we repeated the simulation 20 times and reported the median behavior over the runs.
-At each choice of $\alpha$ (in steps of 0.05)
-we performed a set of 20 simulated runs,
-allowing us to plot various measurements of the system versus $\alpha$.
-
-Sweeping over the range of $\alpha$ values in this way,
-we can immediately see differences in the frequency of simulated operations.
-Figure~\ref{fig:opcounts} shows the upper range of $\alpha$ values where behavior differences appear.
-From the lower values on the left,
-the insert and delete counts are the primary (or only) operations,
-with number of hits relatively constant.
-This corresponds to a simple LRU-based cache.
-The insert count is higher due to cache filling,
-but the two tend to increase in lockstep.
-As $\alpha$ increases,
-image merges become more probable.
-The merge count in the figure steadily increases throughout most of the upper range,
-while inserts and deletes decrease.
-This suggests that at high $\alpha$ values,
-the cache space would be more efficiently used,
-with some of the duplication merged out.
-In the extreme case with $\alpha=1$,
-every request is merged into a single image in cache,
-hence the sudden increase in hit rate and decrease in total merges at the far right of Figure~\ref{fig:opcount}.
-Note that as merging increases,
-the proportion of hits generally also increases.
-This indicates that more of the requests can be satisfied by previously merged images in the cache.
-From the perspective of the system,
-this is beneficial as there is no extra time or compute cost for handling a larger proportion of images.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{curated/comparative/operation_count.pdf}
-\label{fig:opcounts}
-\if 0
-inserts and merges switch as alpha goes up
-more hits with merging
-allows better responsiveness? image is immediately ready
-inserts and deletes go to zero, so storage use will taper off
-drop off: we keep everything, so 500 requests get merged; hit count goes up correspondingly since everything is in cache
-not gonna be close to cache limit
-inserts and deletes track: corresponds to LRU, inserts while filling then balanced
-\fi
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{curated/comparative/distribution_efficiency.pdf}
-\label{fig:dist-eff}
-\if 0
-explain what happened
-metrics: why we care about efficiency?
-cache efficiency: gauge duplication (waste), keeping working set costs lots of storage
-container efficiency: unexpected overhead for each request; compute, network, etc.
-place limits based on site, conditions
-limits exist, we won't tell you
-maybe just buy more cache?
-large scale (10-100 TB), no can do...so need to merge sometimes
-why 100 and 20 on the left
-why they switch places: extreme cases
-strike a balance
-one curve might be more important for you
-effects are not uniform/balanced gentle drop, massive spike in overhead on right
-\fi
-\end{figure}
-
-\subsection{Determining a Viable Range for $\alpha$}
-
-When sweeping over the range of $\alpha$ values,
-there are a large number of metrics available to summarize each simulation run.
-Many, however, are highly coupled with the particular system configuration and difficult to compare as we vary the paramaters of the simulations.
-
-We therefore chose to define two metrics, cache efficiency and container efficiency,
-to indicate the relative utilization of the system.
-Figure~\ref{fig:dist-eff} plots these two metrics for a representative simulation.
-We defined cache efficiency as the ratio of unique data to total data in the cache.
-In our case, this is equivalent to the ratio of the size of the unique packages to the total cache size.
-If many images contain copies of the same packages,
-the cache efficiency decreases.
-This metric captures the extent of duplication within the cache across all images.
-With no merging there is a high degree of duplication,
-so the cache efficiency is low.
-On the other end of the spectrum,
-maintaining a single, large image containing all data results in cache efficiency of 100\%,
-as the entire cache resides in that single image with no duplication
-
-We defined container efficiency as the ratio of the size of the requested container
-(in our case a set of requested packages plus all dependencies)
-to the size of the container the system actually used for the job.
-In the absence of merging,
-these two are equal so the container efficiency is 100\%;
-jobs are run with exactly what was requested.
-By merging to allow for image reuse,
-we include additional, unrequested data in container images.
-The container efficiency measures this difference between requests and containers.
-In the extreme case of $\alpha=1$ with a single large image, for example,
-the container efficiency is poor because the entire repo is used for every request,
-regardless of size.
-
-These two extreme cases, no merging among many images and a single merged image,
-can both be useful in some situations.
-If the requirements of all jobs are known in advance and worker storage and bandwidth are sufficient,
-maintaining a single, large image for all jobs on all workers is an useful strategy.
-After the initial (costly) image creation and transfer,
-there are no continuing costs or delays.
-When resources are limited or requirements change regularly,
-however, this approach becomes prohibitively expensive.
-The high compute and bandwidth overhead apply for every image update,
-which in the worst case could be every job.
-
-The other extreme situation,
-simply caching requests with no merging,
-can also be viable.
-This approach is simpler and does not come with extreme per-job overhead.
-In addition, frequent changes to the jobs and requirements do not degrade the efficiency of each job.
-At large scale, however,
-the overall system efficiency suffers.
-Due to duplication among images,
-larger caches contain less and less unique data.
-Thus to support a given repo,
-it can become necessary to provision a cache much larger than the size of the repo.
-With common CVMFS repos consuming several terabytes of storage each,
-the amount of cache space used can quickly grow.
-
-These two limiting factors,
-the compute and transfer cost in the highly merged case and the cache efficiency in the unmerged case,
-serve as limits on the viable range of $\alpha$ values for the system.
-Figure~\ref{fig:dist-eff} shows two vertical lines indicating these limits on a plot of efficiencies.
-The left line shows a lower limit of around 30\% on the cache efficiency.
-Choosing $\alpha$ too low results in cache efficiency reaching below 20\%.
-On the right, the line gives an upper limit on the likelihood of merging.
-As shown in Figure~\ref{fig:opcounts},
-the amount of I/O and compute to update images becomes much larger if $\alpha$ is set too high.
-These two limits define a range of viable values for $\alpha$.
-There is no general rule for the placement of these limits,
-which depends strongly on the performance characteristics of the execution environment,
-as well as the priorities of the administrators in optimizing the system.
-Rather than defining where these limits fall,
-we give per-site considerations in choosing limits in Section ??.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{curated/comparative/cache_efficiency.pdf}
-\label{fig:cache-eff}
-justification for cache efficiency
-unique vs total
-full x range?
-merging can give more unique data in less cache space
-\end{figure}
-
-Figure~\ref{fig:cache-eff} shows the actual data sizes used to calculate the cache efficiency.
-Without merging, the unique data makes up a small proportion of the cache.
-With increased merging, the amount of unique data increases.
-For sufficiently high $\alpha$,
-merges occur more frequently than cache inserts,
-resulting a drop in total storage size.
-On the far right of Figure~\ref{fig:cache-eff},
-the entire cache contents are merged into a single large image,
-so that the unique and total data sizes become equal.
-The cache efficiency metric is affected both by the increase in unique data and by the decrease in total cache size.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{curated/comparative/container_efficiency.pdf}
-\label{fig:container-eff}
-\if 0
-related to container efficiency: not exact but shows it
-too much jitter from individual requests, so cumulative
-crossover: hit rate allows reuse, so slightly less bytes written than requested
-requested is based on random stream, constant
-actual written goes up fast with alpha, so more overhead
-proxy for compute/network/etc.
-really steep at high alpha
-\fi
-\end{figure}
-
-Figure~\ref{fig:container-eff} shows the cumulative costs of creating and updating images relative to the cost of simply providing the requested images.
-These costs illustrate the effect of container efficiency over the course of a simulated run.
-Without merging, the actual I/O and compute costs in the system closely follow the requests.
-In Figure~\ref{fig:container-eff} the actual I/O is slightly smaller than the requested amount due to caching:
-the system can sometimes reuse images without performing any I/O.
-As $\alpha$ increases, the cost of updating and merging images comes to dominate the total cost.
-Administrators migh decide, for example,
-to spend at most twice the compute and I/O time relative to the the actual requested operations.
-This would correspond to the upper compute limit shown on \label{fig:dist-eff}.
-Depending on the priorities at a particular execution site,
-either the storage utilization or compute and I/O time could be the critical factor in optimizing performance.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{curated/comparative/distribution_efficiency_comp.pdf}
-\label{fig:comp-eff}
-\if 0
-different data patterns cause different effects
-random has high utilization
-pushed right, limited opportunity for merge, since no overlap
-merging doesn't help
-random is not realistic though
-merging is not general solution, but for trees of software, look at this
-
-tree has more duplication, starts merging much sooner
-closure is very non-controvertial, but results in serious changes in efficiency
-common to any software, so opportunity for optimization
-common core that can merge to avoid dups
-including usage stats (dist) amplifies these somewhat
-real usage should be stronger than dist
-
-use dist from now on because reasonable approximation to actual usage
-similar to tree anyway
-mild but noticable difference from tree, more natural approximation to real usage
-\fi
-\end{figure}
-
-We also compared the efficiencies of several types simulated images under a range of $\alpha$ values.
-For the previous discussion, synthetic images were generated taking into account both the tree of package dependencies and usage statistics.
-Figure~\ref{fig:comp-eff} shows a representative simulation with two other synthetic image types included.
-First, we measured images considering only the dependency relationships but not usage.
-These images show similar behavior to those previously discussed,
-albeit with slightly lower efficiencies.
-More realistic images primarily make the same trends more pronounced.
-This suggests that even with random, non-realistic requests,
-the tree structure of package dependencies is sufficient to produce duplication in the cache.
-The other type of synthetic image shown in Figure~\ref{fig:comp-eff} is purely randomly generated.
-The set of packages in each image was chosen uniformly randomly from the full SFT repo.
-These random images do not take into account usage statistics or package dependencies,
-making the software in them entirely nonfunctional.
-For the random images,
-we observe markedly different cache behavior.
-The cache efficiency is much higher,
-indicating a greater proportion of unique data in the cache.
-Since the image contents were selected purely at random,
-this is expected.
-The container efficiency is also high over a much larger range of $\alpha$ values.
-This is because there is effectively no merging until extremely high $\alpha$.
-Since image contents are purely random,
-there is no correlation between different images.
-Thus, it is much more difficult to find images similar enough to merge until the $\alpha$ value is very lax.
-This would indicate that our merging strategy is not applicable to arbitrary collections of data.
-Random images show little to no effect for most $\alpha$ values.
-Our merging strategy, which takes advantage of duplicated content included as a result of dependencies in software,
-would be ill advised for situations that are not known to follow similar patterns of duplication.
-
-\if 0
-\subsection{Quantifying Image Management Costs}
-
-The next step in evaluation is to define metrics describing the collection of cached containers.
-Here we will define metrics in terms of discrete software components/packages to match the simulated streams,
-but the same definitions would apply equally well at the granularity of individual files.
-We will suppose that there is a collection of requested container images,
-from either our simulated stream or from real user requests based on job analysis and/or dynamic tracing.
-It is the responsibility of the image management system to provide a container image satisfying the requirements of each request.
-Let the set $C$ be the union of all components in one or more requests.
-$C$ provides a lower bound on the amount of space required to serve all requests at the same time.
-If the storage space required to hold all of $C$ is greater than the total available storage,
-then images serving new requests will requre removal of container image(s) for previous requests.
-Since we have been considering image management as an online problem,
-we cannot place an upper bound on the sizes of requests and thus require some caching policy
-(in our case LRU) to ensure that $C$ is finite and fits available storage.
-The large, full-repo images tested at NERSC correspond to single images containing all of $C$.
-The image size is very large and any changes/updates are expensive both in terms of compute and bandwidth.
-NERSC's approach does, however,
-avoid duplication of data as there is a single image that serves all users and applications.
-
-We thus define the multiset $A$ as the union (with repetition) of all components in a set of requested images.
-If multiple images request the same component,
-it will occur multiple times in $A$.
-The size required to store $A$,
-including all duplicated elements,
-is the storage space that would be required if a separate image were created for each job request.
-$A$ captures the amount of (potentially) unnecessary duplication in the collection of images.
-The initial approach described in section ?? would have a storage cost of $A$.
-Note that it is always the case that $|C| \leq |A|$,
-and if $|C| = |A|$ then there are no common components among the requested container images.
-While large, full-repo images provide the best storage efficiency, being based solely on $|C|$,
-they are expensive and unweildy to use in practice.
-The initial approach in section ?? is extremely simple and friendly to use in practice,
-but its storage cost is determined by $A$.
-Our goal, therefore, in evaluating our improved image management scheme detailed in section ??
-is to achieve storage cost as close to $C$ as possible without excessive overhead in terms of compute and network bandwidth to keep the image collection up to date.
-We define \emph{cache efficiency} as $|C| / |A|$ to quantify the amount of duplication in the system cache.
-With poor cache efficiency,
-the available storage for the system cache is wasted on repeated copies of the same data.
-Improving cache efficiency allows for more more unique image data within the same storage space,
-or equivalently allows the same set of job requests to be served efficiently from a smaller cache.
-\fi
-
-\subsection{System-specific Effects of $\alpha$ Threshold}
-
-\begin{figure*}
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_cache.pdf}
-\hfill
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_cache.pdf}
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_request.pdf}
-\hfill
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_request.pdf}
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_uses.pdf}
-\hfill
-\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_uses.pdf}
-\label{fig:sensitivity}
-\if 0
-[t] "just buy more cache"
-efficiency is worse, since cache holds on to lots of questionable data
-loads of opportunities to try merges, so we see a lot of unnecessary merges
-less force removing bad choices, so more cruft can accumulate in many images
-better cache utilization with tiny cache, but of course more thrashing
-at scale, bad idea
-large cache is too pricey at scale
-[m] varying reuse doesn't affect too much
-below 100 doesn't seem full/steady state yet
-cache isn't settled? so haven't dropped bad merges; more unique data in cache boosts efficiency
-[b] high number of requests, repeats don't matter
-unique data dominates
-hard to tell difference between [m] and [b], so not too sensitive to these
-once you hit steady state, jobs count/repeats has little effect on efficiency
-general trends of merging, not interesting by themselves
-\fi
-\end{figure*}
-
-In Figure~\ref{fig:sensitivity}, we plot efficiency curves for a range of simulation conditions.
-The left column shows container efficiency,
-while the right column shows cache efficiency.
-In the first row, the number of jobs and the amount of repetition are constant while the cache size is varied.
-In the second row, the number of unique requests is varied while holding the other parameters constant,
-and the third row varies the number of times each job is submitted while holding the rest constant.
-
-The size of the cache has an inverse relationship with both the container and cache efficiency.
-A larger cache can of course hold a larger number of images,
-but since each image contains significant duplicated portions,
-increasing cache size tends to decrease cache efficiency.
-Conversely, small caches more quickly evict images so that ineffective merges and similar images tend not to remain in cache too long.
-A larger cache also allows for more opportunities to merge images,
-leading to decreased container efficiency.
-When deciding how to handle a request,
-a large cache full of images is much more likely to contain an image suitable for merging.
-With a small cache, opportunities to merge are much more dependent on the order in which requests happen to come.
-
-The effect of varying the number of unique jobs is less pronounced than the effect of cache size.
-Streams of 500 and 1000 unique jobs show nearly indistinguishable behavior,
-indicating that by 500 jobs the system has reached a steady state.
-Continuing with an arbitrarily long stream should not result in significant performance changes.
-It appears that 100 unique jobs were not sufficient to fill the cache and reach a steady state.
-In this case the container efficiency is somewhat decreased over a range of $\alpha$ values,
-suggesting that some ineffective merges had not made their way out of the cache yet.
-The cache efficiency in this case is slightly increased.
-This would suggest that before reaching a steady state,
-the contents of the cache are more assorted and some unnecessary data remain in cache.
-
-The effect of varying the number of times each unique job is submitted appear minimal at best.
-As discussed with the previous row,
-a stream of 500 unique images should be sufficient to reach a steady state.
-For a sufficiently large cache, therefore,
-it appears that the efficiency is determined by the amount of unique data and fairly insensitive to how much those pieces of data are used.
-This simulation is somewhat unrealistic in that each job is repeated a uniform number of times.
-In real systems, we expect more variation in the frequency of repeated jobs.
-We predict that a steady, non-uniform distribution of frequencies would result in the same behavior as shown in the bottom row of Figure~\ref{fig:sensitivity},
-while changes in the frequencies of jobs (e.g. a new user starting an experiment)
-would disturb the steady state and produce effects similar to those in the middle row.
-
-\if 0
-Using these metrics for cost and overhead,
-we carried out an evaluation of our container image management scheme as follows.
-First, we somewhat arbitrarily chose a storage limit of 1~TB for our simulated image cache.
-This is not an unreasonable storage cost for an HPC site;
-for the test setup in place at NERSC,
-the full repo images already consume multiple terabytes per worker.
-We experimented with different storage limits around the same order of magnitude,
-but did not find marked differences in the behavioral trends observed below.
-The important aspect is that some definite limit exist.
-This does not take into account the storage used by Shrinkwrap,
-which is automatically deduplicated,
-nor does it take into account the metadata cost of creating projections.
-
-The next step was to generate a synthetic stream of requests for container images.
-These were generated using the procedure described previously.
-To ensure that both the merging strategy and the caching policy come into play,
-and to try to capture the steady state behavior of the system,
-we continued the stream until the total amount of unique data handled was twice the storage capacity.
-Equivalently, the request streams were long enough such that $C$ for the entire stream exceeded 2~TB.
-We then tabulated the duplication ratio $d$ for the image cache,
-as well as the total amount of data written over the course of the simulation.
-Our goal is to minimize both duplication and data written,
-which will be proportional to the compute and bandwidth cost of creating and updating container images.
-We repeated this simulation W times and computed the averages and standard deviations of these metrics.
-We repeated this procedure for a range of $\alpha$ thresholds in the range 0 to 1,
-with the results shown in figure ??.
-
-Looking at the results in figure ??,
-we can see that small values of $\alpha$ result in the highest level of data duplication,
-with $d$ decreasing as $\alpha$ increases.
-Since $\alpha=0$ corresponds to the initial approach of creating a unique container image for each job submission,
-we expect to see the highest level of duplication there.
-Looking at the cost in terms of write activity,
-we see that large values of $\alpha$ result in the most write I/O,
-and thus compute and bandwidth overhead.
-Since $\alpha=1$ allows all requests to be merged into a single large image,
-we expect minimal duplication.
-This configuration, however,
-has substantial overhead for image creation and maintenance.
-This is a real complaint from administrators at NERSC testing full-repo images.
-The high cost of creation means that such large images can only be infrequently created or updated.
-\fi
-
-\subsection{Steps to Determining $\alpha$}
-TODO
-
-\section{Related Work}
-
-\subsection{CVMFS}
-The CernVM File System (CVMFS) is a globally distributed filesystem designed for providing efficient,
-read-only access to scientific software.
-CVMFS is primarily used for distribution of High Energy Physics (HEP) software used as part of the Large Hadron Collider (LHC) at CERN.
-Researchers at CERN use CVMFS as the primary means of distributing the analysis and simulation software they develop to the Worldwide LHC Computing Grid (WLCG).
-
-Each WLCG site commits to providing some fixed amount of computational resources for analyzing particle collision events and running simulations.
-Worker nodes at each site must be configured to mount CVMFS as a prerequisite to performing any computational tasks.
-Since CVMFS is implemented as a FUSE module,
-WLCG sites are expected to support FUSE on their worker nodes.
-CVMFS also relies on global configuration files,
-requiring administrative involvement in preparing worker nodes for WLCG jobs.
-In addition, sites generally provision some infrastructure to support CVMFS,
-such as caching proxies to reduce load created by local workers on publicly accessible filesystem servers.
-
-As a result, it is difficult to contribute computational resources toward the LHC experiments without making some site-wide provisioning and configuration changes.
-This is not an issue for the WLCG sites,
-as they provide dedicated computing resources and can customize their setups to meet the requirements of CVMFS.
-For a professor attempting to harness a campus cluster,
-on the other hand, these requirements present a serious barrier.
-Solutions exist such as
-Parrot, a debugging tool and personal filesystem developed at the University of Notre Dame,
-which sidesteps FUSE but still requires unrestricted network access and
-UnCVMFS which locally store all of CVMFS to avoid network access, 
-but relies on large storage and permission mounting.
-For computing at scale, it is important to provide a caching proxy and concise storage,
-which currently require assistance from administrators.
-
-\subsection{HPC}
-
-With the rapidly increasing demands of the LHC experiments at CERN,
-HPC resources are an appealing source of computing power to supplement the WLCG.
-Leadership class machines such as Summit, at Oak Ridge National Laboratory, will by themselves be able to perform a substantial portion of the work of the entire WLCG today.
-Use of national-scale HPC resources across the United States and Europe such as Cori at NERSC, Sierra at Lawrence Livermore National Laboratory, Piz Diant at Swiss National Supercomputing Centre could be key in meeting the computational demands following the LHC's high-luminosity upgrade.
-
-Unfortunately, HPC sites are generally more restrictive than the previously mentioned campus resources.
-In addition to limited privilege on workers and lack of FUSE support,
-HPC sites often impose restrictions on network activity,
-including outbound firewalls preventing access to CVMFS.
-
-The US collaboration of the ATLAS project is currently taking advantage of computing resources at various supercomputers in the United States including Cori at NERSC.
-For the deployment of their software the \texttt{uncvmfs} utility is used to unpack entire CVMFS repositories into standalone images (usually in ext4 or squashfs format) which in turn can be used to build Shifter or Singularity images.
-While these images can very well be scaled out onto a large number of nodes inside the NERSC infrastructure,
-the image build process takes around 24 hours which makes it difficult to deploy up-to-date versions of the software on a regular basis.
-As these images contain complete copies of everything in CVMFS,
-each takes on the order of terabytes,
-necessitating careful storage management.
-In addition, the process requires administrators for image creation, deployment, and cleanup.
-Currently, this approach is used in production on the ATLAS and CMS experiments.
-As additional experiments want to take advantage of the resources at NERSC,
-the administrative burden of managing multiple CVMFS images on multiple software versions increases accordingly.
-For the experiments to use the same images (saving time and cost at NERSC),
-multiple groups would have to come to agreement about a single common set of software versions to use and adjust their pipelines elsewhere.
-Both solutions introduce new inconveniences and administrative burden.
-
-More broadly, the use of static, manually managed container images breaks user assumptions about the operation of CVMFS.
-When running jobs using CVMFS,
-researchers can specify the version of each repo to use,
-allowing precise control over software and dependencies.
-Each experiment has its own policies and procedures for choosing a software version;
-some may use bleeding-edge analysis software or another experiment may prefer older, stable versions.
-CVMFS is able to provide multiple customized views of the global filesystem without incurring much additional cost.
-CVMFS also supports some more exotic features,
-such as symlink resolution controlled by environment variables.
-CVMFS as a whole does not map well to the semantics of a traditional filesystem.
-Projecting CVMFS into such a filesystem can capture at most a single, static version of a repo and must necessarily eliminate some features,
-such as environment-based symlink resolution at runtime.
-
-\if 0
-% https://indico.cern.ch/event/587955/contributions/2937411/attachments/1683156/2705423/WahidBhimji-CHEP18-NERSC.pdf
-Another similar approach in use at NERSC is DVMFS.
-Since creating static container images containing CVMFS is costly and difficult to customize,
-DVMFS instead uses Cray DVS IO forwarders to mount a copy of CVMFS served over NFS.
-This allows for less administrator involvement,
-since the mirrored copy of CVMFS can be automatically updated on a regular basis.
-After dealing with several implementation problems,
-DVMFS is currently deployed at NERSC as an alternative to static images.
-While this approach affords more flexibility in choice of repos and software versions,
-it increases the job startup time and places additional requirements on the infrastructure.
-To be usable in production, DVMFS requires a sufficient number of DVS servers to be provisioned.
-DVMFS is nonetheless a viable solution as long as the costs of developing/maintaining the service and allocating infrastructure do not grow too high.
-\fi
-
-\section{Conclusion}
-TODO
-
 \section{Reproducibility Data}
@@
 and instructions on how to build and run them.
-
-\section*{Acknowledgment}
+\fi
+
+%\section*{Acknowledgment}
 
